### 2020-03-31 latexnotes/notes.tex
initial commit

\documentclass[11pt]{amsart}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{todonotes}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{problem}{Problem}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{example}{Example}
\newtheorem*{solution}{Solution}

\title{Prisoner's Dilemma}
\author{Arvid Lunnemark}

\begin{document}
\maketitle

\begin{definition}
  The \textit{prisoner's dilemma} is a symmetric two-player game with two actions, cooperate ($C$) and defect ($D$), where, if player 1 plays $a$ and player 2 plays $b$, player 1 gets payoff 
  \todo{phrase this in a nicer way}
  \begin{equation*}
    p(a,b) = \begin{bmatrix}
      R & S \\
      T & P
    \end{bmatrix}
  \end{equation*}
  if player 1 is the row player, and cooperate is the first action. We have $T > R > P > S$, and typically, we have the concrete values $T = 5$, $R = 3$, $P = 1$ and $S = 0$.
  \todo{(could also think about including the $2R > T + S$ condition here)}
\end{definition}

\begin{definition}
  A \textit{strategy} is a Moore machine (finite automaton with outputs) over the input and output alphabet $\{C, D\}$, with probability $1-p$ of following the correct transition and probability $p$ of following the incorrect transition. 
  \todo{(could also define it as probability $p$ of outputting the wrong character, but should be largely equivalent)}
  The strategy can be viewed as a function $s$ from a history string $h$ of the opponents moves to an action, i.e., $s : \{C, D\}^* \to \{C,D\}$.
\end{definition}

\begin{definition}
  The \textit{expected time-average payoff} of strategy $s_1$ when played against strategy $s_2$ is \begin{equation*}
    v(s_1, s_2) = \lim_{T \to \infty} E \left[ \frac{1}{T} \sum_{t = 1}^T p(s_1(h_{t-1, 2}), s_2(h_{t-1, 1})) \right]
  \end{equation*}
  where the history is defined as \begin{equation*}
    \begin{split}
    h_{t, 1} = h_{t-1, 1} \circ s_1(h_{t-1, 2})  \\
    h_{t, 2} = h_{t-1, 2} \circ s_2(h_{t-1, 1}) 
    \end{split}
  \end{equation*}
  \todo{(TODO redo this. the actual definition i'm using is the distribution in the stationary state of the markov chain as explained below)}
\end{definition}

\begin{definition}
  A \textit{population} of strategies $P = (S, f)$ is a set $S$ of strategies and a function $f : S \to (0,1]$ such that $\sum_{s \in S} f(s) = 1$, representing the frequency of each strategy in the population.
\end{definition}

\begin{definition}
  The \textit{fitness} of a strategy $s$ in a population $P = (S, f)$ is \begin{equation*}
    F(s) = \sum_{s' \in S} f(s) v(s, s').
  \end{equation*}
\end{definition}

\begin{definition}
  A strategy $s_2$ can \textit{invade} a strategy $s_1$, if there exists a real number $\alpha \in (0,1]$, such that for all populations $P$ with $S = \{s_1, s_2\}$ and $f(s_2) < \alpha$, we have
  \begin{equation*}
    F(s_2) > F(s_1).
  \end{equation*}
  \todo{a question is what to do with evolutionary drift, that is, if $F(s_2) = F(s_1)$, becuase it could potentially pave the way for more aggressive strategies. NOTE this is much more complicated than I first thought. the way I defined it now says that a strategy is evolutionary stable even if it can be taken over 99\% by another strategy, which feels weird. on the other hand, the way I defined it now implies that an evolutionary stable strategy will always stick around, even in minority, but it probably could easily be replaced after being in the minority. on the other hand, the prevailing definition also does not preclude the possibility of being easily replaced after evolutionary drift has taken place, and one should need to defend against that too if one wants the ultimate evolutionary super stable strategy. I feel like super stable is impossible though because if you were drifted with allC and then invaded by allD allD would take over and the only way to protect against it would be to go allD on allC but that would definitely not work I think. }
\end{definition}
\begin{remark}
  Thus, $s_2$ can invade $s_1$ if and only if it can start as an infinitesimally small part of the population and grow to become a constant fraction $\alpha$ of it.
\end{remark}

\begin{definition}
  A strategy $s_1$ is \textit{evolutionary stable} if it cannot be invaded by any strategy $s_2$.
\end{definition}

\begin{corollary}
  A strategy $s_1$ is evolutionary stable if and only if \begin{equation*}
    v(s_1,s_1) > v(s_2,s_1)
  \end{equation*}
  or \begin{equation*}
    v(s_1,s_1) = v(s_2, s_1) \text{ and } v(s_1,s_2) \geq v(s_2,s_2)
  \end{equation*}
  for all $s_2$.
  \todo{apparently the last inequality is strict in the traditional definition. given what I've seen with Pavlov it may even be possible to have that stricter definition be in place}
\end{corollary}
\begin{proof}
  Yeah this is true, the proof is easy by examination of cases. I just don't want to write it now.
\end{proof}

\begin{theorem}
  Suppose a strategy $s$ is evolutionary stable. Then $v(s,s) = R$. In other words, $s$ is efficient with itself.
  \todo{This looks like a very similar result to the theorem presented in the original paper. In fact, it looks like merely a simplication of it. To the contrary, however, it is very different, which stems from the definition of $v(\cdot,\cdot)$ — it captures the entire behavior, the actual expected value, instead of some weird artificial value of pretending that no mistakes occur at all which doesn't really make sense. Therefore, I think that this definition of evolutionary stable and $v$ will even exclude TFT from being evolutionary stable, which is a highly interesting result actually. Not sure if it is true though because that would mean that there would be an invader for TFT but I think that that is definitely a possibility, and that this model is actually more similar to real life than all other models presented in the literature on this subject so far. nice.}
\end{theorem}

OK we need a different notion of what it means to be efficient with oneself now. \textit{Probably.} In fact, maybe we don't. Because AllC will get $R$. But neither TFT nor Pavlov will. And, in fact, any strategy that is not stupid, will have opposite behavior when $p = 0$ and when $p = 1$ which suggests that it cannnot possibly get $R$ in both situations. COULD MAKE IT ONLY WORK IN THE LIMIT AS $p \to 0$.

wait..... maybe allC can invade TFT now. seems to depend on the specific values of $R, S, T, P$ which is the worst possible situation one can end up in. I think AllC can invade Pavlov too, which is not very promising

hmm okay so Pavlov is better against itself than TFT against itself, which, however, we have seen doesn't matter much.

I think the proof strategy is like this: suppose $v(s_1,s_1) < R$. then create a strategy $s_2$ that is more or less exactly like $s_1$, so $v(s_2,s_1) = v(s_1,s_2) = v(s_1,s_1)$, but $s_2$ identifies itself and has $v(s_2,s_2) = R$.

I think that captures the idea of an outside species pretending to be the same as everyone else but secretly getting a lot of value from cooperating with other members of the infiltrators, which I think makes a lot of sense from the anthropomorphic viewpoint.

A reasonable condition is \begin{equation*}
  \argmax_v \left[ (R, T, S, P) \cdot v \right] = (1, 0, 0, 0).
\end{equation*}
I think this is needed and crucial. Maybe one should even require that it's the unique maximum but I'm not exactly sure. This also seems like a very reasonable thing to assume, and is a generalization of the commonly used $R \geq (S + T)/2$. I like this. omg wait this condition makes no sense at all.

If we require unique maximum, then TFT can be invaded by allC, which kinda makes sense.

NICEEEE: As $p$ tends to 0, Pavlov tends to characteristic vector $(1,0,0,0)$ (i.e. payoff $R$) so I actually think that this theorem might be true (especially considering that AllC fares poorly against Pavlov). Since TFT have characteristic vector $(1/4,1/4,1/4,1/4)$, and would thus not be evolutionary stable here, this theorem indicates a stronger result than the theorem that it is based on.

omg wait the condition I proposed makes no sense at all hmmmm

oh wow read https://search.proquest.com/docview/235762955?accountid=12492 they came up with exactly the same Markov chain as I did and seems to have done more analysis on it. they seem to not do much with their results at all, however.

\end{document}

### 2020-03-31 work.py

@@ -1,0 +1,25 @@
+#%%
+# imports
+import numpy as np
+import sympy as sym
+
+# %%
+# constants
+p = sym.Symbol('p')
+
+# %%
+# pavlov vs allC
+M = sym.Matrix([[(1-p)**2,    (1-p)*p,    p*(1-p),    p**2],
+              [p**2,        p*(1-p),    (1-p)*p,    (1-p)**2],
+              [p**2,        p*(1-p),    (1-p)*p,    (1-p)**2],
+              [(1-p)**2,    p*(1-p),    (1-p)*p,    p**2]])
+
+# %%
+sym.simplify(M**10)
+
+# %%
+M_pavlovVSallc = sym.Matrix([[1-p, p],[p,1-p]])
+
+sym.simplify(M_pavlovVSallc**10)
+
+# %%

### 2020-04-05 latexnotes/notes.tex
a major problem

@@
 
+\section*{A Major Problem}
+
+The definitions in the preceding section were all taken from the existing literature.
+In particular, the corollary is the traditional definition of evolutionary stability (and the underlying definitions were reversed engineered from that).
+However, these definitions lead to the following big problem: \textit{AllD can no longer be invaded}.
+The reason is simple: Suppose strategy S can invade AllD.
+In the stationary distribution, there can be no state of S that cooperates.
+Thus, one can contract all independent connected components into single nodes of D, reachable by different branches.
+One can then easily see that running S against itself also always defects, so it must effectively just be an AllD copy, and will thus not be able to invade.
+
 \end{document}

### 2020-04-09 realnotes/notes.tex
add better notes

\documentclass[11pt]{amsart}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{todonotes}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{problem}{Problem}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{example}{Example}
\newtheorem*{solution}{Solution}

\title{Prisoner's Dilemma}
\author{Arvid Lunnemark}

\begin{document}
\maketitle

\begin{definition}
  The \textit{prisoner's dilemma} is a symmetric two-player game with two actions, cooperate ($C$) and defect ($D$), where, if player 1 plays $a$ and player 2 plays $b$, player 1 gets payoff 
  \begin{equation*}
    p(a,b) = \begin{cases}
      R &\text{if $a = C, b = C$} \\
      T &\text{if $a = D, b = C$} \\
      S &\text{if $a = C, b = D$} \\
      P &\text{if $a = D, b = D$}
    \end{cases}
  \end{equation*}
  We have $T > R > P > S$, and typically, we have the concrete values $T = 5$, $R = 3$, $P = 1$ and $S = 0$.
\end{definition}

\begin{definition}
  A \textit{strategy} is a Moore machine (finite automaton with outputs) over the input and output alphabet $\{C, D\}$, with probability $1-p$ of following the correct transition and probability $p$ of following the incorrect transition. 
\end{definition}

Note: this models an error probability in \textit{perception}. One could also think of an error probability in \textit{outcome}, but it is easy to see that the two are equivalent up to a change of the values of $R, S, T, P$.

\begin{definition}
  Suppose strategy $s_1$ plays strategy $s_2$. This defines a \textit{$s_1$-$s_2$ graph} which is a Markov chain where each node represents a pair of states $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$. The transition probabilities are defined in the obvious way.
\end{definition}

\begin{definition}
  Let $\pi$ be the stationary distribution achieved by starting in the start state of the \textit{$s_1$-$s_2$ graph}. The \textit{payoff} of strategy $s_1$ when played against strategy $s_2$ is \begin{equation*}
    v_{s_1}(s_2) = \sum \pi_{c_1, c_2} \cdot p(c_1, c_2).
  \end{equation*}
\end{definition}

Note: The graph might be periodic in which case we we will not get a stationary distribution. I need to think about this special case but my intuition is that it shouldn't matter.

\begin{definition}
  A \textit{population} of strategies $P = (S, f)$ is a set $S$ of strategies and a function $f : S \to (0,1]$ such that $\sum_{s \in S} f(s) = 1$, representing the frequency of each strategy in the population.
\end{definition}

\begin{definition}
  The \textit{fitness} of a strategy $s$ in a population $P = (S, f)$ is \begin{equation*}
    F(s) = \sum_{s' \in S} f(s') v_s(s').
  \end{equation*}
\end{definition}

\begin{definition}
  A strategy $s_1$ is \textit{$\epsilon$-invadable} if there exists a strategy $s_2$ such that in all populations $P$ with $S = \{s_1,s_2\}$ and $f(s_2) \geq \epsilon$, we have \begin{equation}
    \label{fitnesscond}
    F(s_2) > F(s_1)
  \end{equation}
\end{definition}

\begin{definition}
  A strategy $s_1$ is \textit{evolutionary stable} if there exists an $\alpha \in (0,1)$ such that for all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
\end{definition}

Note: it is easy to see that this is just equivalent to saying that there exists some $\epsilon$ for which $s_1$ is not $\epsilon$-invadable.

\begin{theorem}
  Suppose $s_1$ is evolutionary stable. Then $v_{s_1}(s_1) \geq \frac{S + T}{2}$.
\end{theorem}

\begin{theorem}
  Suppose $s_1$ is evolutionary stable as $p$ goes to 0 (i.e., that it is evolutionary stable if condition \ref{fitnesscond} is replaced by $\lim_{p \to 0} (F (s_2)-F(s_1)) > 0$). Then $v_{s_1}(s_1) = R$. In other words, $s_1$ is utilitarian.
\end{theorem}

\begin{theorem}
  The Pavlov strategy, displayed in Figure ??, is evolutionary stable as $p$ goes to 0.
\end{theorem}

\begin{remark}
  TFT, displayed in Figure ??, is not evolutionary stable as $p$ goes to 0. It has the stationary distribution $(1/4,1/4,1/4,1/4)$ which is smaller than $R$.
\end{remark}

\begin{figure}
  \includegraphics[width=4cm]{tft.jpg}
  \centering
  \caption{TFT.}
  \end{figure}
\begin{figure}
  \includegraphics[width=4cm]{pavlov.jpg}
  \centering
  \caption{Pavlov.}
  \end{figure}

\end{document}

### 2020-04-23 realnotes2/notes.tex
modified theorems

--- 2020-04-09 realnotes/notes.tex
@@
 \usepackage{todonotes}
+\usepackage{cleveref}
 
@@
 \begin{definition}
-  Suppose strategy $s_1$ plays strategy $s_2$.
-This defines a \textit{$s_1$-$s_2$ graph} which is a Markov chain where each node represents a pair of states $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$.
+  Suppose strategy $s_1$ plays against strategy $s_2$.
+This defines an \textit{$s_1$-$s_2$ graph} which is a Markov chain where each node represents a pair of states $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$.
 The transition probabilities are defined in the obvious way.
@@
 
-Note: The graph might be periodic in which case we we will not get a stationary distribution. I need to think about this special case but my intuition is that it shouldn't matter.
+Note: The graph might be periodic in which case we we will not get a stationary distribution — in that case, however, because of the linearity of the payoffs, we can replace the see that the periodic distribution gets the same time-average value as the stationary distribution that is the mean of the distributions of the periodic end state.
 
@@
 \begin{definition}
-  A strategy $s_1$ is \textit{evolutionary stable} if there exists an $\alpha \in (0,1)$ such that for all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
+  A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\alpha$, both in $(0,1)$, such that for all $p < p_0$, and all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
 \end{definition}
 
-Note: it is easy to see that this is just equivalent to saying that there exists some $\epsilon$ for which $s_1$ is not $\epsilon$-invadable.
+Note: it is easy to see that this is just equivalent to saying that for all $p < p_0$, there exists some $\epsilon$ for which $s_1$ is not $\epsilon$-invadable.
 
 \begin{theorem}
-  Suppose $s_1$ is evolutionary stable. Then $v_{s_1}(s_1) \geq \frac{S + T}{2}$.
+  \label{evolutionarystable1}
+  Suppose $s_1$ is evolutionarily stable. Then $\lim_{p \to 0} v_{s_1}(s_1) = R$.
 \end{theorem}
@@
 \begin{theorem}
-  Suppose $s_1$ is evolutionary stable as $p$ goes to 0 (i.e., that it is evolutionary stable if condition \ref{fitnesscond} is replaced by $\lim_{p \to 0} (F (s_2)-F(s_1)) > 0$).
-Then $v_{s_1}(s_1) = R$.
-In other words, $s_1$ is utilitarian.
+  \label{pavlovtheorem}
+  The Pavlov strategy, displayed in \cref{pavlovfigure}, is evolutionarily stable.
 \end{theorem}
 
-\begin{theorem}
-  The Pavlov strategy, displayed in Figure ??, is evolutionary stable as $p$ goes to 0.
-\end{theorem}
-
-\begin{remark}
-  TFT, displayed in Figure ??, is not evolutionary stable as $p$ goes to 0. It has the stationary distribution $(1/4,1/4,1/4,1/4)$ which is smaller than $R$.
-\end{remark}
-
 \begin{figure}
+  \label{tftfigure}
   \includegraphics[width=4cm]{tft.jpg}
@@
 \begin{figure}
+  \label{pavlovfigure}
   \includegraphics[width=4cm]{pavlov.jpg}
@@
 
+\begin{remark}
+  TFT, displayed in \cref{tftfigure}, is not evolutionarily stable. It has the stationary distribution $(1/4,1/4,1/4,1/4)$ which is smaller than $R$.
+\end{remark}
+
+    \begin{proof}[Proof of \cref{evolutionarystable1}]
+      Suppose $s_1$ is such that it is not true that \begin{equation*}
+        \lim_{p \to 0 } v_{s_1}(s_1) = R
+      \end{equation*}
+      Note first that $v_{s_1}(s_1)$ is a polynomial in $p$, so the limit exists \todo{Formally prove this!}.
+Also, by the symmetry of the game, we know that $v_{s_1}(s_1) \leq R$ is always true, so this must mean that $\lim_{p \to 0} v_{s_1}(s_1) < R$ \todo{Formally prove this!}.
+In particular, then, there exists some $\beta > 0$ such that $v_{s_1}(s_1) = R - \beta$.
+
+      We want to prove that $s_1$ is not evolutionarily stable, and to do that, we want to prove that for any sufficiently small $\epsilon$, there exists a strategy $s_2$ that can invade $s_1$.
+
+      We create the strategy $s_2$ as follows.
+First, copy the entire $s_1$ machine into $s_2$.
+Suppose the state corresponding to the start state of $s_1$ is $c_s$.
+Let the output at $c_s$ be $G(c_s)$.
+Let the node it goes to upon perceiving the opponent move $G(c_s)$ be $T(c_s, G(c_s))$.
+Then, create a new state $c_0$ that outputs $\lnot G(c_s)$ and has transition $T(c_0, G(c_s)) = T(c_s, G(c_s))$.
+Create another new state $c_1$.
+Let $T(c_0, \lnot G(c_s)) = c_1$.
+Let $T(c_2, \cdot) = c_2$.
+Let $G(c_2) = C$.
+This completely describes $s_2$.
+
+      Now, we claim that the payoffs are as follows.
+      \begin{equation}
+        v_{s_2}(s_1) = (1 - p) \gamma  + p S
+      \end{equation}
+      \begin{equation}
+        v_{s_2}(s_2) = (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma
+      \end{equation}
+      \begin{equation}
+        v_{s_1}(s_1) = \gamma
+      \end{equation}
+      \begin{equation}
+        v_{s_1}(s_2) = (1-p) \gamma + p T
+      \end{equation}
+
+      Now, we simply compute $F(s_2) - F(s_1)$, and want to show that it is greater than 0 for all $\alpha$ that are sufficiently small.
+
+      \begin{align*}
+        F(s_2) - F(s_1) &= \\
+        &= (1 - \alpha) \cdot v_{s_2}(s_1) + \alpha \cdot v_{s_2}(s_2) - (1 - \alpha) \cdot v_{s_1}(s_1) - \alpha \cdot v_{s_1}(s_2) \\
+        &= 
+      \end{align*}
+
+      We can easily see that this works for some small $p$.
+
+    \end{proof}
+
+    \begin{proof}[Proof of \cref{pavlovtheorem}]
+
+    \end{proof}
+
 \end{document}

### 2020-05-06 realnotes2/notes.tex
write first draft

--- 2020-04-09 realnotes/notes.tex
@@
 \title{Prisoner's Dilemma}
-\author{Arvid Lunnemark}
+\author{Arvid Lunnemark. \\ \\
+Supervised by: Michael Sipser}
 
@@
 
+\section{Introduction}
+
+Lots to add here.
+
+Should some of the definitions be in the introduction?
+
+\section{Setup}
+
 \begin{definition}
-  The \textit{prisoner's dilemma} is a symmetric two-player game with two actions, cooperate ($C$) and defect ($D$), where, if player 1 plays $a$ and player 2 plays $b$, player 1 gets payoff 
+  The \textit{prisoner's dilemma} is a symmetric two-player game with two actions, cooperate ($C$) and defect ($D$), where, if player 1 selects action $a$ and player 2 selects action $b$, player 1 gets reward
   \begin{equation*}
-    p(a,b) = \begin{cases}
+    r(a,b) = \begin{cases}
       R &\text{if $a = C, b = C$} \\
@@
   \end{equation*}
-  We have $T > R > P > S$, and typically, we have the concrete values $T = 5$, $R = 3$, $P = 1$ and $S = 0$.
+  We require $T > R > P > S$ and $2R > T + S$. 
 \end{definition}
 
+A common choice in simulations of the iterated prisoner's dilemma is $T = 5$, $R = 3$, $P = 1$ and $S = 0$.
+
+We want to study the \textit{iterated} prisoner's dilemma, for which we can define strategies that determine their next move based on the history of previous moves.
+As discussed previously, we want to restrict ourselves to strategies with finite memory.
+
 \begin{definition}
-  A \textit{strategy} is a Moore machine (finite automaton with outputs) over the input and output alphabet $\{C, D\}$, with probability $1-p$ of following the correct transition and probability $p$ of following the incorrect transition.
+  A \textit{strategy} $s$ is a Moore machine (finite automaton with outputs) over the input and output alphabet $\{C, D\}$. 
 \end{definition}
 
-Note: this models an error probability in \textit{perception}.
-One could also think of an error probability in \textit{outcome}, but it is easy to see that the two are equivalent up to a change of the values of $R, S, T, P$.
+Notation-wise, we will use $c$ to denote states in $s$, $G_s(c)$ to denote the output at state $c$, and $T_s(c, a)$ to denote the state that $c$ transitions to upon receiving input $a$.
+For simplicity, we will also define the $\lnot$ operator such that $\lnot C = D$ and $\lnot D = C$, and $c_{\text{start}}(s)$ to be the start state of $s$.
+
+We will consider strategies in the presence of noise.
+To model that, we will assume that a strategy has a probability $1-p$ of following the correct transition and a probability $p$ of following the incorrect transition, at every step.
+Note that this models noise in \textit{perception}.
+One could also imagine modeling noise in \textit{action taken}, but it is easy to see that the two are equivalent up to a change of the values of $R, S, T, P$.
 
 \begin{definition}
-  Suppose strategy $s_1$ plays strategy $s_2$.
-This defines a \textit{$s_1$-$s_2$ graph} which is a Markov chain where each node represents a pair of states $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$.
-The transition probabilities are defined in the obvious way.
+  Suppose that strategy $s_1$ plays against strategy $s_2$.
+This defines an \textit{$s_1$-$s_2$ Markov chain} where each state $x$ is the vector $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$.
+The transition probabilities are defined in the obvious way, using the error probability $p$.
 \end{definition}
 
+We use the notation $G_{s_1,s_2}(c_1,c_2)$ to refer to the vector $(G_{s_1}(c_1), G_{s_2}(c_2))$, and we use $S_{s_1,s_2}$ to refer to the set of all states in the Markov chain.
+
 \begin{definition}
-  Let $\pi$ be the stationary distribution achieved by starting in the start state of the \textit{$s_1$-$s_2$ graph}.
-The \textit{payoff} of strategy $s_1$ when played against strategy $s_2$ is \begin{equation*}
-    v_{s_1}(s_2) = \sum \pi_{c_1, c_2} \cdot p(c_1, c_2).
+  \label{strategypayoffs}
+  Let $X_t$ be the random variable designating which state the $s_1$-$s_2$ Markov chain is in at time $t$. Then, the \text{payoff} of strategy $s_1$ when played against strategy $s_2$ is 
+  \begin{equation*}
+    v_{s_1}(s_2) = E \left[ \lim_{T \to \infty} \frac{1}{T} \sum_{t=0}^T r(G_{s_1,s_2} (X_t)) \Bigm| X_0 = (c_{\text{start}}(s_1), c_{\text{start}}(s_2)) \right]
   \end{equation*}
@@
 
-Note: The graph might be periodic in which case we we will not get a stationary distribution. I need to think about this special case but my intuition is that it shouldn't matter.
+That is, when $s_1$ plays against $s_2$, we define its payoff to be the average payoff over all possible infinite sequences of moves.
+Note that the expectation is taken over the infinite sequence $(X_0, X_1, \ldots)$.
+The limit inside is thus simply a normal time-average limit of bounded real numbers, which clearly exists.
+
+We will now introduce the notion of a time average distribution which will lead us to a second way of defining the payoff $v_{s_1}(s_2)$.
+\todo{is this really necessary??
+should i pick one or the other??
+should i move one of them to the next section?
+which one is easier to understand?
+do they trivially say the same thing?}
+
+\begin{definition}
+  \label{timeaveragedistribution}
+  The \textit{time average distribution} of the $s_1$-$s_2$ Markov chain given the start state $(a, b)$, denoted $\pi^{(a, b)}$, is the distribution such that \begin{equation*}
+    \pi_{c_1, c_2}^{(a, b)} = E \left[ \text{fraction of time in state $(c_1,c_2)$} \mid \text{initial state is } (a, b)  \right]
+  \end{equation*}
+  where the fraction of time is taken over the infinite sequence $(X_0, X_1, \ldots)$.
+\end{definition}
+
+\begin{lemma}
+  \label{payofftimeaverage}
+  The payoff when $s_1$ plays against $s_2$ is
+  \begin{equation*}
+    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} \pi_{c_1,c_2}^{(c_{\text{start}}(s_1), c_{\text{start}}(s_2))} \cdot r(c_1,c_2).
+  \end{equation*}
+\end{lemma}
+
+\begin{proof}
+  The key idea is that a time average sum where each element is one of finitely many values can be written as a frequency-weighted finite sum instead.
+Let $I_{c_1,c_2,t}$ be the indicator variable that is 1 if $G_{s_1,s_2}(X_t) = (c_1,c_2)$ and 0 otherwise.
+Then, we can write \begin{equation*}
+    \lim_{T \to \infty} \frac{1}{T} \sum_{t = 0}^T r(G_{s_1,s_2}(X_t)) = 
+    \lim_{T \to \infty}\frac{1}{T} \sum_{t = 0}^T \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1,c_2) \cdot I_{c_1,c_2,t}
+  \end{equation*}
+  We may now exchange the order of summation and move the finite sum out of the limit, to get \begin{equation*}
+    \lim_{T \to \infty}\frac{1}{T} \sum_{t = 0}^T \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1,c_2) \cdot I_{c_1,c_2,t}
+    = \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1,c_2) \cdot \lim_{T \to \infty} \sum_{t=0}^T \frac{I_{c_1,c_2,t} }{T}
+  \end{equation*}
+  We can now use \cref{strategypayoffs} and linearity of expectation to find that
+  \begin{equation*}
+    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1, c_2) E \left[ \lim_{T \to \infty} \sum_{t=0}^T \frac{I_{c_1,c_2,t}}{T} \bigm| X_0 = (c_{\text{start}}(s_1), c_{\text{start}}(s_2))
+    \right]
+  \end{equation*}
+  Finally, we note that this is exactly the statement of \cref{payofftimeaverage}, which proves our lemma.
+\end{proof}
+
+Appendix A contains more details on time average distributions.
+In particular, if a unique stationary distribution exists, it is equal to the time-average distribution, which enables us to quickly find the time-average distribution in many cases.
+
+We're now ready to look at how strategies interact.
 
@@
 
+One can think of this as saying that we have infinitely many members of the population, and that they all interact with everyone else.
+This justifies the usage of expectation when definining $v_{s_1}(s_2)$.
+
+We can now use the fitness of a strategy to compare it with other strategies in the same population.
+If a strategy $s_1$ has a higher fitness than another strategy $s_2$, that means that the frequency of $s_1$ will increase on the expense of the frequency of $s_2$, in the next step of the evolutionary process.
+This is getting us close to how we want to define stable strategies; our next move is looking not only at a single evolutionary step, but the entire evolutionary process.
+
 \begin{definition}
-  A strategy $s_1$ is \textit{$\epsilon$-invadable} if there exists a strategy $s_2$ such that in all populations $P$ with $S = \{s_1,s_2\}$ and $f(s_2) \geq \epsilon$, we have \begin{equation}
+  A strategy $s_1$ is \textit{$\epsilon$-invadable} if there exists a strategy $s_2$ such that in all populations $P$ with $S = \{s_1,s_2\}$ and $f(s_2) \geq \epsilon$, we have 
+  \begin{equation*}
     \label{fitnesscond}
     F(s_2) > F(s_1)
-  \end{equation}
+  \end{equation*}
 \end{definition}
 
+That is, if $s_1$ is $\epsilon$-invadable, there exists a strategy $s_2$ that can start as only a tiny fraction $\epsilon$ of the total population, and consistently have higher fitness than $s_1$, eventually causing overtaking $s_1$ completely.
+We are now finally ready to state our main definition.
+
 \begin{definition}
-  A strategy $s_1$ is \textit{evolutionary stable} if there exists an $\alpha \in (0,1)$ such that for all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
+  A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\alpha$, both in $(0,1)$, such that for all $p < p_0$, and all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
 \end{definition}
 
-Note: it is easy to see that this is just equivalent to saying that there exists some $\epsilon$ for which $s_1$ is not $\epsilon$-invadable.
+That is, a strategy $s_1$ is evolutionarily stable if it can withstand invasion attempts from any strategy that starts off in low numbers, as the probability of noise tends to 0.
+
+\section{Results}
+
+We can now state our results! Together, the following two theorems prove that in the setup described here, mutual cooperation arises as the only stable choice.
 
 \begin{theorem}
-  Suppose $s_1$ is evolutionary stable. Then $v_{s_1}(s_1) \geq \frac{S + T}{2}$.
+  \label{evolutionarystable1}
+  Suppose that a strategy $s_1$ is evolutionarily stable. Then $\lim_{p \to 0} v_{s_1}(s_1) = R$.
 \end{theorem}
@@
 \begin{theorem}
-  Suppose $s_1$ is evolutionary stable as $p$ goes to 0 (i.e., that it is evolutionary stable if condition \ref{fitnesscond} is replaced by $\lim_{p \to 0} (F (s_2)-F(s_1)) > 0$).
-Then $v_{s_1}(s_1) = R$.
-In other words, $s_1$ is utilitarian.
+  \label{pavlovtheorem}
+  The Pavlov strategy is evolutionarily stable.
 \end{theorem}
 
-\begin{theorem}
-  The Pavlov strategy, displayed in Figure ??, is evolutionary stable as $p$ goes to 0.
-\end{theorem}
-
-\begin{remark}
-  TFT, displayed in Figure ??, is not evolutionary stable as $p$ goes to 0. It has the stationary distribution $(1/4,1/4,1/4,1/4)$ which is smaller than $R$.
-\end{remark}
-
 \begin{figure}
+  \label{tftfigure}
   \includegraphics[width=4cm]{tft.jpg}
@@
 
+\begin{remark}
+  Tit-for-tat, displayed in \cref{tftfigure}, is not evolutionarily stable.
+It has the stationary distribution $(1/4,1/4,1/4,1/4)$ in its own Markov chain, which has a payoff that is significantly smaller than $R$.
+This should be intuitive: if tit-for-tat makes one mistake, it goes into a defection cycle that it doesn't break out of until it makes a second mistake.
+\end{remark}
+
+\section{Proofs}
+
+\subsection{Helpful Lemmas}
+
+\begin{lemma}
+  \label{limitvexists}
+  The limit \begin{equation*}
+    \lim_{p \to 0} v_{s_1}(s_2)
+  \end{equation*}
+  exists, for any strategies $s_1$ and $s_2$.
+\end{lemma}
+\begin{proof}
+  Hmmmm this was harder than I expected.
+  \todo{prove this; seems obvious}
+\end{proof}
+
+\begin{lemma}
+  \label{neverbetterthanr}
+  For any strategy $s$,
+  \begin{equation*}
+    v_{s}(s) \leq R
+  \end{equation*}
+\end{lemma}
+\begin{proof}
+  For notational simplicity, we will let $s_1$ and $s_2$ be two copies of strategy $s$. Then, $v_{s}(s) = v_{s_1}(s_2) = v_{s_2}(s_1)$. By definition, we have
+  \begin{equation*}
+    v_{s_1}(s_2) = \sum \pi_{c_1,c_2} \cdot r(c_1,c_2)
+  \end{equation*}
+  and
+  \begin{equation*}
+    v_{s_2}(s_1) = \sum \pi_{c_2,c_1} \cdot r(c_2,c_1).
+  \end{equation*}
+  Note that $\pi_{c_1,c_2}$ and $\pi_{c_2,c_1}$ refer to the same state, so we thus have \begin{equation*}
+    v_{s_1}(s_2) + v_{s_2}(s_1) = \sum \pi_{c_1,c_2} \cdot (r(c_1,c_2) + r(c_2,c_1))
+  \end{equation*}
+  which implies that \begin{equation*}
+    v_s(s) = \sum \left( \pi_{c_1,c_2} \cdot \frac{r(c_1,c_2) + r(c_2,c_1)}{2} \right).
+  \end{equation*}
+  Now, note that $r(c_1,c_2) + r(c_2,c_1) \in \{R + R, S + T, T+S, P + P\}$. Since $P < R$ and $T + S < 2R$, we thus find that \begin{equation*}
+    v_s(s) \leq \sum \pi_{c_1,c_2} \cdot R = R \sum \pi_{c_1,c_2} = R,
+  \end{equation*}
+  as desired.
+\end{proof}
+
+\subsection{Evolutionary Stability Implies Utilitarianism}
+
+With these lemmas, we are now ready to prove our first theorem.
+
+    \begin{proof}[Proof of \cref{evolutionarystable1}]
+      Suppose that the strategy $s_1$ is such that it is \textit{not} true that \begin{equation*}
+        \lim_{p \to 0 } v_{s_1}(s_1) = R
+      \end{equation*}
+      By \cref{limitvexists} and \cref{neverbetterthanr}, this assumption implies that the limit is strictly less than $R$. Define $\gamma = v_{s_1}(s_1)$. We thus know that \begin{equation*}
+        \gamma < R.
+      \end{equation*}
+
+      We want to prove that $s_1$ is not evolutionarily stable.
+
+      To do that, we want to prove that for all $p_0, \alpha \in (0,1)$, there exists $p < p_0$ and $\epsilon < \alpha$, such that $s_1$ is $\epsilon$-invadable.
+We choose $\epsilon = \alpha / 2$, and present a strategy $s_2$ that can invade $s_1$ for sufficiently small $p$.
+
+      We create the strategy $s_2$ as follows.
+First, copy the entire $s_1$ machine into $s_2$.
+Suppose that the state corresponding to the start state of $s_1$ is $c_s$.
+Recall that the output at $c_s$ is $G(c_s)$, and that the state $s$ goes to upon perceiving the opponent move $G(c_s)$ is $T(c_s, G(c_s))$.
+Now, create two new states: $c_0$ and $c_1$.
+Define the transitions as \begin{align*}
+        T(c_0, G(c_s)) &= T(c_s, G(c_s))\\
+        T(c_0, \lnot G(c_s)) &= c_1 \\
+        T(c_1, \cdot) &= c_1
+      \end{align*}
+      and the outputs as \begin{align*}
+        G(c_0) &= \lnot G(c_2)\\
+        G(c_1) &= C.
+      \end{align*}
+      Let the start state of $s_2$ be $c_0$.
+
+      Given this construction, we claim that the rewards are as follows:
+      \begin{align*}
+        v_{s_1}(s_1) &= \gamma \\
+        v_{s_1}(s_2) &\leq (1-p) \gamma + p T \\
+        v_{s_2}(s_1) &\geq (1 - p) \gamma  + p S \\
+        v_{s_2}(s_2) &\geq (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma 
+      \end{align*}
+      \todo{prove this! using time-average distributions this becomes relatively trivial}
+
+      Now, we simply compute $F(s_2) - F(s_1)$, which we want to show is greater than 0.
+
+      \begin{align*}
+        F(s_2) - F(s_1) &= \\
+        &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
+        &= (1 - \epsilon) (\gamma + p(\ldots)) + \epsilon (R + p(\ldots)) - (1-\epsilon) \gamma - \epsilon (\gamma + p(\ldots)) \\
+        &= \epsilon (R - \gamma) + p(\ldots)
+      \end{align*}
+
+      $R - \gamma > 0$ by our assumption.
+Clearly, since $(\ldots)$ is some polynomial in $p$, given an $\epsilon$ we can find a sufficiently small $p$ such that the full expression is positive.
+This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is not $\epsilon$-invadable for this value of $p$.
+In conclusion, then $s_1$ is not evolutionarily stable, which concludes the proof of \cref{evolutionarystable1}.
+
+    \end{proof}
+
+    \subsection{Evolutionarily Stable Strategies Exist}
+
+    \begin{proof}[Proof of \cref{pavlovtheorem}]
+
+      TBC.
+
+    \end{proof}
+
+    \section{Discussion of Model}
+
+    \subsection{Potential Other Models}
+
+    Right now we have only modeled noise in perception.
+One could think of another possible kind of noise: a ``failure of the mind,'' which perhaps could be modeled instead by a probability $p$ of being transported to any random state, instead.
+This would create ergodicity which is nice.
+
+    \section{Appendix: Non-stationary limiting distributions}
+
+    We might have periodicity, but for our purposes, we might as well extend the definition and look at periodic distributions as stationary too. The following two lemmas help with that.
+
+    \begin{lemma}
+      Given a starting distribution $v$ and a Markov matrix $M$, for every $\epsilon > 0$, there will exist a $k$ such that $|vP^{nk} - vP^{mk}| < \epsilon$ for all $n$ and $m$ $> 0$.
+    \end{lemma}
+
+    This proves that a Markov chain will always reach a periodic state.
+
+    \begin{lemma}
+      Suppose distributions form a chain $p_1 \to p_2 \to \cdots \to p_n \to p_1$. Then $\pi = \frac{p_1 + \ldots p_n}{n}$ is stationary.
+    \end{lemma}
+
+    This proves that we're able to talk about stationary distributions even when they don't really actually exist.
+
 \end{document}

### 2020-05-06 realnotes2/notes.tex
almost finish proof of theorem1

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{claim}{Claim}
 \newtheorem{problem}{Problem}
@@
 
-Lots to add here.
+Introduce the prisoner's dilemma, the iterated prisoner's dilemma and the goal of finding out how cooperation can appear. Introduce Axelrod along with some other previous results.
 
@@
   \label{strategypayoffs}
-  Let $X_t$ be the random variable designating which state the $s_1$-$s_2$ Markov chain is in at time $t$. Then, the \text{payoff} of strategy $s_1$ when played against strategy $s_2$ is 
+  Let $X_t$ be the random variable designating which state the $s_1$-$s_2$ Markov chain is in at time $t$. The \text{payoff} of strategy $s_1$ when played against strategy $s_2$ is 
   \begin{equation*}
@@
 
-\begin{proof}
+We may also make $r(c_1,c_2)$ into a vector, denoted by $r$, and write this as \begin{equation*}
+  v_{s_1}(s_2) = \pi^{(c_{\text{start}}(s_1), c_{\text{start}}(s_2))} \cdot r.
+\end{equation*}
+
+\begin{proof}[Proof of \cref{payofftimeaverage}]
   The key idea is that a time average sum where each element is one of finitely many values can be written as a frequency-weighted finite sum instead.
@@
 \end{figure}
+  \todo{use tikz for these figures!}
 
@@
   Tit-for-tat, displayed in \cref{tftfigure}, is not evolutionarily stable.
-It has the stationary distribution $(1/4,1/4,1/4,1/4)$ in its own Markov chain, which has a payoff that is significantly smaller than $R$.
+It has the stationary distribution $(\frac{1}{4},\frac{1}{4},\frac{1}{4},\frac{1}{4})$ in its own Markov chain, which has a payoff that is significantly smaller than $R$.
 This should be intuitive: if tit-for-tat makes one mistake, it goes into a defection cycle that it doesn't break out of until it makes a second mistake.
+
 \end{remark}
@@
 \begin{proof}
-  Hmmmm this was harder than I expected.
+  Use the fact that this is the dot product of a stationary distribution with a fixed coefficient vector, and then say something about how any stationary distribution with the correct sum needs to be a polynomial in $p$ maybe?
   \todo{prove this; seems obvious}
@@
       \end{align*}
-      Let the start state of $s_2$ be $c_0$.
+      Let the start state of $s_2$ be $c_0$. \todo{add a figure!! the construction is simple but this description is deceivingly hard}
 
-      Given this construction, we claim that the rewards are as follows:
+      \begin{claim}
+        \label{claimpayoffs}
+        Given the above construction of $s_2$, the following inequalities hold:
       \begin{align*}
-        v_{s_1}(s_1) &= \gamma \\
+        v_{s_1}(s_1) &\leq (1-p)^2 \gamma + 2(1-p)p R + p^2  R \\ 
         v_{s_1}(s_2) &\leq (1-p) \gamma + p T \\
         v_{s_2}(s_1) &\geq (1 - p) \gamma  + p S \\
-        v_{s_2}(s_2) &\geq (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma 
+        v_{s_2}(s_2) &\geq (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma .
       \end{align*}
-      \todo{prove this! using time-average distributions this becomes relatively trivial}
+      \end{claim}
+
+      We will first finish the proof of the theorem assuming this claim, and then prove it.
 
       Now, we simply compute $F(s_2) - F(s_1)$, which we want to show is greater than 0.
-
       \begin{align*}
@@
         &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
-        &= (1 - \epsilon) (\gamma + p(\ldots)) + \epsilon (R + p(\ldots)) - (1-\epsilon) \gamma - \epsilon (\gamma + p(\ldots)) \\
+        &= (1 - \epsilon) (\gamma + p(\ldots)) + \epsilon (R + p(\ldots)) - (1-\epsilon) (\gamma + p(\ldots)) - \epsilon (\gamma + p(\ldots)) \\
         &= \epsilon (R - \gamma) + p(\ldots)
@@
 
-      $R - \gamma > 0$ by our assumption.
+      We know that $R - \gamma > 0$ by our initial assumption.
 Clearly, since $(\ldots)$ is some polynomial in $p$, given an $\epsilon$ we can find a sufficiently small $p$ such that the full expression is positive.
@@
 
+    \end{proof}
+
+    \begin{proof}[Proof of \cref{claimpayoffs}]
+      We can prove this using either of the two definitions.
     \end{proof}

### 2020-05-06 paper/paper.tex
rename to paper again
add title
change title

--- 2020-05-06 realnotes2/notes.tex
@@
 
-\title{Prisoner's Dilemma}
+\title{Cooperation Provably Prevails in a Version of the Noisy Iterated Prisoner's Dilemma}
 \author{Arvid Lunnemark. \\ \\
@@
   \begin{equation*}
-    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} \pi_{c_1,c_2}^{(c_{\text{start}}(s_1), c_{\text{start}}(s_2))} \cdot r(c_1,c_2).
+    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} \pi_{c_1,c_2}^{(c_{\text{start}}(s_1), c_{\text{start}}(s_2))} \cdot r(G_{s_1}(c_1),G_{s_2}(c_2)).
   \end{equation*}
@@
 
-We may also make $r(c_1,c_2)$ into a vector, denoted by $r$, and write this as \begin{equation*}
+We may also make $r(G_{s_1}(c_1),G_{s_2}(c_2))$ into a vector, denoted by $r$, and write this as \begin{equation*}
   v_{s_1}(s_2) = \pi^{(c_{\text{start}}(s_1), c_{\text{start}}(s_2))} \cdot r.
@@
 
-      We will first finish the proof of the theorem assuming this claim, and then prove it.
+      Before proving this claim, we will use it to finish our proof of \cref{evolutionarystable1}.
 

### 2020-05-16 paper/paper.tex
change terminology
add tikz figures

@@
 \usepackage{cleveref}
+\usepackage{tikz}
+\usetikzlibrary{backgrounds,arrows,automata,positioning}
 
@@
 \begin{definition}
-  The \textit{prisoner's dilemma} is a symmetric two-player game with two actions, cooperate ($C$) and defect ($D$), where, if player 1 selects action $a$ and player 2 selects action $b$, player 1 gets reward
+  The \textit{prisoner's dilemma} is a symmetric two-player game with two actions, cooperate ($C$) and defect ($D$), where, if player 1 selects action $a$ and player 2 selects action $b$, player 1 gets the reward
   \begin{equation*}
@@
 Notation-wise, we will use $c$ to denote states in $s$, $G_s(c)$ to denote the output at state $c$, and $T_s(c, a)$ to denote the state that $c$ transitions to upon receiving input $a$.
-For simplicity, we will also define the $\lnot$ operator such that $\lnot C = D$ and $\lnot D = C$, and $c_{\text{start}}(s)$ to be the start state of $s$.
+For simplicity, we will also define the $\bar{\;}$ operator such that $\bar{C} = D$ and $\bar{D} = C$, and $c_{\text{start}}(s)$ to be the start state of $s$.
 
@@
 Note that the expectation is taken over the infinite sequence $(X_0, X_1, \ldots)$.
-The limit inside is thus simply a normal time-average limit of bounded real numbers, which clearly exists.
+The limit inside is thus simply a normal time average limit of bounded real numbers, which clearly exists.
 
@@
   \end{equation*}
-  where the fraction of time is taken over the infinite sequence $(X_0, X_1, \ldots)$.
+  where the expected fraction of time is taken over the infinite sequence $(X_0, X_1, \ldots)$.
 \end{definition}
+
+We will use $\pi$ to refer to $\pi^{(c_{\text{start}}(s_1), c_\text{start}(s_2))}$.
 
@@
   \begin{equation*}
-    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} \pi_{c_1,c_2}^{(c_{\text{start}}(s_1), c_{\text{start}}(s_2))} \cdot r(G_{s_1}(c_1),G_{s_2}(c_2)).
+    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} \pi_{c_1,c_2} \cdot r(G_{s_1}(c_1),G_{s_2}(c_2)).
   \end{equation*}
@@
 
-We may also make $r(G_{s_1}(c_1),G_{s_2}(c_2))$ into a vector, denoted by $r$, and write this as \begin{equation*}
-  v_{s_1}(s_2) = \pi^{(c_{\text{start}}(s_1), c_{\text{start}}(s_2))} \cdot r.
+We may also make $r(G_{s_1}(c_1),G_{s_2}(c_2))$ into a vector, denoted by $r$, and write this as the dot product \begin{equation*}
+  v_{s_1}(s_2) = \pi \cdot r.
 \end{equation*}
@@
 \begin{definition}
-  A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\alpha$, both in $(0,1)$, such that for all $p < p_0$, and all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
+  A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\epsilon_0$, both in $(0,1)$, such that for all $p < p_0$, and all $\epsilon < \epsilon_0$, $s_1$ is not $\epsilon$-invadable.
 \end{definition}
@@
 \begin{figure}
-  \label{tftfigure}
-  \includegraphics[width=4cm]{tft.jpg}
-  \centering
-  \caption{TFT.}
-\end{figure}
-\begin{figure}
   \label{pavlovfigure}
-  \includegraphics[width=4cm]{pavlov.jpg}
+\begin{tikzpicture}[node distance=2cm]
+  \node[state,initial] (s_0) {$C$};
+  \node[state] (s_1) [right of=s_0] {$D$};
+
+  \path[->] (s_0) edge [loop above] node {$C$} ();
+  \path[->] (s_0) edge [bend left] node [above] {$D$} (s_1);
+  \path[->] (s_1) edge [loop above] node {$C$} ();
+  \path[->] (s_1) edge [bend left] node [below] {$D$} (s_0);
+\end{tikzpicture}
   \centering
@@
 \end{figure}
-  \todo{use tikz for these figures!}
+
+\begin{figure}
+  \label{tftfigure}
+\begin{tikzpicture}[node distance=2cm]
+  \node[state,initial] (s_0) {$C$};
+  \node[state] (s_1) [right of=s_0] {$D$};
+
+  \path[->] (s_0) edge [loop above] node {$C$} ();
+  \path[->] (s_0) edge [bend left] node [above] {$D$} (s_1);
+  \path[->] (s_1) edge [loop above] node {$D$} ();
+  \path[->] (s_1) edge [bend left] node [below] {$C$} (s_0);
+\end{tikzpicture}
+  \centering
+  \caption{Tit for tat.}
+\end{figure}
+  % \begin{tikzpicture}[scale=0.8, show background rectangle]
+  %   \draw (0,0) -- (2,1);
+  %   \draw (1,1) circle (3);
+  %   \path[draw] (0,0) -- (1,1) -- (2,0);
+  % \end{tikzpicture}
 
@@
 
-      To do that, we want to prove that for all $p_0, \alpha \in (0,1)$, there exists $p < p_0$ and $\epsilon < \alpha$, such that $s_1$ is $\epsilon$-invadable.
-We choose $\epsilon = \alpha / 2$, and present a strategy $s_2$ that can invade $s_1$ for sufficiently small $p$.
+      To do that, we want to prove that for all $p_0, \epsilon_0 \in (0,1)$, there exists $p < p_0$ and $\epsilon < \epsilon_0$, such that $s_1$ is $\epsilon$-invadable.
+We choose $\epsilon = \epsilon_0 / 2$, and present a strategy $s_2$ that can invade $s_1$ for sufficiently small $p$.
 
@@
 First, copy the entire $s_1$ machine into $s_2$.
-Suppose that the state corresponding to the start state of $s_1$ is $c_s$.
-Recall that the output at $c_s$ is $G(c_s)$, and that the state $s$ goes to upon perceiving the opponent move $G(c_s)$ is $T(c_s, G(c_s))$.
+Suppose that the state corresponding to the start state of $s_1$ is $c$, and that the output at $c$ is $\alpha$, and that the state $s$ goes to upon perceiving the opponent move $\alpha$ is $c' = T(c, \alpha)$.
 Now, create two new states: $c_0$ and $c_1$.
 Define the transitions as \begin{align*}
-        T(c_0, G(c_s)) &= T(c_s, G(c_s))\\
-        T(c_0, \lnot G(c_s)) &= c_1 \\
+        T(c_0, \alpha) &= c'\\
+        T(c_0, \bar{\alpha}) &= c_1 \\
         T(c_1, \cdot) &= c_1
@@
       \end{align*}
-      Let the start state of $s_2$ be $c_0$. \todo{add a figure!! the construction is simple but this description is deceivingly hard}
+      Let the start state of $s_2$ be $c_0$.
+
+      \begin{figure}
+        \centering
+        \begin{tikzpicture}[node distance=2cm]
+          \node[state, initial] (c_0) {$\bar{\alpha}$};
+          \node[state] (c_1) [below right of=c_0] {$C$};
+          \node[state] (c_prime) [above right of=c_0] {};
+          \node[state] (c) [above left of=c_prime] {$\alpha$};
+
+          \path[->] (c_0) edge node [above right] {$\bar{\alpha}$} (c_1);
+          \path[->] (c_1) edge [loop right] node [right] {$C,D$} ();
+          \path[->] (c_0) edge node [above left] {$\alpha$} (c_prime);
+          \path[->] (c) edge node [below left] {$\alpha$} (c_prime);
+        \end{tikzpicture}
+        \caption{Constructon of invasion strategy, used in the proof of \cref{evolutionarystable1}}
+        \label{invasionstrategy}
+      \end{figure}
 
@@
 
-    \section{Appendix: Non-stationary limiting distributions}
+    \section{Appendix: Time average distributions}
 

### 2020-05-20 paper/paper.tex
fix things

@@
 \theoremstyle{definition}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{proposition}{Proposition}
-\newtheorem{corollary}{Corollary}
-\newtheorem{lemma}{Lemma}
-\newtheorem{claim}{Claim}
-\newtheorem{problem}{Problem}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{conjecture}[theorem]{Conjecture}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{claim}[theorem]{Claim}
 \theoremstyle{remark}
@@
 \newtheorem*{example}{Example}
-\newtheorem*{solution}{Solution}
 
@@
 
+\todo{clean up the julia notebook and refer to it for people who want to play with the setup}
+
 \section{Setup}
@@
   \end{equation*}
-  We require $T > R > P > S$ and $2R > T + S$. 
+  We require $T > R > P > S$.
 \end{definition}
@@
 
-\begin{theorem}
+\begin{conjecture}
   \label{pavlovtheorem}
-  The Pavlov strategy is evolutionarily stable.
-\end{theorem}
+  Suppose $2R > T + P$. Then, the Pavlov strategy is evolutionarily stable.
+\end{conjecture}
 
@@
 \begin{proof}
-  Use the fact that this is the dot product of a stationary distribution with a fixed coefficient vector, and then say something about how any stationary distribution with the correct sum needs to be a polynomial in $p$ maybe?
-  \todo{prove this; seems obvious}
+  Note that $v_{s_1}(s_2)$ is a weighted average of the rewards $R, T, S$ and $P$. Therefore, $S \leq v_{s_1}(s_2) \leq T$. \todo{eughhhh fix this somehow}
 \end{proof}
@@
 
-    \begin{proof}[Proof of \cref{pavlovtheorem}]
+    Unfortunately, we have no proof of \cref{pavlovtheorem}.
 
-      TBC.
-
-    \end{proof}
+    We note that the $2R > T + P$ condition is necessary.
+Otherwise, the AllD strategy would be able to invade Pavlov.
+We see this by noting that $\lim_{p \to 0} v_{s_2}(s_1) = T + P$ if $s_2$ is AllD and $s_1$ is Pavlov, and that AllD is better against itself than Pavlov is against it.
 
@@
 
-    \subsection{Potential Other Models}
+    \subsection{Other Potential Models}
 
@@
 
+    \section{Appendix: Probablistic Automata}
+
+    In this paper, we have considered strategies that make a deterministic move based on what they perceive.
+One could also imagine strategies that attaches a certain probability distribution to a perceived input, and chooses their next action based on that.
+In this appendix we show that these can be reduced to the deterministic ones, and thus that all results for the deterministic ones also hold for the probabilistic ones.
+
+    Proof idea: we can use cycles in the Markov chain with n total outputs, x of which are to state 1, to model getting to state 1 with probability x / n.
+this assumes that the outputs are of low enough probability, which can be achieved by chaining together lots of (1-p) transitions, which go to 0.
+
+    the hard part of this is showing that modifying the finite automaton like this won't hurt us.
+in fact, it would certainly not hurt us if not every state had to give an output.
+but that doesn't work for our model i think.
+so there are certainly things to think about here.
+
 \end{document}

### 2020-05-20 paper/paper.tex
some progress

@@
 
-\title{Cooperation Provably Prevails in a Version of the Noisy Iterated Prisoner's Dilemma}
+\newenvironment{subproof}[1][\proofname]{%
+  \renewcommand{\qedsymbol}{$\blacksquare$}%
+  \begin{proof}[#1]%
+}{%
+  \end{proof}%
+}
+
+\title{Cooperation Is Provably Required in a Version of the Noisy Iterated Prisoner's Dilemma}
 \author{Arvid Lunnemark. \\ \\
@@
 
+In this section we thoroughly define our problem setup.
+
 \begin{definition}
@@
   \end{equation*}
-  We require $T > R > P > S$.
+  We require $T > R > P > S$ and $2R > T + S$.
 \end{definition}
@@
 Notation-wise, we will use $c$ to denote states in $s$, $G_s(c)$ to denote the output at state $c$, and $T_s(c, a)$ to denote the state that $c$ transitions to upon receiving input $a$.
-For simplicity, we will also define the $\bar{\;}$ operator such that $\bar{C} = D$ and $\bar{D} = C$, and $c_{\text{start}}(s)$ to be the start state of $s$.
+For simplicity, we will also define the $\bar{\;}$ operator such that $\bar{C} = D$ and $\bar{D} = C$.
 
@@
 \end{equation*}
+
+\todo{I think the best way to handle this is to remove the expected value definition altogether.
+just do the time average distribution and skip the proof of the lemma.
+it is honestly a very natural definition}
 
@@
 
-\subsection{Helpful Lemmas}
+\subsection{The time average distribution}
+
+Before we prove \cref{evolutionarystable1}, we need to understand what the payoff $v_{s_1}(s_2)$ really means.
+In this subsection, we prove a series of lemmas that characterize the time average distribution, and consequently $v_{s_1}(s_2)$.
+
+\begin{lemma}
+  \label{timeaverageisstationary}
+  The time average distribution $\pi^{(a,b)}$, for any starting state $(a,b)$, is a stationary distribution of the Markov chain.
+\end{lemma}
+
+We state this lemma without proof, as it is a fairly standard result.
+Important to note is that the time average distribution is not necessarily a \textit{unique} stationary distribution, as we make no assumptions that our Markov chain be ergodic.
+
+\begin{definition}
+  A \textit{strongly connected component} of a directed graph is a subgraph where there is a path from every node to every other node.
+\end{definition}
+
+\begin{definition}
+  An \textit{absorbing component} of a directed graph is a subgraph where there are no edges from vertices inside the component to vertices outside it.
+\end{definition}
+
+We may also put both of the terms together and talk about absorbing strongly connected components, which, as shown by the next few lemmas, are useful.
+
+\begin{lemma}
+  An absorbing strongly connected component has a unique time average distribution, i.e., the time average distribution does not depend on the start state.
+\end{lemma}
+
+\begin{lemma}
+  \label{timeaveragedistributiondecomposition}
+  Let $\mathcal{S}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain.
+For every $C \in \mathcal{S}$, let $\pi^{(C)}$ be its unique time average distribution.
+Then, for some probabilities $p_C$ with $\sum_{C \in \mathcal{S}} p_C = 1$, depending only on the condensation of the Markov chain, we have \begin{equation*}
+    \pi = \sum_{C \in \mathcal{S}} p_C \cdot \pi^{(C)}.
+  \end{equation*}
+\end{lemma}
+
+Finally, we can characterize the absorbing strongly connected components of the $s_1$-$s_2$ Markov chain in terms of the absorbing SCCs of $s_1$ and $s_2$ separately.
+
+\begin{lemma}
+  \label{absorbingstronglyconnectedcomponentscartesianproduct}
+  Let $\mathcal{C}_1$ be the set of absorbing SCCs of $s_1$, and similarly, let $\mathcal{C}_2$ be the set of absorbing SCCs of $s_2$.
+If $\mathcal{C}$ is the set of absorbing SCCs of the $s_1$-$s_2$ Markov chain, then it is the cartesian product of the two, i.e.,
+  \begin{equation*}
+    \mathcal{C} = \mathcal{C}_1 \times \mathcal{C}_2.
+  \end{equation*}
+\end{lemma}
 
@@
 \begin{proof}
-  Note that $v_{s_1}(s_2)$ is a weighted average of the rewards $R, T, S$ and $P$. Therefore, $S \leq v_{s_1}(s_2) \leq T$. \todo{eughhhh fix this somehow}
+  By definition, \begin{equation*}
+    v_{s_1}(s_2) = \pi \cdot r.
+  \end{equation*}
+
+  By \cref{timeaverageisstationary}, $\pi$ is a stationary distribution.
+In particular, if $M$ is the transition matrix for the $s_1$-$s_2$ Markov chain, then $\pi$ is an eigenvector of $M$ with eigenvalue 1.
+This implies that $\pi$ is in the nullspace of $M - I$.
+Thus, we can find $\pi$ by solving for $X$ in $(M - I)X = 0$.
+If we solve this using Gaussian elimination and back substitution, it is clear that the entries of $\pi$ will be on the form $\frac{f(p)}{g(p)}$ where $f$ and $g$ are polynomials in $p$, since each
+  entry of $M$ is a second-degree polynomial in $p$.
+This is continuous for all $p$ where $g(p) \neq 0$, and since $g$ will have a finite degree it is therefore continuous in a small right neighborhood of 0.
+Finally, note that $v_{s_1}(s_2)$ is between $S$ and $T$, which in conclusion means that the limit of it as $p$ goes to 0 tends to a finite number, as desired.
 \end{proof}
@@
 
-With these lemmas, we are now ready to prove our first theorem.
-
     \begin{proof}[Proof of \cref{evolutionarystable1}]
@@
       \end{equation*}
-      By \cref{limitvexists} and \cref{neverbetterthanr}, this assumption implies that the limit is strictly less than $R$. Define $\gamma = v_{s_1}(s_1)$. We thus know that \begin{equation*}
+      By \cref{limitvexists} and \cref{neverbetterthanr}, this assumption implies that the limit is strictly less than $R$. Define $\gamma = v_{s_1}(s_1)$. Then,
+      \begin{equation*}
         \gamma < R.
@@
 
-      We want to prove that $s_1$ is not evolutionarily stable.
+      We want to prove that $s_1$ is not evolutionarily stable. This would prove the theorem.
 
@@
 
+      The underlying idea is to create $s_2$ such that $s_1$ can see no difference between itself and $s_2$, while $s_2$, on the other hand, can.
+In that case, we would have $v_{s_1}(s_2) = v_{s_2}(s_1) = v_{s_1}(s_1)$, and could construct $s_2$ such that it always cooperates when it recognizes itself, thereby yielding $v_{s_2}(s_2) = R$.
+This would give $s_2$ a higher fitness than $s_1$.
+The rest of this proof executes this plan in detail.
+
       We create the strategy $s_2$ as follows.
+First, we copy all of $s_1$ into $s_2$.
+Let $\mathcal{C}_1$ be the absorbing strongly connected components of $s_1$.
+The key idea, now, is to replace each original absorbing strongly connected component $C_1 \in \mathcal{C}_1$ with a new absorbing component $C_2$, which has the capability of self identifying.
+
+      We construct $C_2$ as follows. First, it finds out in finite time what action $A$ that $s_1$ will take at some large finite future time $T$ (with probability $1 - O(p)$). 
+      After that, it responds with the opposite action (which we will denote by $\bar{A}$) right after time $T$. 
+      If $s_2$ subsequently perceives that its opponent takes action $A$, it will transition into a copy of $C_1$, the absorbing strongly connected component that $C_2$ replaces.
+On the other hand, if $s_2$ perceives its opponent taking action $\bar{A}$, then it will transition into an absorbing strongly connected component that always cooperates.
+A schematic view of $C_2$ is presented in figure ????.
+This concludes the construction of $s_2$.
+
+      \todo{add picture here when low on focus}
+
+      We now want to prove that $s_2$ can in fact invade $s_1$ for small $p$.
+First, we will assume that $C_2$ can find out the action $A$ that $s_1$ would take at time $T$, and finish the proof of our theorem using this assumption.
+After that, we will prove that that is indeed the case.
+
+      \begin{claim}
+        \label{claimpayoffs}
+        Given the above construction of $s_2$, the payoffs are as follows.
+      \begin{align*}
+        v_{s_1}(s_1) &= \gamma \\
+        v_{s_1}(s_2) &= \gamma + O(p) \\
+        v_{s_2}(s_1) &= \gamma + O(p) \\
+        v_{s_2}(s_2) &= R + O(p)
+      \end{align*}
+      \end{claim}
+      \begin{subproof}
+        By definition, $v_{s_1}(s_1) = \gamma$. By \cref{timeaveragedistributiondecomposition}, the time average distribution of the $s_1$-$s_1$ Markov chain $\pi$ is \begin{equation*}
+          \pi = \sum_{(C_1, C_1') \in \mathcal{C}_1 \times \mathcal{C}_1} p_{(C_1, C_1')} \cdot \pi^{(C_1, C_1')}
+        \end{equation*}
+        When creating $s_2$, we replaced each $C_1$ by a $C_2$, containing two absorbing strongly connected components: a copy of $C_1$, which we call $C_{1, \text{copy}}$ and the always cooperating $C_c$.
+This means that the condensation of the $s_1$-$s_2$ Markov chain is the same as the condensation of the $s_1$-$s_1$ Markov chain, except that each absorbing strongly connected component has been replaced by two.
+That is, the absorbing strongly connected component $(C_1, C_1')$ of the $s_1$-$s_1$ chain has become both $(C_1, C_{1, \text{copy}})$ and $(C_1, C_c)$ in the $s_1$-$s_2$ chain.
+Then, we find by our construction, that
+        \begin{align}
+          \label{c1c1copy}
+          p_{(C_1, C_{1, \text{copy}})} &= (1 - O(p)) p_{(C_1, C_1')} \\
+          p_{(C_1, C_c)} &= O(p) p_{(C_1, C_1')}
+        \end{align}
+        This equation is true because when $s_2$ plays against $s_1$, it identifies the action $A$ that $s_1$ will take at time $A$ with probability $(1 - O(p))$, and upon seeing that transitions to $C_{1, \text{copy}}$ with probability $(1-p)$.
+
+        Now, we can calculate $\tau$, the time average distribution of the $s_1$-$s_2$ Markov chain:
+        \begin{align*}
+          \tau &= \sum_{(C_1, C') \in \mathcal{C}_1 \times \mathcal{C}_2} p_{(C_1, C')} \cdot \pi^{(C_1, C')} \\
+          &= \sum_{(C_1, C_{1, \text{copy}})} p_{(C_1, C_{1, \text{copy}})} \cdot \pi^{(C_1, C_{1, \text{copy}})} + \sum_{(C_1, C_c)} p_{(C_1, C_c)} \cdot \pi^{(C_1, C_c)} \\
+          &= \sum_{(C_1, C_1') \in \mathcal{C}_1 \times \mathcal{C}_1} p_{(C_1, C_1')} \cdot \pi^{(C_1, C_1')} + O(p) \\
+          &= \pi + O(p)
+        \end{align*}
+        where the second-to-last equality follows by \cref{c1c1copy}. \todo{we're kinda abusing notation here} This proves that $v_{s_1}(s_2) = \gamma + O(p)$ and $v_{s_2}(s_1) = \gamma + O(p)$.
+
+        For the last part of the claim, note that at time $T$, $s_2$ and its copy $s_2$ will be in the exact same state with probability at least $(1-p)^T = 1 - O(p)$, since if there are no mistakes they will both go to the exact same states.
+Thus, when $s_2$ outputs $\bar{A}$ to identify $s_1$, both of them will do that, and thus transition to an always cooperating state with probability $1 - O(p)$.
+Therefore, $v_{s_2}(s_2) = R - O(p)$, which finishes the proof of the claim.
+      \end{subproof}
+
+      Given \cref{claimpayoffs}, we simply compute $F(s_2) - F(s_1)$, which we want to show is greater than 0.
+      \begin{align*}
+        F(s_2) - F(s_1) &= \\
+        &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
+        &= (1 - \epsilon) (\gamma - O(p)) + \epsilon (R - O(p)) - (1-\epsilon) \gamma - \epsilon (\gamma + O(p)) \\
+        &\leq \epsilon (R - \gamma) + O(p)
+      \end{align*}
+
+      We know that $R - \gamma > 0$ by our initial assumption.
+For a sufficiently small $p$, thus, $F(s_2) - F(s_1) > 0$.
+This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is not $\epsilon$-invadable for this value of $p$.
+In conclusion, then, $s_1$ is not evolutionarily stable, which concludes the proof of \cref{evolutionarystable1}.
+
+    \end{proof}
+
+      We now return to the part we left out: how to construct $C_2$ such that it can find out which action that $s_1$ would take at time $T$.
+
+      \begin{claim}
+        There is a $T$ and a procedure implementable in $C_2$ that runs in a finite number of steps and can determine the action $A$ of $s_1$ at time $T$ with probability $1 - O(p)$.
+      \end{claim}
+
+      \begin{proof}
+          Let $\mathcal{C}_1$ be the set of all absorbing strongly connected components in $s_1$.
+           Let $M$ be the set of all tuples of $(C, c)$ where $C \in \mathcal{C}_1$ and $c$ is a possible state of $C$.
+Let $M'$ be a subset of $M$ where a tuple $(C, c)$ has been removed if there is another tuple $(C', c') \in M'$ produces exactly the same output on all inputs, when only following the $1-p$ transitions.
+
+        The procedure works as follows:
+        \begin{enumerate}
+          \item Simulate $C_1$ for $N$ steps.
+(This can be done by e.g.
+duplicating $C_1$ $N$ times and having all edges advance to corresponding state in the next copy.) Choose $N$ so that after $N$ timesteps, $s_1$ will be in an absorbing strongly connected component with probability $1 - O(p)$.
+\todo{move this out of the claim???
+but it's annoying and needs a space somewhere}
+          \item Iterate over every pair of tuples $(C,c)$ and $(C', c')$ in $M$:
+          \begin{enumerate}
+            \item There is a string $S$ of actions on which $C$ starting in $c$ and $C'$ starting in $c'$ will produce a different output at a time $t$.
+Now, output the string $S$, and determine if the sequence of perceived actions corresponds to $(C,c)$ or $(C', c')$.
+          \end{enumerate}
+        \end{enumerate}
+        Choose $T = \max t$ over all $t$.
+         With probability $1 - O(p)$, there will be exactly 1 pair $(C, c)$ that corresponds to all perceived actions in each of its comparisons.
+We know what $C$, when starting in $c$, outputs at time $T$.
+Thus, with probability $1 - O(p)$, we know what action $A$ that $s_1$ outputs at time $T$.
+
+        It should be noted that this procedure is easily implementable on a finite automaton, using e.g. a decision tree structure.
+      \end{proof}
+
+      \todo{KEY IMPORTANT DETAIL: we need to say that at a large finite $T$ then $s_1$ will be within an absorbing strongly connected component.
+probably need to look at all possible starting states of all possible ASCCs, not only every ASCC.
+uhhhhhhhhhhhh also need to modify our language a little bit to think about the fact that absorbing strongly connected components of $s_2$ actually DO AFFECT the condensation probabilities.
+however, since $C_2$ behaves as $C_1$ for many steps, it won't affect it meaningfully.
+}
+
+    \begin{proof}
+      \begin{claim}
+        Suppose that there are no mistakes, i.e., that $p = 0$. We can create a $C_2$ such that 
+      \end{claim}
+
       First, copy the entire $s_1$ machine into $s_2$.
@@
       \begin{claim}
-        \label{claimpayoffs}
         Given the above construction of $s_2$, the following inequalities hold:

### 2020-05-20 paper/paper.tex
some more changes

@@
       We construct $C_2$ as follows. First, it finds out in finite time what action $A$ that $s_1$ will take at some large finite future time $T$ (with probability $1 - O(p)$). 
+      \todo{THIS DOESN'T WORK.
+The problem is the circulation states: if $s_2$ goes into an absorbing SCC before $s_1$ does, our modified $C_2$ might affect the probabilities of $s_1$ going into certain absorbing SCCs.
+I think a way to fix that is as follows: You take $C_1$, and replace each of its individual states by an SCC, targeting a particular non-absorbing SCC of $s_1$.
+Make sure that they circulate in sync with each other.
+That means that we will be able to delay getting into the identification stage of $C_2$ which is exactly what we want.}
       After that, it responds with the opposite action (which we will denote by $\bar{A}$) right after time $T$. 

### 2020-05-21 paper/paper.tex
write introduction

@@
 \usepackage{tikz}
+\usepackage{blkarray}
 \usetikzlibrary{backgrounds,arrows,automata,positioning}
@@
 
-Introduce the prisoner's dilemma, the iterated prisoner's dilemma and the goal of finding out how cooperation can appear. Introduce Axelrod along with some other previous results.
+The prisoner's dilemma is a classic symmetric two-player game, where both players acting in their own best interests leads to an outcome that is not optimal for either of them.
+There are two actions, cooperate ($C$) and defect ($D$), with the payoffs represented by the matrix
+\begin{equation*}
+  \begin{matrix}
+      &  C & D \\ 
+    C &  (R, R) & (S, T) \\
+    D & (T, S) & (P, P)
+  \end{matrix}
+  % \begin{blockarray}{cccccc}
+  %   a & b & c & d & e \\
+  %   \begin{block}{c(ccccc)}
+  %     1 & 1 & 1 & 1 & 1 & f \\
+  %     0 & 1 & 0 & 0 & 1 & g \\
+  %     0 & 0 & 1 & 0 & 1 & h \\
+  %     0 & 0 & 0 & 1 & 1 & i \\
+  %     0 & 0 & 0 & 0 & 1 & j \\
+  %   \end{block}
+  %   \end{blockarray}
+\end{equation*}
+where $T > R > P > S$ (the row player gets the first in each pair of numbers).
+Given the other player's action, a player always maximizes their payoff by choosing $D$, but with both doing so, the players end up in the $(D, D)$ state, which is worse for both than the $(C, C)$ state since $R > P$.
+The idea of this game has been applied to many societal situations, including climate change, etc.
 
-Should some of the definitions be in the introduction?
+To make the game more interesting, one can consider playing it multiple times in a row with multiple players.
+One might loosely connect the repeated multiplayer game to evolution: why have humans evolved to cooperate with each other?
+In a seminal paper in 1980, Axelrod presented evidence that in a repeated setting with multiple players facing off in a tournament, cooperation can arise as the strategy of choice even for a selfish player \cite{axelrod1980effective}.
+In particular, the mostly cooperating Tit-for-tat strategy, depicted in \cref{tftfigure}, won his tournament.
+The idea is that even though Tit-for-tat loses when played against defecting strategies, that is outweighed by it being heavily rewarded when cooperating with other cooperating strategies.
 
-\todo{clean up the julia notebook and refer to it for people who want to play with the setup}
+\begin{figure}
+\begin{tikzpicture}[node distance=2cm]
+  \node[state,initial] (s_0) {$C$};
+  \node[state] (s_1) [right of=s_0] {$D$};
+
+  \path[->] (s_0) edge [loop above] node {$C$} ();
+  \path[->] (s_0) edge [bend left] node [above] {$D$} (s_1);
+  \path[->] (s_1) edge [loop above] node {$D$} ();
+  \path[->] (s_1) edge [bend left] node [below] {$C$} (s_0);
+\end{tikzpicture}
+  \centering
+  \caption{The Tit-for-tat strategy.
+At any one point, it is in one of two states, taking the action corresponding to the label of the state.
+Upon perceiving its opponent's move, it decides to switch state if the opponent does something different.}
+  \label{tftfigure}
+\end{figure}
+
+Clearly, however, the winning strategy in a tournament depends on the composition of strategies in that tournament.
+If all strategies had been of the type to always defect, Tit-for-tat would not have won the Axelrod tournament.
+Therefore, follow-up tournaments and simulations have been run, and perhaps surprisingly, most provide further evidence that cooperation is the prevailing strategy, as Axelrod summarizes in a 1981 book \cite{axelrod1981evolution}.
+There, he also introduces an evolutionary model that can be used for analyzing the game theoretically: strategies are considered to exist in a population that evolves over time in an evolutionary way (more successful strategies reproduce, and mutations can happen).
+In that context, a strategy is said to be evolutionarily stable if, supposing it controls a large share of the population, it resists being overtaken by any other strategy entering the population in small numbers.
+That is, populations of evolutionarily stable strategies form the equilibruim states of the evolutionary process, and thus, one may say that evolutionarily stable strategies are the ``best'' strategies.
+
+In light of Axelrod's tournaments highlighting the effectiveness of cooperation, it has long been a goal to prove that a cooperating strategy like Tit-for-tat is evolutionarily stable.
+A few variants of this has been shown: First, Nowak has shown that in the finitely repeated game, a strategy that always defects is sometimes not evolutionarily stable \cite{nowak2004emergence}.
+Second, Binmore has shown that in the infinitely repeated game where strategies are modeled as finite automatons where having more states comes at a cost, a strategy needs to cooperate with itself to be stable \cite{binmore1992evolutionary}.
+And third, Fundenberg and Maskin have shown that a strategy needs to cooperate with itself to be stable also in the infinitely repeated game in the presence of a certain notion of infinitesimally small noise \cite{fundenberg1990evolution}.
+
+Notably, all these results impose additional restrictions on the setup, and as noted by Fundenberg and Maskin, one has to do that: in the deterministic infinitely repeated game, one could create a strategy that can self-identify — if the opponent ever deviates from the pattern, the strategy resorts to defecting for the rest of the game.
+This strategy will be evolutionarily stable, but is not cooperative.
+
+In this paper, we, like Fundenberg and Maskin, choose the addition of noise as our restriction.
+In contrast to them, however, we present a model where, at every step, each strategy has a tiny probability $p$ of doing the wrong thing, which is arguably the most natural way of modeling noise.
+In \cref{sectionsetup}, we define the setup of our version of the problem in detail.
+Then, in \cref{sectionresults}, we state the two main results: that a strategy needs to be cooperative to be evolutionarily stable, and that such evolutionarily stable strategies exist.
+In \cref{sectionproofs}, we prove our results.
+Finally, in \cref{sectiondiscussion} we briefly discuss other potential ways of modeling the problem.
 
 \section{Setup}
+\label{sectionsetup}
 
-In this section we thoroughly define our problem setup.
+In this section we meticulously define our setup.
 
@@
 \section{Results}
+\label{sectionresults}
 
@@
 
-\begin{figure}
-  \label{tftfigure}
-\begin{tikzpicture}[node distance=2cm]
-  \node[state,initial] (s_0) {$C$};
-  \node[state] (s_1) [right of=s_0] {$D$};
-
-  \path[->] (s_0) edge [loop above] node {$C$} ();
-  \path[->] (s_0) edge [bend left] node [above] {$D$} (s_1);
-  \path[->] (s_1) edge [loop above] node {$D$} ();
-  \path[->] (s_1) edge [bend left] node [below] {$C$} (s_0);
-\end{tikzpicture}
-  \centering
-  \caption{Tit for tat.}
-\end{figure}
   % \begin{tikzpicture}[scale=0.8, show background rectangle]
@@
 \section{Proofs}
+\label{sectionproofs}
 
@@
     \section{Discussion of Model}
+    \label{sectiondiscussion}
 
@@
 
+    \bibliography{bibliography.bib}
+    \bibliographystyle{acm}
+
 \end{document}

### 2020-05-21 paper/paper.tex
finish setup section

@@
   %   \end{blockarray}
-\end{equation*}
+\end{equation*} \todo{copy the table from axelrod's book because it's much better and much clearer!!!}
 where $T > R > P > S$ (the row player gets the first in each pair of numbers).
 Given the other player's action, a player always maximizes their payoff by choosing $D$, but with both doing so, the players end up in the $(D, D)$ state, which is worse for both than the $(C, C)$ state since $R > P$.
-The idea of this game has been applied to many societal situations, including climate change, etc.
+The game can been used to model many real-world situations, including countries failing to act to stop climate change, doping in sport, and economic competition.
 
@@
 
-In this section we meticulously define our setup.
+In this section we define our setup.
+In summary, we consider the infinitely repeated prisoner's dilemma played by finite automata in infinite populations, evolving evolutionarily in the presence of noise.
+
+\subsection{Formal Definition}
+
+First, we define the reward function.
 
@@
 
-A common choice in simulations of the iterated prisoner's dilemma is $T = 5$, $R = 3$, $P = 1$ and $S = 0$.
-
-We want to study the \textit{iterated} prisoner's dilemma, for which we can define strategies that determine their next move based on the history of previous moves.
-As discussed previously, we want to restrict ourselves to strategies with finite memory.
+When we study the \textit{iterated} prisoner's dilemma, we want to look at strategies that determine their next move based on the history of previous moves.
+We restrict ourselves to strategies that can be implemented on a computer with finite memory.
 
@@
 
-Notation-wise, we will use $c$ to denote states in $s$, $G_s(c)$ to denote the output at state $c$, and $T_s(c, a)$ to denote the state that $c$ transitions to upon receiving input $a$.
-For simplicity, we will also define the $\bar{\;}$ operator such that $\bar{C} = D$ and $\bar{D} = C$.
-
 We will consider strategies in the presence of noise.
@@
 Note that this models noise in \textit{perception}.
-One could also imagine modeling noise in \textit{action taken}, but it is easy to see that the two are equivalent up to a change of the values of $R, S, T, P$.
+One could also imagine modeling noise in \textit{action taken}, but it is easy to see that the two are equivalent up to a change in the values of $R, T, S$ and $P$.
+
+We can now begin to define the outcome of two strategies playing against each other in the infinitely repeated game. We do this using Markov chains.
 
@@
   Suppose that strategy $s_1$ plays against strategy $s_2$.
-This defines an \textit{$s_1$-$s_2$ Markov chain} where each state $x$ is the vector $(c_1,c_2)$ where $c_1$ is a state in $s_1$ and $c_2$ is a state in $s_2$.
-The transition probabilities are defined in the obvious way, using the error probability $p$.
+This defines an \textit{$s_1$-$s_2$ Markov chain} where each state is a tuple $(c_1,c_2)$ with $c_1$ being a state in $s_1$ and $c_2$ a state in $s_2$.
+The transition probabilities are defined in the obvious way:
+  \begin{equation*}
+    p_{(c_1,c_2) \to (c_1', c_2')} = p_{c_1 \to c_1'} \cdot p_{c_2 \to c_2'},
+  \end{equation*}
+  where $p_{c_1 \to c_1'}$ is $1-p$ if the output of $s_2$ at state $c_2$ causes $c_1$ to transition to $c_1'$, and is $p$ if the opposite of that output causes that transition; $p_{c_2 \to c_2'}$ is defined similarly.
 \end{definition}
 
+Markov chains naturally leads themselves to the study of limiting cases.
+
+\begin{definition}
+  \label{timeaveragedistribution}
+  The \textit{time average distribution} of the $s_1$-$s_2$ Markov chain given the start state $(a, b)$, denoted $\pi^{(a, b)}$, is the distribution such that \begin{equation*}
+    \pi_{c_1, c_2}^{(a, b)} = E \left[ \text{fraction of time in state $(c_1,c_2)$} \mid \text{initial state is } (a, b)  \right]
+  \end{equation*}
+  where the expected fraction of time is taken over the infinite sequence $(X_0, X_1, \ldots)$.
+\end{definition}
+
+We will often use $\pi$ to refer to $\pi^{(a_0, b_0)}$ where $a_0$ is the initial state of $s_1$ and $b_0$ is the initial state of $s_2$.
+
+\begin{definition}
+  \label{payofftimeaverage}
+  The payoff that $s_1$ gets when playing against $s_2$ is
+  \begin{equation*}
+    v_{s_1}(s_2) = \sum_{(c_1,c_2)} \pi_{c_1,c_2} \cdot r(G_{s_1}(c_1),G_{s_2}(c_2))
+  \end{equation*}
+  where $G_s(c)$ is the output of strategy $s$ at state $c$. The sum is taken over all states $(c_1,c_2)$ in the $s_1$-$s_2$ Markov chain.
+\end{definition}
+
+For notational convenience, we may also make $r(G_{s_1}(c_1),G_{s_2}(c_2))$ into a vector, denoted by $r$, and write this as the dot product \begin{equation*}
+  v_{s_1}(s_2) = \pi \cdot r.
+\end{equation*}
+
+That is, the payoff that $s_1$ gets when playing against $s_2$ is simply an average of the reward it gets in each possible state of the Markov chain weighted by the fraction of time that is spent there.
+
+\iffalse
 We use the notation $G_{s_1,s_2}(c_1,c_2)$ to refer to the vector $(G_{s_1}(c_1), G_{s_2}(c_2))$, and we use $S_{s_1,s_2}$ to refer to the set of all states in the Markov chain.
@@
 
+\fi
+
 We're now ready to look at how strategies interact.
@@
 
-One can think of this as saying that we have infinitely many members of the population, and that they all interact with everyone else.
-This justifies the usage of expectation when definining $v_{s_1}(s_2)$.
+One can think of this as saying that we have infinitely many members of the population, interacting with each other evenly, and that the fitness of a strategy is its expected payoff.
+Having infinitely many interactions like this justifies the usage of expectation when definining $v_{s_1}(s_2)$.
 
 We can now use the fitness of a strategy to compare it with other strategies in the same population.
-If a strategy $s_1$ has a higher fitness than another strategy $s_2$, that means that the frequency of $s_1$ will increase on the expense of the frequency of $s_2$, in the next step of the evolutionary process.
+If a strategy $s_1$ has a higher fitness than another strategy $s_2$, we say that the frequency of $s_1$ will increase at the expense of the frequency of $s_2$, in the next step of the evolutionary process.
 This is getting us close to how we want to define stable strategies; our next move is looking not only at a single evolutionary step, but the entire evolutionary process.
@@
 
-That is, if $s_1$ is $\epsilon$-invadable, there exists a strategy $s_2$ that can start as only a tiny fraction $\epsilon$ of the total population, and consistently have higher fitness than $s_1$, eventually causing overtaking $s_1$ completely.
-We are now finally ready to state our main definition.
+That is, if $s_1$ is $\epsilon$-invadable, there exists a strategy $s_2$ that can start as only a tiny fraction $\epsilon$ of the total population, and consistently have higher fitness than $s_1$, eventually causing $s_2$ to overtake and then eliminate $s_1$ completely.
+We are now finally ready to define evolutionary stability.
 
 \begin{definition}
-  A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\epsilon_0$, both in $(0,1)$, such that for all $p < p_0$, and all $\epsilon < \epsilon_0$, $s_1$ is not $\epsilon$-invadable.
+  A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\epsilon_0$ with $0 < p_0, \epsilon_0 < 1$, such that for all $p < p_0$, and all $\epsilon < \epsilon_0$, $s_1$ is not $\epsilon$-invadable.
 \end{definition}
 
-That is, a strategy $s_1$ is evolutionarily stable if it can withstand invasion attempts from any strategy that starts off in low numbers, as the probability of noise tends to 0.
+\todo{separate this definition out into a pure evolutionarily stable definition, and then an evolutionarily stable for infinitesimally small probabilities definition??
+would make interpretation much clearer.}
+
+That is, a strategy $s_1$ is evolutionarily stable if, as the noise probability $p$ goes to 0, it can withstand invasion attempts from any strategy that starts off as a tiny fraction of the population.
+
+\subsection{Interpretation}
+
+Having concluded our formal definition of evolutionary stability, it might be helpful to see examples of concrete setups in which this definition makes sense.
+
+Consider the following world.
+
+\todo{WRITE THIS SECTION??
+I'M NOT SURE THAT MY MODEL MAKES MUCH SENSE ANYMORE THOUGH, BECAUSE IT REQUIRES PEOPLE TO PLAY AGAINST THEMSELVES WHICH IS A NONO.
+also the probability going to 0 does not really make sense}
 
@@
 
-We can now state our results! Together, the following two theorems prove that in the setup described here, mutual cooperation arises as the only stable choice.
+We can now state our results! Together, the following two conjectures imply that in the setup described here, mutual cooperation arises as the only stable choice.
 
-\begin{theorem}
+\begin{conjecture}
   \label{evolutionarystable1}
   Suppose that a strategy $s_1$ is evolutionarily stable. Then $\lim_{p \to 0} v_{s_1}(s_1) = R$.
-\end{theorem}
+\end{conjecture}
 
@@
   \label{pavlovtheorem}
-  Suppose $2R > T + P$. Then, the Pavlov strategy is evolutionarily stable.
+  Suppose $2R > T + P$. Then, the Pavlov strategy, depicted in \cref{pavlovfigure}, is evolutionarily stable.
 \end{conjecture}
@@
   \centering
-  \caption{Pavlov.}
+  \caption{The Pavlov strategy. Similar to but not the same as the Tit-for-tat strategy.}
 \end{figure}
 
-  % \begin{tikzpicture}[scale=0.8, show background rectangle]
-  %   \draw (0,0) -- (2,1);
-  %   \draw (1,1) circle (3);
-  %   \path[draw] (0,0) -- (1,1) -- (2,0);
-  % \end{tikzpicture}
-
 \begin{remark}
-  Tit-for-tat, displayed in \cref{tftfigure}, is not evolutionarily stable.
-It has the stationary distribution $(\frac{1}{4},\frac{1}{4},\frac{1}{4},\frac{1}{4})$ in its own Markov chain, which has a payoff that is significantly smaller than $R$.
-This should be intuitive: if tit-for-tat makes one mistake, it goes into a defection cycle that it doesn't break out of until it makes a second mistake.
-
+  Tit-for-tat, depicted in \cref{tftfigure}, is not evolutionarily stable. 
+  When played against itself, it ends up spending just as much time in the defection state as in the cooperation state, because as soon as one mistake is made, it goes into a mutual defection cycle with its clone that is only broken by an additional mistake.
+  Thus, $v_s(s)$ where $s$ is Tit-for-tat is much smaller than $R$, which means that it is not evolutionarily stable by \cref{evolutionarystable1}.
 \end{remark}
@@
 \label{sectionproofs}
+
+In this section, we prove our results from the previous section.
 
@@
 
+    \iffalse
+
     \section{Appendix: Time average distributions}
@@
 
+    \fi
+
     \section{Appendix: Probablistic Automata}
+
+    OHHHHHHH. THIS WILL HELP WITH MY PROOF????? THIS IS EXACTLY WHAT I'M TRYING TO DO IN MY PROOF RIGHT????? YEAHHHH, NO.
 

### 2020-05-21 paper/paper.tex
prove the main theorem

@@
 \usepackage{blkarray}
-\usetikzlibrary{backgrounds,arrows,automata,positioning}
+\usetikzlibrary{backgrounds,arrows,automata,positioning,shapes,shapes.geometric}
 
@@
 \begin{document}
+\tikzset{elliptic state/.style={draw,ellipse, text centered, text width=10em}}
+\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
+    text width=5em, text centered, rounded corners, minimum height=4em]
+
 \maketitle
@@
 
+  The following proof is relatively involved, and we believe that a simpler proof exists, but haven't been able to find one.
+
     \begin{proof}[Proof of \cref{evolutionarystable1}]
@@
       \end{equation*}
-      By \cref{limitvexists} and \cref{neverbetterthanr}, this assumption implies that the limit is strictly less than $R$. Define $\gamma = v_{s_1}(s_1)$. Then,
+      By \cref{limitvexists} and \cref{neverbetterthanr}, this assumption implies that the limit is strictly less than $R$. Define $\gamma = \lim_{p \to 0} v_{s_1}(s_1)$. Then,
       \begin{equation*}
@@
 
-      We want to prove that $s_1$ is not evolutionarily stable. This would prove the theorem.
+      We want to prove that $s_1$ is not evolutionarily stable.
 
       To do that, we want to prove that for all $p_0, \epsilon_0 \in (0,1)$, there exists $p < p_0$ and $\epsilon < \epsilon_0$, such that $s_1$ is $\epsilon$-invadable.
-We choose $\epsilon = \epsilon_0 / 2$, and present a strategy $s_2$ that can invade $s_1$ for sufficiently small $p$.
+We choose $\epsilon = \epsilon_0 / 2$, and present a strategy $s_2$ that can invade $s_1$ for all $p$ that are sufficiently small.
 
       The underlying idea is to create $s_2$ such that $s_1$ can see no difference between itself and $s_2$, while $s_2$, on the other hand, can.
-In that case, we would have $v_{s_1}(s_2) = v_{s_2}(s_1) = v_{s_1}(s_1)$, and could construct $s_2$ such that it always cooperates when it recognizes itself, thereby yielding $v_{s_2}(s_2) = R$.
+In that case, we would have that $v_{s_1}(s_2) = v_{s_2}(s_1) = v_{s_1}(s_1)$, and could construct $s_2$ such that it always cooperates when it recognizes itself, thereby yielding $v_{s_2}(s_2) = R$.
 This would give $s_2$ a higher fitness than $s_1$.
@@
 
-      We create the strategy $s_2$ as follows.
-First, we copy all of $s_1$ into $s_2$.
-Let $\mathcal{C}_1$ be the absorbing strongly connected components of $s_1$.
+      We create the strategy $s_2$ as follows. First, we copy all of $s_1$ into $s_2$. Let $\mathcal{C}_1$ be the set of
+       absorbing strongly connected components of $s_1$.
 The key idea, now, is to replace each original absorbing strongly connected component $C_1 \in \mathcal{C}_1$ with a new absorbing component $C_2$, which has the capability of self identifying.
 
-      We construct $C_2$ as follows. First, it finds out in finite time what action $A$ that $s_1$ will take at some large finite future time $T$ (with probability $1 - O(p)$). 
-      \todo{THIS DOESN'T WORK.
-The problem is the circulation states: if $s_2$ goes into an absorbing SCC before $s_1$ does, our modified $C_2$ might affect the probabilities of $s_1$ going into certain absorbing SCCs.
-I think a way to fix that is as follows: You take $C_1$, and replace each of its individual states by an SCC, targeting a particular non-absorbing SCC of $s_1$.
-Make sure that they circulate in sync with each other.
-That means that we will be able to delay getting into the identification stage of $C_2$ which is exactly what we want.}
-      After that, it responds with the opposite action (which we will denote by $\bar{A}$) right after time $T$. 
-      If $s_2$ subsequently perceives that its opponent takes action $A$, it will transition into a copy of $C_1$, the absorbing strongly connected component that $C_2$ replaces.
-On the other hand, if $s_2$ perceives its opponent taking action $\bar{A}$, then it will transition into an absorbing strongly connected component that always cooperates.
-A schematic view of $C_2$ is presented in figure ????.
+      We now construct $C_2$, given $C_1$.
+\Cref{c2flowchart} shows the high-level construction of $C_2$.
+All of the steps succeed with probability $1 - O(p)$.
+The first step makes sure that $C_2$ does not deviate from $s_1$ at all, until $s_1$ has reached an absorbing strongly connected component.
+Once $s_1$ is in an absorbing strongly connected component, we already know that the time average distribution does not depend on the particular state in $s_1$, which means that even if $s_1$ now ``figures out'' that it is not playing against itself anymore, there is nothing it can do about it.
+The next step makes sure that, if $s_2$ is playing against itself, it is in sync; i.e., with probability $1 - O(p)$ they will both copies of $s_2$ will be in the same state after this step.
+The third step now figures out what action $s_1$ would take at some large finite time $T$, and then outputs the exact opposite action at time $T$.
+At that point, if its opponent outputs $s_1$'s expected action, it transitions into the absorbing strongly connected component $C_1$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and enters an always cooperating state.
+
+      \begin{figure}
+        \centering
+        \begin{tikzpicture}[node distance=0.9cm,initial text=]
+          \node[elliptic state, initial, initial where=above] (c_0) {1. Wait until $s_1$ is in an absorbing SCC, while simulating $C_1$.};
+          \node[elliptic state] (c_1) [below=of c_0] {2. Sync $s_2$.};
+          \node[elliptic state] (c_3) [below=of c_1] {3. Figure out what action that $s_1$ will take right after this.};
+          \node[state] (c_4) [below left=of c_3] {$C$};
+          \node[state] (c_5) [below right=of c_3] {$D$};
+          \node[state] (c_6) [below=of c_5] {$C$};
+          \node[state] (c_7) [below=of c_4] {$C_1$};
+
+          \path[->] (c_0) edge (c_1);
+          \path[->] (c_1) edge (c_3);
+          \path[->] (c_3) edge node [above left] {$s_1$ will take $D$} (c_4);
+          \path[->] (c_3) edge node [above right] {$s_1$ will take $C$} (c_5);
+          \path[->] (c_4) edge node [below, pos=0.1] {$C$} (c_6);
+          \path[->] (c_5) edge node [right] {$D$} (c_6);
+          \path[->] (c_4) edge node [left] {$D$} (c_7);
+          \path[->] (c_5) edge node [below, pos=0.1] {$C$} (c_7);
+          \path[->] (c_6) edge [loop right] (c_6);
+          % \path[->] (c_0) edge node [above right] {$\bar{\alpha}$} (c_1);
+          % \path[->] (c_1) edge [loop right] node [right] {$C,D$} ();
+          % \path[->] (c_0) edge node [above left] {$\alpha$} (c_prime);
+          % \path[->] (c) edge node [below left] {$\alpha$} (c_prime);
+        \end{tikzpicture}
+        \caption{High level construction of $C_2$, the absorbing component replacing $C_1$ when creating $s_2$ in the proof of \cref{evolutionarystable1}.
+Note that $C_2$ contains exactly 2 absorbing strongly connected components.}
+        \label{c2flowchart}
+      \end{figure}
+
       This concludes the construction of $s_2$.
+We now want to prove that this $s_2$ can in fact invade $s_1$.
+To do that, we will first assume that the below 3 claims are correct, i.e., that the construction is possible and works as specified, and finish the proof of our theorem using those assumptions.
+Then, we will prove the 3 claims.
 
-      \todo{add picture here when low on focus}
+      \begin{claim}
+        \label{claimcanwaitfors1}
+        We can construct a non-absorbing component with indistinguishable output from the component $C_1$, such that when we leave the component, $s_1$ will be in an absorbing strongly connected component with probability $1 - O(p)$.
+      \end{claim}
 
-      We now want to prove that $s_2$ can in fact invade $s_1$ for small $p$.
-First, we will assume that $C_2$ can find out the action $A$ that $s_1$ would take at time $T$, and finish the proof of our theorem using this assumption.
-After that, we will prove that that is indeed the case.
+      \begin{claim}
+        \label{claimcansyncs2}
+        We can construct a non-absorbing component placed after the first component, such that if $s_2$ plays against itself, both clones transition out of the component at the exact same time with probability $1 - O(p)$.
+      \end{claim}
+
+      \begin{claim}
+        \label{claimcanfigureout}
+        We can construct a non-absorbing component with a finite number of maximum steps that, supposing that $s_2$ plays against $s_1$, can figure out what action $s_1$ will take at the exact timestep following the transition out of the component.
+      \end{claim}
+
+      We now finish our proof.
 
@@
       \begin{align*}
-        v_{s_1}(s_1) &= \gamma \\
+        v_{s_1}(s_1) &= \gamma + O(p) \\
         v_{s_1}(s_2) &= \gamma + O(p) \\
@@
       \begin{subproof}
-        By definition, $v_{s_1}(s_1) = \gamma$. By \cref{timeaveragedistributiondecomposition}, the time average distribution of the $s_1$-$s_1$ Markov chain $\pi$ is \begin{equation*}
-          \pi = \sum_{(C_1, C_1') \in \mathcal{C}_1 \times \mathcal{C}_1} p_{(C_1, C_1')} \cdot \pi^{(C_1, C_1')}
-        \end{equation*}
-        When creating $s_2$, we replaced each $C_1$ by a $C_2$, containing two absorbing strongly connected components: a copy of $C_1$, which we call $C_{1, \text{copy}}$ and the always cooperating $C_c$.
+
+        That $v_{s_1}(s_1) = \gamma + O(p)$ follows by our definition of $\gamma$.
+
+        Consider first the case of $s_1$ playing against $s_2$.
+By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ reaches an absorbing strongly connected component $C_1'$.
+This means that the probability for reaching the absorbing component $(C_1', C_2)$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we hadn't replaced $C_1$ by $C_2$, which is just the probability of reaching the absorbing component $(C_1', C_1)$ in the $s_1$-$s_1$ Markov chain.
+Thus,
+        \begin{multline*}
+          P(\text{$s_1$-$s_2$ chain ends up in $(C_1', C_2)$}) \\ = (1 - O(p)) P(\text{$s_1$-$s_1$ chain ends up in $(C_1', C_1)$})
+        \end{multline*}
+        Now, by \cref{claimcanfigureout}, the probability that $s_2$ enters the absorbing strongly connected component $C_1$ after having gotten to $C_2$ is $1 - O(p)$, and that it enters the always cooperating component $C_c$ is $O(p)$.
+Thus,
+        \begin{align*}
+          P(\text{$s_1$-$s_2$ chain ends up in $(C_1',C_1)$}) &= P(\text{$s_1$-$s_1$ chain ends up in $(C_1',C_1)$}) + O(p) \\
+          P(\text{$s_1$-$s_2$ chain ends up in $(C_1',C_c)$}) &= O(p)
+        \end{align*}
+        That is, when $s_1$ plays against $s_2$, the absorbing strongly connected components are mostly the same, with mostly the same probabilities. 
+        Now, we can calculate $\tau$, the time average distribution of the $s_1$-$s_2$ Markov chain:
+
+        \todo{this is nonsensical here!! i should really try to explain this better!!!!!}
+
+        % By definition, $v_{s_1}(s_1) = \gamma$. By \cref{timeaveragedistributiondecomposition}, the time average distribution of the $s_1$-$s_1$ Markov chain $\pi$ is \begin{equation*}
+        %   \pi = \sum_{(C_1, C_1') \in \mathcal{C}_1 \times \mathcal{C}_1} p_{(C_1, C_1')} \cdot \pi^{(C_1, C_1')}
+        % \end{equation*}
+        % When creating $s_2$, we replaced each $C_1$ by a $C_2$, containing two absorbing strongly connected components: a copy of $C_1$, which we call $C_{1, \text{copy}}$ and the always cooperating $C_c$.
 This means that the condensation of the $s_1$-$s_2$ Markov chain is the same as the condensation of the $s_1$-$s_1$ Markov chain, except that each absorbing strongly connected component has been replaced by two.
@@
 Then, we find by our construction, that
-        \begin{align}
-          \label{c1c1copy}
-          p_{(C_1, C_{1, \text{copy}})} &= (1 - O(p)) p_{(C_1, C_1')} \\
-          p_{(C_1, C_c)} &= O(p) p_{(C_1, C_1')}
-        \end{align}
-        This equation is true because when $s_2$ plays against $s_1$, it identifies the action $A$ that $s_1$ will take at time $A$ with probability $(1 - O(p))$, and upon seeing that transitions to $C_{1, \text{copy}}$ with probability $(1-p)$.
+        % \begin{align}
+        %   \label{c1c1copy}
+        %   p_{(C_1, C_{1, \text{copy}})} &= (1 - O(p)) p_{(C_1, C_1')} \\
+        %   p_{(C_1, C_c)} &= O(p) p_{(C_1, C_1')}
+        % \end{align}
+        % This equation is true because when $s_2$ plays against $s_1$, it identifies the action $A$ that $s_1$ will take at time $A$ with probability $(1 - O(p))$, and upon seeing that transitions to $C_{1, \text{copy}}$ with probability $(1-p)$.
 
-        Now, we can calculate $\tau$, the time average distribution of the $s_1$-$s_2$ Markov chain:
         \begin{align*}
@@
 
-        For the last part of the claim, note that at time $T$, $s_2$ and its copy $s_2$ will be in the exact same state with probability at least $(1-p)^T = 1 - O(p)$, since if there are no mistakes they will both go to the exact same states.
-Thus, when $s_2$ outputs $\bar{A}$ to identify $s_1$, both of them will do that, and thus transition to an always cooperating state with probability $1 - O(p)$.
+        For the last part of the claim, consider the $s_2$-$s_2$ Markov chain.
+By \cref{claimcansyncs2}, both clones will be in the exact same position after the sync step.
+The third step takes only a finite number of steps by \cref{claimcanfigureout}.
+Thus, after leaving the third step of $C_2$, if no mistakes have happened since leaving state 2, both clones of $s_2$ will be in the exact same state.
+This happens with probability $(1-p)^N$ where $N$ is the finite number of states in step 3, and thus with probability $1 - O(p)$.
+Thus, in the identifying stage, both clones will output the same thing, and as we can see in \cref{c2flowchart}, they will then both transition to the always cooperating state with probability $1 - O(p)$.
 Therefore, $v_{s_2}(s_2) = R - O(p)$, which finishes the proof of the claim.
@@
 
-      We now return to the part we left out: how to construct $C_2$ such that it can find out which action that $s_1$ would take at time $T$.
+      We now return to the three claims we left out that detail the construction of $s_2$.
 

### 2020-05-21 paper/paper.tex
some more work

@@
   \label{payofftimeaverage}
-  The payoff that $s_1$ gets when playing against $s_2$ is
+  Let $r(c_1, c_2)$ refer to the reward that $s_1$ gets when $s_1$ takes the action in state $c_1$ and $s_2$ takes the action in state $c_2$.
+Then, the \textit{payoff that $s_1$ gets when playing against $s_2$} is
   \begin{equation*}
-    v_{s_1}(s_2) = \sum_{(c_1,c_2)} \pi_{c_1,c_2} \cdot r(G_{s_1}(c_1),G_{s_2}(c_2))
+    v_{s_1}(s_2) = \sum_{(c_1,c_2)} \pi_{c_1,c_2} \cdot r(c_1, c_2),
   \end{equation*}
-  where $G_s(c)$ is the output of strategy $s$ at state $c$. The sum is taken over all states $(c_1,c_2)$ in the $s_1$-$s_2$ Markov chain.
+  where the sum is taken over all states $(c_1,c_2)$ in the $s_1$-$s_2$ Markov chain.
 \end{definition}
 
-For notational convenience, we may also make $r(G_{s_1}(c_1),G_{s_2}(c_2))$ into a vector, denoted by $r$, and write this as the dot product \begin{equation*}
+For notational convenience, we may also make $r(c_1, c_2)$ into a vector, denoted by $r$, and write this as the dot product \begin{equation*}
   v_{s_1}(s_2) = \pi \cdot r.
@@
 
-We can now state our results! Together, the following two conjectures imply that in the setup described here, mutual cooperation arises as the only stable choice.
+We can now state our results! Together, the following theorem and conjecture imply that in the setup described here, mutual cooperation arises as the only stable choice.
 
-\begin{conjecture}
+\begin{theorem}
   \label{evolutionarystable1}
   Suppose that a strategy $s_1$ is evolutionarily stable. Then $\lim_{p \to 0} v_{s_1}(s_1) = R$.
-\end{conjecture}
+\end{theorem}
 
@@
   \label{pavlovtheorem}
-  Suppose $2R > T + P$. Then, the Pavlov strategy, depicted in \cref{pavlovfigure}, is evolutionarily stable.
+  Suppose that $2R > T + P$. Then, the Pavlov strategy, depicted in \cref{pavlovfigure}, is evolutionarily stable.
 \end{conjecture}
@@
 \begin{figure}
-  \label{pavlovfigure}
 \begin{tikzpicture}[node distance=2cm]
@@
   \caption{The Pavlov strategy. Similar to but not the same as the Tit-for-tat strategy.}
+  \label{pavlovfigure}
 \end{figure}
@@
   The time average distribution $\pi^{(a,b)}$, for any starting state $(a,b)$, is a stationary distribution of the Markov chain.
+That is, if $M$ is the transition matrix of the Markov chain, we have $M \pi^{(a,b)} = \pi^{(a,b)}$.
 \end{lemma}
@@
 \begin{lemma}
+  \label{lemmaabsorbinghasuniquetimeaverage}
   An absorbing strongly connected component has a unique time average distribution, i.e., the time average distribution does not depend on the start state.
@@
 
+\begin{proof}
+  It is well known that a strongly connected (also known as irreducible) Markov chain has a unique stationary distribution.
+Since the time average distribution is stationary by \cref{timeaverageisstationary}, it thus also has to be unique.
+\end{proof}
+
 \begin{lemma}
   \label{timeaveragedistributiondecomposition}
-  Let $\mathcal{S}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain.
-For every $C \in \mathcal{S}$, let $\pi^{(C)}$ be its unique time average distribution.
-Then, for some probabilities $p_C$ with $\sum_{C \in \mathcal{S}} p_C = 1$, depending only on the condensation of the Markov chain, we have \begin{equation*}
-    \pi = \sum_{C \in \mathcal{S}} p_C \cdot \pi^{(C)}.
+  Let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain.
+For every $S \in \mathcal{A}$, let $\pi^{(S)}$ be its unique time average distribution.
+Then, for some probabilities $p_{S}$ with $\sum_{S \in \mathcal{A}} p_{S} = 1$, corresponding to the time average probabilities of ending up in the absorbing strongly connected component $S$, we have \begin{equation*}
+    \pi = \sum_{S \in \mathcal{A}} p_{S} \cdot \pi^{(S)}.
   \end{equation*}
@@
   \label{absorbingstronglyconnectedcomponentscartesianproduct}
-  Let $\mathcal{C}_1$ be the set of absorbing SCCs of $s_1$, and similarly, let $\mathcal{C}_2$ be the set of absorbing SCCs of $s_2$.
-If $\mathcal{C}$ is the set of absorbing SCCs of the $s_1$-$s_2$ Markov chain, then it is the cartesian product of the two, i.e.,
+  Let $\mathcal{A}_1$ be the set of absorbing strongly connected components of $s_1$, and similarly, let $\mathcal{A}_2$ be the set of absorbing strongly connected components of $s_2$.
+If $\mathcal{A}$ is the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain, then
   \begin{equation*}
-    \mathcal{C} = \mathcal{C}_1 \times \mathcal{C}_2.
+    \mathcal{A} = \{H \times K \mid H \in \mathcal{A}_1 \text{ and } K \in \mathcal{A}_2\},
   \end{equation*}
+  where $A \times B$ is the cartesian product $A \times B = \{(a, b) \mid a \in A \text{ and } b \in B\}$.
 \end{lemma}
+
+Now that we know more about the time average distribution, we are also interested in proving that the payoff function behaves nicely.
 
@@
   \begin{equation*}
-    v_{s_1}(s_2) = \sum \pi_{c_1,c_2} \cdot r(c_1,c_2)
+    v_{s_1}(s_2) = \sum \pi_{c_1,c_2} \cdot r(c_1, c_2)
   \end{equation*}
@@
   \begin{equation*}
-    v_{s_2}(s_1) = \sum \pi_{c_2,c_1} \cdot r(c_2,c_1).
+    v_{s_2}(s_1) = \sum \pi_{c_2,c_1} \cdot r(c_2, c_1).
   \end{equation*}
@@
 
-\subsection{Evolutionary Stability Implies Utilitarianism}
+\subsection{Evolutionary Stability Requires Cooperation}
 
-  The following proof is relatively involved, and we believe that a simpler proof exists, but haven't been able to find one.
+  We are now finally ready to prove that to be evolutionarily stable, a strategy needs to cooperate with a clone of itself.
 
@@
       Suppose that the strategy $s_1$ is such that it is \textit{not} true that \begin{equation*}
-        \lim_{p \to 0 } v_{s_1}(s_1) = R
+        \lim_{p \to 0 } v_{s_1}(s_1) = R.
       \end{equation*}
@@
 
-      The underlying idea is to create $s_2$ such that $s_1$ can see no difference between itself and $s_2$, while $s_2$, on the other hand, can.
-In that case, we would have that $v_{s_1}(s_2) = v_{s_2}(s_1) = v_{s_1}(s_1)$, and could construct $s_2$ such that it always cooperates when it recognizes itself, thereby yielding $v_{s_2}(s_2) = R$.
+      The underlying idea is to construct $s_2$ such that $s_1$ can see no difference between itself and $s_2$, while $s_2$, on the other hand, can.
+If we succeed in doing so, we will have that $v_{s_1}(s_2) = v_{s_2}(s_1) = v_{s_1}(s_1)$, and can construct $s_2$ such that it always cooperates when it recognizes itself, thereby yielding $v_{s_2}(s_2) = R$.
 This would give $s_2$ a higher fitness than $s_1$.
@@
 
-      We create the strategy $s_2$ as follows. First, we copy all of $s_1$ into $s_2$. Let $\mathcal{C}_1$ be the set of
+      We create the strategy $s_2$ as follows. First, we copy all of $s_1$ into $s_2$. Let $\mathcal{A}_1$ be the set of
        absorbing strongly connected components of $s_1$.
-The key idea, now, is to replace each original absorbing strongly connected component $C_1 \in \mathcal{C}_1$ with a new absorbing component $C_2$, which has the capability of self identifying.
+The key idea, now, is to replace each original absorbing strongly connected component $H \in \mathcal{A}_1$ with a new absorbing component $K_H$, which has the capability of identifying itself.
 
-      We now construct $C_2$, given $C_1$.
-\Cref{c2flowchart} shows the high-level construction of $C_2$.
-All of the steps succeed with probability $1 - O(p)$.
-The first step makes sure that $C_2$ does not deviate from $s_1$ at all, until $s_1$ has reached an absorbing strongly connected component.
-Once $s_1$ is in an absorbing strongly connected component, we already know that the time average distribution does not depend on the particular state in $s_1$, which means that even if $s_1$ now ``figures out'' that it is not playing against itself anymore, there is nothing it can do about it.
-The next step makes sure that, if $s_2$ is playing against itself, it is in sync; i.e., with probability $1 - O(p)$ they will both copies of $s_2$ will be in the same state after this step.
-The third step now figures out what action $s_1$ would take at some large finite time $T$, and then outputs the exact opposite action at time $T$.
-At that point, if its opponent outputs $s_1$'s expected action, it transitions into the absorbing strongly connected component $C_1$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and enters an always cooperating state.
+      We now construct a $K_H$ for each $H$. \Cref{c2flowchart} shows the high-level construction of $K_H$.
+       All of the steps succeed with probability $1 - O(p)$. The first step makes sure that $K_H$ does not deviate from $s_1$ at all
+        until $s_1$ has reached an absorbing strongly connected component. 
+       Once $s_1$ is in an absorbing strongly connected component, \cref{lemmaabsorbinghasuniquetimeaverage} tells us that the time average distribution does not depend on the particular start state in that component, which means that even if $s_1$ after this point ``figures out'' that it is not playing against itself anymore, there is nothing it can do about it (since it also cannot escape the component).
+      The second step in the construction of $K_H$ makes sure that if $s_2$ is playing against itself, it is in sync with its clone; that is, with probability $1 - O(p)$, both copies of $s_2$ will be in the exact same state after the second step.
+      The third step now figures out what action $s_1$ would take at some large finite time $T$. 
+      Then, $K_H$ outputs the exact opposite action at time $T$. 
+      At that point, if its opponent outputs $s_1$'s expected action, it transitions into the absorbing strongly connected component $H_\text{copy}$, which is just an exact copy of $H$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and enters an always cooperating state.
 
@@
         \begin{tikzpicture}[node distance=0.9cm,initial text=]
-          \node[elliptic state, initial, initial where=above] (c_0) {1. Wait until $s_1$ is in an absorbing SCC, while simulating $C_1$.};
+          \node[elliptic state, initial, initial where=above] (c_0) {1. Wait until $s_1$ is in an absorbing SCC, while simulating $H$.};
           \node[elliptic state] (c_1) [below=of c_0] {2. Sync $s_2$.};
@@
           \node[state] (c_6) [below=of c_5] {$C$};
-          \node[state] (c_7) [below=of c_4] {$C_1$};
+          \node[elliptic state] (c_7) [below=of c_4] {$H_\text{copy}$ \\(an exact copy of $H$)};
 
@@
         \end{tikzpicture}
-        \caption{High level construction of $C_2$, the absorbing component replacing $C_1$ when creating $s_2$ in the proof of \cref{evolutionarystable1}.
-Note that $C_2$ contains exactly 2 absorbing strongly connected components.}
+        \caption{High level construction of $K_H$, the absorbing component replacing $H$ when creating $s_2$ in the proof of \cref{evolutionarystable1}.
+Note that $K_H$ contains exactly 2 absorbing strongly connected components.}
         \label{c2flowchart}
@@
 
-      This concludes the construction of $s_2$.
+      This concludes the high-level construction of $s_2$.
 We now want to prove that this $s_2$ can in fact invade $s_1$.
-To do that, we will first assume that the below 3 claims are correct, i.e., that the construction is possible and works as specified, and finish the proof of our theorem using those assumptions.
-Then, we will prove the 3 claims.
+To do that, we will first assume that the below 3 claims are correct, i.e., that the construction is possible and works as specified, and then finish the proof of our theorem using those assumptions.
+Once that's done, we will prove the 3 claims.
 
@@
         \label{claimcanwaitfors1}
-        We can construct a non-absorbing component with indistinguishable output from the component $C_1$, such that when we leave the component, $s_1$ will be in an absorbing strongly connected component with probability $1 - O(p)$.
+        We can construct a non-absorbing component with behavior indistinguishable from the behavior of component $H$, such that when we leave the component, $s_1$ will be in an absorbing strongly connected component with probability $1 - O(p)$.
       \end{claim}
@@
 
-        That $v_{s_1}(s_1) = \gamma + O(p)$ follows by our definition of $\gamma$.
+        By our definition of $\gamma$, we have that $v_{s_1}(s_1) = \gamma + O(p)$.
 
-        Consider first the case of $s_1$ playing against $s_2$.
-By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ reaches an absorbing strongly connected component $C_1'$.
+        Consider now the case of $s_1$ playing against $s_2$.
+Suppose that $s_2$ ends up in the absorbing component $K_H$.
+By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ also reaches an absorbing strongly connected component, which we call $H'$.
 This means that the probability for reaching the absorbing component $(C_1', C_2)$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we hadn't replaced $C_1$ by $C_2$, which is just the probability of reaching the absorbing component $(C_1', C_1)$ in the $s_1$-$s_1$ Markov chain.
@@
         &= (1 - \epsilon) (\gamma - O(p)) + \epsilon (R - O(p)) - (1-\epsilon) \gamma - \epsilon (\gamma + O(p)) \\
-        &\leq \epsilon (R - \gamma) + O(p)
+        &\geq \epsilon (R - \gamma) - O(p)
       \end{align*}

### 2020-05-21 paper/paper.tex
finish main theorem

@@
 \usepackage{blkarray}
-\usetikzlibrary{backgrounds,arrows,automata,positioning,shapes,shapes.geometric}
+\usetikzlibrary{backgrounds,arrows,automata,positioning,shapes,shapes.geometric,calc, arrows.meta, bending, decorations.pathreplacing}
+
+\usepackage{witharrows}
 
@@
   \label{timeaveragedistributiondecomposition}
-  Let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain.
-For every $S \in \mathcal{A}$, let $\pi^{(S)}$ be its unique time average distribution.
-Then, for some probabilities $p_{S}$ with $\sum_{S \in \mathcal{A}} p_{S} = 1$, corresponding to the time average probabilities of ending up in the absorbing strongly connected component $S$, we have \begin{equation*}
-    \pi = \sum_{S \in \mathcal{A}} p_{S} \cdot \pi^{(S)}.
+  Let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain. For every $S \in \mathcal{A}$, let $\pi^{(S)}$ be its unique time average distribution. Then
+  \begin{equation*}
+    \pi = \sum_{S \in \mathcal{A}} P(\text{$s_1$-$s_2$ chain ends up in $S$}) \cdot \pi^{(S)}.
   \end{equation*}
@@
       \begin{align*}
-        v_{s_1}(s_1) &= \gamma + O(p) \\
-        v_{s_1}(s_2) &= \gamma + O(p) \\
-        v_{s_2}(s_1) &= \gamma + O(p) \\
-        v_{s_2}(s_2) &= R + O(p)
+        v_{s_1}(s_1) &= \gamma \pm O(p) \\
+        v_{s_1}(s_2) &= \gamma \pm O(p) \\
+        v_{s_2}(s_1) &= \gamma \pm O(p) \\
+        v_{s_2}(s_2) &= R \pm O(p)
       \end{align*}
@@
 
-        By our definition of $\gamma$, we have that $v_{s_1}(s_1) = \gamma + O(p)$.
+        By our definition of $\gamma$, we have that $v_{s_1}(s_1) = \gamma \pm O(p)$.
 
@@
 Suppose that $s_2$ ends up in the absorbing component $K_H$.
-By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ also reaches an absorbing strongly connected component, which we call $H'$.
-This means that the probability for reaching the absorbing component $(C_1', C_2)$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we hadn't replaced $C_1$ by $C_2$, which is just the probability of reaching the absorbing component $(C_1', C_1)$ in the $s_1$-$s_1$ Markov chain.
+By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ reaches an absorbing strongly connected component, which we call $H'$.
+        This means that the probability of reaching the absorbing component $H' \times K_H$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we hadn't replaced $H$ by $K_H$ in $s_2$, which is just the probability of reaching the absorbing component $H' \times H$ in the $s_1$-$s_1$ Markov chain.
 Thus,
         \begin{multline*}
-          P(\text{$s_1$-$s_2$ chain ends up in $(C_1', C_2)$}) \\ = (1 - O(p)) P(\text{$s_1$-$s_1$ chain ends up in $(C_1', C_1)$})
+          P(\text{$s_1$-$s_2$ ends up in $H' \times K_H$}) = (1 - O(p)) P(\text{$s_1$-$s_1$ ends up in $H' \times H$})
         \end{multline*}
-        Now, by \cref{claimcanfigureout}, the probability that $s_2$ enters the absorbing strongly connected component $C_1$ after having gotten to $C_2$ is $1 - O(p)$, and that it enters the always cooperating component $C_c$ is $O(p)$.
+        Now, by \cref{claimcanfigureout}, the probability that $s_2$ enters the absorbing strongly connected component $H_\text{copy}$ after having gotten to $K_H$ is $1 - O(p)$ when playing against $s_1$.
 Thus,
-        \begin{align*}
-          P(\text{$s_1$-$s_2$ chain ends up in $(C_1',C_1)$}) &= P(\text{$s_1$-$s_1$ chain ends up in $(C_1',C_1)$}) + O(p) \\
-          P(\text{$s_1$-$s_2$ chain ends up in $(C_1',C_c)$}) &= O(p)
-        \end{align*}
-        That is, when $s_1$ plays against $s_2$, the absorbing strongly connected components are mostly the same, with mostly the same probabilities. 
-        Now, we can calculate $\tau$, the time average distribution of the $s_1$-$s_2$ Markov chain:
-
-        \todo{this is nonsensical here!! i should really try to explain this better!!!!!}
+        \begin{equation}
+          \label{equationhhcopy}
+          P(\text{$s_1$-$s_2$ ends up in $H' \times H_\text{copy}$}) = P(\text{$s_1$-$s_1$ ends up in $H' \times H$}) \pm O(p).
+        \end{equation}
+        Let $C_H$ be the always cooperating absorbing state of $K_H$. Then, this equation implies that 
+        \begin{equation}
+          \label{equationhch}
+          P(\text{$s_1$-$s_2$ ends up in $H' \times C_H$}) = O(p).
+        \end{equation}
+        Thus, when $s_1$ plays against $s_2$, the absorbing strongly connected components are mostly the same, with mostly the same probabilities. 
+        Now, let $\pi$ be the time average distribution of the $s_1$-$s_1$ Markov chain and let $\tau$ be the time average distribution of the $s_1$-$s_2$ Markov chain.
+Let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_1$ chain, and let $\mathcal{B}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ chain.
+Using \cref{timeaveragedistributiondecomposition} and \cref{absorbingstronglyconnectedcomponentscartesianproduct}, we can now calculate $\tau$:
 
@@
         \begin{align*}
-          \tau &= \sum_{(C_1, C') \in \mathcal{C}_1 \times \mathcal{C}_2} p_{(C_1, C')} \cdot \pi^{(C_1, C')} \\
-          &= \sum_{(C_1, C_{1, \text{copy}})} p_{(C_1, C_{1, \text{copy}})} \cdot \pi^{(C_1, C_{1, \text{copy}})} + \sum_{(C_1, C_c)} p_{(C_1, C_c)} \cdot \pi^{(C_1, C_c)} \\
-          &= \sum_{(C_1, C_1') \in \mathcal{C}_1 \times \mathcal{C}_1} p_{(C_1, C_1')} \cdot \pi^{(C_1, C_1')} + O(p) \\
-          &= \pi + O(p)
+          \WithArrowsOptions{displaystyle}
+          \begin{WithArrows}
+          \tau &= \sum_{S \in \mathcal{B}} P(\text{$s_1$-$s_2$ ends up in $S$}) \cdot \pi^{(S)} \Arrow{\cref{absorbingstronglyconnectedcomponentscartesianproduct}}\\
+          &= \sum_{H', C_H} P(\text{$s_1$-$s_2$ ends up in $H' \times C_H$}) \cdot \pi^{(H' \times C_H)}\\ &\:\:\:\:\:\:\:+ \sum_{H', H_\text{copy}} P(\text{$s_1$-$s_2$ ends up in $H' \times H_\text{copy}$}) \cdot \pi^{(H' \times H_\text{copy})} \Arrow{\cref{equationhch}}\\
+          &= \sum_{H', H_\text{copy}} P(\text{$s_1$-$s_2$ ends up in $H' \times H_\text{copy}$}) \cdot \pi^{(H' \times H_\text{copy})} \pm O(p) \Arrow{\cref{equationhhcopy}}\\
+          &= \sum_{H', H} P(\text{$s_1$-$s_1$ ends up in $H' \times H$}) \cdot \pi^{(H' \times H)} \pm O(p) \Arrow{\cref{absorbingstronglyconnectedcomponentscartesianproduct}}\\
+          &= \sum_{S \in \mathcal{A}} P(\text{$s_1$-$s_1$ ends up in $S$}) \cdot \pi^{(S)} \pm O(p)\\
+          &= \pi \pm O(p)
+          \end{WithArrows}
         \end{align*}
-        where the second-to-last equality follows by \cref{c1c1copy}. \todo{we're kinda abusing notation here} This proves that $v_{s_1}(s_2) = \gamma + O(p)$ and $v_{s_2}(s_1) = \gamma + O(p)$.
+        Thus, $s_1$-$s_1$ and $s_1$-$s_2$ have almost the same time average distribution. This proves that $v_{s_1}(s_2) = \gamma \pm O(p)$ and $v_{s_2}(s_1) = \gamma \pm O(p)$.
 
@@
 The third step takes only a finite number of steps by \cref{claimcanfigureout}.
-Thus, after leaving the third step of $C_2$, if no mistakes have happened since leaving state 2, both clones of $s_2$ will be in the exact same state.
-This happens with probability $(1-p)^N$ where $N$ is the finite number of states in step 3, and thus with probability $1 - O(p)$.
+Thus, after leaving the third step of $K_H$, if no mistakes have happened since leaving state 2, both clones of $s_2$ will be in the exact same state.
+This happens with probability at least $(1-p)^N$ where $N$ is the finite number of states in step 3, and thus with probability $1 - O(p)$.
 Thus, in the identifying stage, both clones will output the same thing, and as we can see in \cref{c2flowchart}, they will then both transition to the always cooperating state with probability $1 - O(p)$.
-Therefore, $v_{s_2}(s_2) = R - O(p)$, which finishes the proof of the claim.
+Therefore, $v_{s_2}(s_2) = R \pm O(p)$, which finishes the proof of the claim.
       \end{subproof}
@@
         &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
-        &= (1 - \epsilon) (\gamma - O(p)) + \epsilon (R - O(p)) - (1-\epsilon) \gamma - \epsilon (\gamma + O(p)) \\
-        &\geq \epsilon (R - \gamma) - O(p)
+        &= (1 - \epsilon) (\gamma \pm O(p)) + \epsilon (R \pm O(p)) - (1-\epsilon) (\gamma \pm O(p)) - \epsilon (\gamma \pm O(p)) \\
+        &\geq \epsilon (R - \gamma) \pm O(p)
       \end{align*}
@@
       We know that $R - \gamma > 0$ by our initial assumption.
-For a sufficiently small $p$, thus, $F(s_2) - F(s_1) > 0$.
-This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is not $\epsilon$-invadable for this value of $p$.
+For all sufficiently small $p$, thus, $F(s_2) - F(s_1) > 0$.
+This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is not $\epsilon$-invadable some value of $p < p_0$.
 In conclusion, then, $s_1$ is not evolutionarily stable, which concludes the proof of \cref{evolutionarystable1}.

### 2020-05-22 paper/paper.tex
proof of everything but lemma 3

@@
        Once $s_1$ is in an absorbing strongly connected component, \cref{lemmaabsorbinghasuniquetimeaverage} tells us that the time average distribution does not depend on the particular start state in that component, which means that even if $s_1$ after this point ``figures out'' that it is not playing against itself anymore, there is nothing it can do about it (since it also cannot escape the component).
-      The second step in the construction of $K_H$ makes sure that if $s_2$ is playing against itself, it is in sync with its clone; that is, with probability $1 - O(p)$, both copies of $s_2$ will be in the exact same state after the second step.
+      The second step in the construction of $K_H$ makes sure that if $s_2$ is playing against itself, it is in sync with its clone; that is, with probability $1 - O(p)$, both copies of $s_2$ will transition to the third component at the exact same time (although not necessarily in the same $K_H$).
       The third step now figures out what action $s_1$ would take at some large finite time $T$. 
@@
 We now want to prove that this $s_2$ can in fact invade $s_1$.
-To do that, we will first assume that the below 3 claims are correct, i.e., that the construction is possible and works as specified, and then finish the proof of our theorem using those assumptions.
-Once that's done, we will prove the 3 claims.
+To do that, we will first assume that the below 3 lemmas are correct, i.e., that the construction is possible and works as specified, and then finish the proof of our theorem using those assumptions.
+Once that's done, we will prove the 3 lemmas.
 
-      \begin{claim}
+      \begin{lemma}
         \label{claimcanwaitfors1}
         We can construct a non-absorbing component with behavior indistinguishable from the behavior of component $H$, such that when we leave the component, $s_1$ will be in an absorbing strongly connected component with probability $1 - O(p)$.
-      \end{claim}
+      \end{lemma}
 
-      \begin{claim}
+      \begin{lemma}
         \label{claimcansyncs2}
-        We can construct a non-absorbing component placed after the first component, such that if $s_2$ plays against itself, both clones transition out of the component at the exact same time with probability $1 - O(p)$.
-      \end{claim}
+        We can construct a non-absorbing component placed after the first component, such that if $s_2$ plays against itself, both clones transition from the second component to the third component of their respective $K_H$ at the exact same time, with probability $1 - O(p)$.
+      \end{lemma}
 
-      \begin{claim}
+      \begin{lemma}
         \label{claimcanfigureout}
         We can construct a non-absorbing component with a finite number of maximum steps that, supposing that $s_2$ plays against $s_1$, can figure out what action $s_1$ will take at the exact timestep following the transition out of the component.
-      \end{claim}
+This component is the same for all $H$.
+      \end{lemma}
 
-      We now finish our proof.
+      We now finish our proof, starting with a lemma concerning the payoffs.
 
-      \begin{claim}
+      \begin{lemma}
         \label{claimpayoffs}
@@
       \end{align*}
-      \end{claim}
+      \end{lemma}
       \begin{subproof}
@@
         For the last part of the claim, consider the $s_2$-$s_2$ Markov chain.
-By \cref{claimcansyncs2}, both clones will be in the exact same position after the sync step.
-The third step takes only a finite number of steps by \cref{claimcanfigureout}.
-Thus, after leaving the third step of $K_H$, if no mistakes have happened since leaving state 2, both clones of $s_2$ will be in the exact same state.
+By \cref{claimcansyncs2}, both clones will transition enter the third component at the exact same time.
+The third step takes only a finite number of steps by \cref{claimcanfigureout}, and is the same regardless of which $K_H$ the strategies are in.
+Thus, after leaving the third step of $K_H$, if no mistakes have happened since leaving state 2, both clones of $s_2$ will be in exactly corresponding states, since they will always mirror each other.
 This happens with probability at least $(1-p)^N$ where $N$ is the finite number of states in step 3, and thus with probability $1 - O(p)$.
 Thus, in the identifying stage, both clones will output the same thing, and as we can see in \cref{c2flowchart}, they will then both transition to the always cooperating state with probability $1 - O(p)$.
-Therefore, $v_{s_2}(s_2) = R \pm O(p)$, which finishes the proof of the claim.
+Therefore, $v_{s_2}(s_2) = R \pm O(p)$, which finishes the proof of the lemma.
       \end{subproof}
@@
 
-      We now return to the three claims we left out that detail the construction of $s_2$.
+      We now return to the three lemmas that we left out, which detail the construction of $s_2$.
 
-      \begin{claim}
-        There is a $T$ and a procedure implementable in $C_2$ that runs in a finite number of steps and can determine the action $A$ of $s_1$ at time $T$ with probability $1 - O(p)$.
-      \end{claim}
+      \begin{proof}[Proof of \cref{claimcanwaitfors1}]
 
-      \begin{proof}
-          Let $\mathcal{C}_1$ be the set of all absorbing strongly connected components in $s_1$.
-           Let $M$ be the set of all tuples of $(C, c)$ where $C \in \mathcal{C}_1$ and $c$ is a possible state of $C$.
+        Let $T_1$ be the random variable designating the time at which $s_1$ transitions into an absorbing strongly connected component, counting from the time $s_2$ entered $K_H$.
+(In particular, if $s_1$ is already in an absorbing strongly connected component, $T_1 = 0$.)
+        Let $T_2$ be the random variable designating the time at which $s_2$ transitions out of component 1 of $K_H$.
+        To prove our lemma, we want to prove that with probability $1 - O(p)$, $T_1 \leq T_2$, and that during all the time that we are in component 1 of $K_H$, the output of $s_2$ is indistinguishable from $H$.
+
+        First, we deal with the indistinguishableness part.
+We create the component by copying $H$.
+Next, we define a new graph $W$, that we will design in the subsequent paragraphs, which will wait for $s_1$ to get into an absorbing component.
+This graph will have edges defined with the labels $C$ and $D$ still, but have no outputs at the states.
+The graph $D$ will also have one designated ``end state.'' We will then construct our component 1 for $K_H$ by replacing every state $h \in H$ with a copy of $D$ that we call $D_h$.
+There will be a $C$ edge from node $u \in D_h$ to node $v \in D_{h'}$ if and only if there is a $C$ edge from $h$ to $h'$, and from $u$ to $v$ in $D$.
+Similarly for the $D$ edge.
+This is commonly known as the Kronecker or tensor product of the graphs $D$ and $H$.
+The output at every state $u \in D_h$ is the same as the output of $h$.
+For the end state $e \in D_h$ for each $h$, we will add an edge with label both $C$ and $D$, transitioning out of this component and into component 2.
+
+        It is clear that as long as we don't transition out of this component, the Kronecker product will ensure that the output will be indistinguishable from $H$.
+We now want to design $D$ such that $T_1 \leq T_2$ with high probability.
+
+        Let $d$ be the maximum length of a path from any state in $s_1$ to its closest absorbing strongly connected component.
+Then, for each state in $s_1$, there is a sequence of perceived inputs of length $\leq d$ (the \textit{special sequence} of that state), such that $s_1$ upon seeing that sequence ends up in an absorbing strongly connected component.
+For all states, the probability that $s_1$
+        perceives its special sequence is $\geq p^d$, since, in the worst case, $s_1$ would need to make a mistake on exactly all 
+        inputs to perceive the special sequence. That means that at any single point in time, there is a probability $\geq p^d$ that $s_1$ 
+        will go into an absorbing strongly connected component within the subsequent $d$ moves.
+Consider the geometric random variable $X_1 = d + \operatorname{Geom} (p^d)$.
+For all $t$, then, we have that $P(X_1 = t) \leq P(T_1 = t)$.
+Thus, we may see $X_1$ as some sort of upper bound to $T_1$.
+4
+        We now describe the construction of $D$.
+Consider a state $c_1 \in s_1$ and a state $h \in H$, and suppose that we let them evolve with no mistakes happening at all.
+Then, the outputs of $c_1$ will form an infinite string $w(c_1, h)$.
+Let there be $|s_1|$ states in $s_1$ and $|H|$ states in $H$.
+Now, create a string $w^*$ of length $|s_1| \cdot |H|$, which differs in at least 1 place from each $w(c_1,h)$.\footnote{This is the famous technique of diagonalization.} Thus, if $s_1$ ever outputs the string $w^*$ when we are in this component, we know that at least one mistake has occurred, somewhere.
+Consequently, the probability that, at each timestep, $w^*$ will be perceived by $s_2$ as the next sequence of outputs of $s_1$ is $\leq p$.
+        If we repeat $w^*$ $d+1$ times to form $(w^*)^{d+1}$, the probability that that string is perceived by $s_2$ as the next sequence of outputs of $s_1$ is $\leq p^{d+1}$.
+        Now, create $D$ as follows: it is a chain of states $d_1,d_2,\ldots,d_{(d+1)|w^*|+1}$ where $d_i$ transitions to $d_{i+1}$ on seeing the $i$th character of the string $(w^*)^{d+1}$; $d_i$ transitions to $d_1$ otherwise.
+$d_{(d+1)|w^*|+1}$ is the designated end state of $D$.
+Now, it is clear that each step, the probability that we will reach the end state in the next $(d+1)|w^*|+1$ steps is $\leq p^{d+1}$.
+Consider the geometric random variable $X_2 = (d+1)|w^*| + 1 + \operatorname{Geom} (p^{d+1})$.
+For all $t$, then, we have that $P(X_2 = t) \geq P(T_2 = t)$.
+Thus, we may see $X_2$ as a lower bound to $T_2$.
+
+        We now claim that the probability that $X_1 \leq X_2$ is $1 - O(p)$, where $X_1$ and $X_2$ are independent.
+It suffices to show that $\operatorname{Geom}(p^d) \leq \operatorname{Geom}(p^{d+1})$ with high probability, and it further suffices to show that $\operatorname{Geom}(x) > \operatorname{Geom}(px)$ with probability $O(p)$, for some arbitrary $x$.
+This is easy to show.
+Consider the geometric variable with parameter $px$ to consist of first an event of probability $x$ happening, and then that counts as a real success with a probability $p$.
+Then, when comparing the two geometric variables, it is clear that both are equally likely to have an event of probability $x$ happening first.
+Thus, with probability $\frac{1}{2}$, $\operatorname{Geom}(x)$ gets the first real success, with probability $\frac{1}{2} \cdot p$, $\operatorname{Geom}(px)$ gets the first real success, and with probability $\frac{1}{2} (1 - p)$, we saw a fake success and will continue waiting for the next potential success.
+Clearly, then, $\operatorname{Geom}(x)$ will get the first success with probability $\frac{1}{2} / (\frac{1}{2} + \frac{1}{2}p)$.
+Thus, $\operatorname{Geom}(px)$ will get the first success with probability $p / (1 + p) = O(p)$, which is exactly what we wanted to show.
+
+        This concludes the proof that component 1 is such that with probability $1 - O(p)$, $s_1$ will have reached an absorbing strongly connected component before $s_2$ continues to component 2 of $K_H$.
+      \end{proof}
+
+      \begin{proof}[Proof of \cref{claimcansyncs2}]
+        This construction is relatively simple.
+It is a chain of states where each state transitions to the next on both input $C$ and $D$ (that is, deterministically).
+Let $w_H^*$ be the string used in the proof of \cref{claimcanwaitfors1} for component $H$.
+Let $W = \sum_{H} (d+1) | w_H^* |$.
+Then, the outputs of the first $W$ states form the string $w_H^*$ repeated $d+1$ times, for all $w_H$.
+The following $W+1$ states always output $D$.
+This is followed by one last state outputting $C$.
+The $C$ state has a transition into itself on input $D$, and out of this component (to component 3 of $K_H$) on input $C$.
+
+        Suppose that $s_2$ plays against a clone $s_2'$ of itself, and that $s_2$ enters an absorbing component $K_H$ first.
+Then, by \cref{claimcanwaitfors1}, using the fact that $s_2'$ is a copy of $s_1$ before it reaches an absorbing component, $s_2$ will not enter component 2 of $K_H$ until $s_2'$ enters an absorbing component $K_{H'}$, with probability $1 - O(p)$.
+When $s_2$ enters component 2, it will at some point output the entirety of the string $w_{H'}^*$.
+With probability $1 - O(p)$, $s_2'$ will then, by our construction of component 1, enter component 2.
+That is, $s_2'$ enters the first part of component 2 before $s_2$ leaves the first part of component 2, which means that $s_2$ is at most $W$ steps ahead of $s_2$.
+This means that $s_2'$ will be in the defection part of the chain when $s_2$ reaches the cooperation state.
+Assuming no mistakes, then, $s_2$ will wait until $s_2'$ also reaches the cooperation state, at which point they will both simultaneously leave the component.
+This happens with probability $1 - O(p)$, as desired.
+      \end{proof}
+
+      \begin{proof}[Proof of \cref{claimcanfigureout}]
+
+          Let $\mathcal{A}_1$ be the set of all absorbing strongly connected components in $s_1$.
+           Let $M$ be the set of all tuples of $(C, c)$ where $H \in \mathcal{A}_1$ and $c$ is a possible state of $C$.
 Let $M'$ be a subset of $M$ where a tuple $(C, c)$ has been removed if there is another tuple $(C', c') \in M'$ produces exactly the same output on all inputs, when only following the $1-p$ transitions.

### 2020-05-22 paper/paper.tex
almost done almost done

@@
 \usepackage{todonotes}
+\usepackage[hidelinks]{hyperref}
 \usepackage{cleveref}
@@
 \title{Cooperation Is Provably Required in a Version of the Noisy Iterated Prisoner's Dilemma}
-\author{Arvid Lunnemark. \\ \\
-Supervised by: Michael Sipser}
+\author{Arvid Lunnemark}
 
@@
 In \cref{sectionproofs}, we prove our results.
-Finally, in \cref{sectiondiscussion} we briefly discuss other potential ways of modeling the problem.
+Finally, in \cref{sectiondiscussion} we briefly discuss potential directions for future work.
 
@@
   \end{equation*}
-  where $p_{c_1 \to c_1'}$ is $1-p$ if the output of $s_2$ at state $c_2$ causes $c_1$ to transition to $c_1'$, and is $p$ if the opposite of that output causes that transition; $p_{c_2 \to c_2'}$ is defined similarly.
+  where $p_{c_1 \to c_1'}$ is 1 if $s_1$ always transitions from $c_1$ to $c_1'$, 0 if it never does, $1-p$ if the output of $s_2$ at state $c_2$ causes $s_1$ to transition from $c_1$ to $c_1'$, and $p$ if the opposite of that output causes the transition; $p_{c_2 \to c_2'}$ is defined similarly.
 \end{definition}
 
-Markov chains naturally leads themselves to the study of limiting cases.
+Markov chains naturally lead themselves to the study of limiting cases.
 
@@
   \label{timeaveragedistribution}
-  The \textit{time average distribution} of the $s_1$-$s_2$ Markov chain given the start state $(a, b)$, denoted $\pi^{(a, b)}$, is the distribution such that \begin{equation*}
-    \pi_{c_1, c_2}^{(a, b)} = E \left[ \text{fraction of time in state $(c_1,c_2)$} \mid \text{initial state is } (a, b)  \right]
+  Let $a_0$ be the start state of $s_1$ and $b_0$ the start state of $s_2$.
+The \textit{time average distribution} of the $s_1$-$s_2$ Markov chain, denoted $\pi$, is the distribution such that \begin{equation*}
+    \pi_{c_1, c_2} = E \left[ \text{fraction of time in state $(c_1,c_2)$} \mid \text{start in state } (a_0, b_0)  \right]
   \end{equation*}
-  where the expected fraction of time is taken over the infinite sequence $(X_0, X_1, \ldots)$.
+  where the expected fraction of time is taken over the infinite sequence of states representing a random walk.
 \end{definition}
 
-We will often use $\pi$ to refer to $\pi^{(a_0, b_0)}$ where $a_0$ is the initial state of $s_1$ and $b_0$ is the initial state of $s_2$.
+Note that the time average distribution given a start state always exists, regardless of the structure of the Markov chain. Now, we can define the payoff.
 
@@
 \begin{definition}
+  \label{definitionevolutionarystability}
   A strategy $s_1$ is \textit{evolutionarily stable} if there exists parameters $p_0$ and $\epsilon_0$ with $0 < p_0, \epsilon_0 < 1$, such that for all $p < p_0$, and all $\epsilon < \epsilon_0$, $s_1$ is not $\epsilon$-invadable.
@@
   Thus, $v_s(s)$ where $s$ is Tit-for-tat is much smaller than $R$, which means that it is not evolutionarily stable by \cref{evolutionarystable1}.
+Notably, this shows that our results are different, and perhaps stronger, than Fudenberg and Maskin's: their criterion for evolutionary stability did not rule out Tit-for-tat \cite{fundenberg1990evolution}.
 \end{remark}
@@
 
-Before we prove \cref{evolutionarystable1}, we need to understand what the payoff $v_{s_1}(s_2)$ really means.
+Before we prove \cref{evolutionarystable1}, we need to understand the mechanisms behind the payoff $v_{s_1}(s_2)$.
 In this subsection, we prove a series of lemmas that characterize the time average distribution, and consequently $v_{s_1}(s_2)$.
@@
   \label{lemmaabsorbinghasuniquetimeaverage}
-  An absorbing strongly connected component has a unique time average distribution, i.e., the time average distribution does not depend on the start state.
+  An absorbing strongly connected component has a unique time average distribution. That is, the time average distribution does not depend on the start state.
 \end{lemma}
@@
   \label{timeaveragedistributiondecomposition}
-  Let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain. For every $S \in \mathcal{A}$, let $\pi^{(S)}$ be its unique time average distribution. Then
+  Let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain.
+For every component $S \in \mathcal{A}$, let $\pi^{(S)}$ be its unique time average distribution.
+Then,
   \begin{equation*}
@@
   \end{equation*}
+  By $P(s_1\text{-}s_2 \text{ chain ends up in } S)$, we mean the probability of ending up in $S$ given that we start in the initial states of the two strategies and perform a random walk in the Markov chain.
 \end{lemma}
+
+\begin{proof}
+  Condition all of the following on starting in the start state of $s_1$ and $s_2$.
+Let $F$ be the fraction of time spent in state $(c_1,c_2)$.
+If $(c_1,c_2)$ is not in an absorbing strongly connected component, it is well known that the fraction of time in that state $F$ will be 0.
+This is also what the lemma says.
+Suppose now that $(c_1,c_2)$ is in an absorbing strongly connected component $S$.
+  Let $X$ be the event that the $s_1$-$s_2$ Markov chain ends up in $S$. By the law of total probability, \begin{equation*}
+    E[F] = E[F \mid X] \cdot P(X) + E[F \mid \text{not } X] \cdot P(\text{not } X)
+  \end{equation*}
+   If the random walk on the $s_1$-$s_2$ chain does not end up in the component $S$, but rather in another absorbing strongly connected component, the expected fraction of time in state $(c_1,c_2)$ is clearly 0.
+Also, note that $E[F \mid X]$ is exactly $\pi^{(S)}_{c_1,c_2}$, and that $E[F]$ is exactly $\pi_{c_1,c_2}$.
+We thus find that \begin{equation*}
+     \pi_{c_1,c_2} = P(\text{$s_1$-$s_2$ chain ends up in $S$}) \cdot \pi^{(S)}_{c_1,c_2}.
+   \end{equation*}
+   Note that $\pi^{(S')}_{c_1,c_2}$ is 0 for all other absorbing strongly connected components $S'$. With this, we have proved the lemma.
+\end{proof}
 
@@
 \end{lemma}
+
+\begin{proof}
+  First, we note that $H \times K$ for all $H \in \mathcal{A}_1$ and $K \in \mathcal{A}_2$ is an absorbing strongly connected component.
+If there is an edge from $(h,k)$ to $(u,v)$ in the $s_1$-$s_2$ Markov chain, there must, by definition, be an edge from $h$ to $u$ and $k$ to $v$, so $(u,v) \in H \times K$.
+It is also easy to see that if there is a path
+\end{proof}
 
@@
 By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ reaches an absorbing strongly connected component, which we call $H'$.
-        This means that the probability of reaching the absorbing component $H' \times K_H$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we hadn't replaced $H$ by $K_H$ in $s_2$, which is just the probability of reaching the absorbing component $H' \times H$ in the $s_1$-$s_1$ Markov chain.
+        This means that the probability of reaching the absorbing component $H' \times K_H$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$, the same as if we hadn't replaced $H$ by $K_H$ in $s_2$, which is just the probability of reaching the absorbing component $H' \times H$ in the $s_1$-$s_1$ Markov chain.
 Thus,
@@
           Let $\mathcal{A}_1$ be the set of all absorbing strongly connected components in $s_1$.
-           Let $M$ be the set of all tuples of $(C, c)$ where $H \in \mathcal{A}_1$ and $c$ is a possible state of $C$.
-Let $M'$ be a subset of $M$ where a tuple $(C, c)$ has been removed if there is another tuple $(C', c') \in M'$ produces exactly the same output on all inputs, when only following the $1-p$ transitions.
+           Let $L$ be the set of all tuples of $(H, h)$ where $H \in \mathcal{A}_1$ and $h$ is a possible state of $H$.
 
-        The procedure works as follows:
+        The third component is implemented as follows. When we talk about ``at compile time'' we refer to when we construct the finite automaton.
         \begin{enumerate}
-          \item Simulate $C_1$ for $N$ steps.
-(This can be done by e.g.
-duplicating $C_1$ $N$ times and having all edges advance to corresponding state in the next copy.) Choose $N$ so that after $N$ timesteps, $s_1$ will be in an absorbing strongly connected component with probability $1 - O(p)$.
-\todo{move this out of the claim???
-but it's annoying and needs a space somewhere}
-          \item Iterate over every pair of tuples $(C,c)$ and $(C', c')$ in $M$:
+          \item Iterate over every pair of tuples $(H,h)$ and $(H', h')$ in $L$:
           \begin{enumerate}
-            \item There is a string $S$ of actions on which $C$ starting in $c$ and $C'$ starting in $c'$ will produce a different output at a time $t$.
-Now, output the string $S$, and determine if the sequence of perceived actions corresponds to $(C,c)$ or $(C', c')$.
+            \item At compile time, we know what actions have been taken so far in this component, and thus know, if $s_1$ were in $H$ in state $h$ when $s_2$ started out this component, the state $g$ that $s_1$ is in now, with high probability.
+Similarly supposing that $s_1$ were in $H'$ in state $h'$ when $s_2$ started out this component, we know the state $g'$ where $s_1$ would be now.
+            \item At compile time, we also know whether there exists an input string $w$ such that $s_1$ starting in state $g$ perceiving $w$ produces a different output sequence than if it started in state $g'$, following only $1-p$ transitions.
+            \item If no such $w$ exists, continue the loop here.
+            \item If such a $w$ exists, then output $w$ deterministically here. Determine if the sequence of perceived actions corresponds to what would be expected from $(H,h)$ or from $(H', h')$.
           \end{enumerate}
         \end{enumerate}
-        Choose $T = \max t$ over all $t$.
-         With probability $1 - O(p)$, there will be exactly 1 pair $(C, c)$ that corresponds to all perceived actions in each of its comparisons.
-We know what $C$, when starting in $c$, outputs at time $T$.
-Thus, with probability $1 - O(p)$, we know what action $A$ that $s_1$ outputs at time $T$.
+        In the end, there will be a subset $U$ of $L$ that corresponds to all tuples $(H,h)$ that corresponded to the perceived sequence of actions in each comparison it was part of.
+By the way the loop is constructed, all $(H,h)$ in $U$ will produce the same high probability output on any input string; for our purposes, we can therefore just pick one canonical $(H,h)$ from $U$.
 
-        It should be noted that this procedure is easily implementable on a finite automaton, using e.g. a decision tree structure.
+        Note that when we implement this procedure on a finite automaton, it becomes a large decision tree, which means that in the end, there will be one branch for each canonical $(H, h)$, and we will be in one corresponding to where $s_1$ actually started off in the beginning of this component with probability $1 - O(p)$.
+In each such branch, we know, at compile time, which state $s_1$ is in at the end of the branch, and thus know what output that $s_1$ will have in the next step, with probability $1 - O(p)$.
+This is exactly what we wanted, so we transition out of component 3.
       \end{proof}
 
-      \todo{KEY IMPORTANT DETAIL: we need to say that at a large finite $T$ then $s_1$ will be within an absorbing strongly connected component.
+    %   \todo{KEY IMPORTANT DETAIL: we need to say that at a large finite $T$ then $s_1$ will be within an absorbing strongly connected component.
 probably need to look at all possible starting states of all possible ASCCs, not only every ASCC.
@@
 
-    \begin{proof}
-      \begin{claim}
-        Suppose that there are no mistakes, i.e., that $p = 0$. We can create a $C_2$ such that 
-      \end{claim}
+    % \begin{proof}
+    %   \begin{claim}
+    %     Suppose that there are no mistakes, i.e., that $p = 0$. We can create a $C_2$ such that 
+    %   \end{claim}
 
-      First, copy the entire $s_1$ machine into $s_2$.
+    %   First, copy the entire $s_1$ machine into $s_2$.
 Suppose that the state corresponding to the start state of $s_1$ is $c$, and that the output at $c$ is $\alpha$, and that the state $s$ goes to upon perceiving the opponent move $\alpha$ is $c' = T(c, \alpha)$.
@@
 Define the transitions as \begin{align*}
-        T(c_0, \alpha) &= c'\\
-        T(c_0, \bar{\alpha}) &= c_1 \\
-        T(c_1, \cdot) &= c_1
-      \end{align*}
-      and the outputs as \begin{align*}
-        G(c_0) &= \lnot G(c_2)\\
-        G(c_1) &= C.
-      \end{align*}
-      Let the start state of $s_2$ be $c_0$.
+    %     T(c_0, \alpha) &= c'\\
+    %     T(c_0, \bar{\alpha}) &= c_1 \\
+    %     T(c_1, \cdot) &= c_1
+    %   \end{align*}
+    %   and the outputs as \begin{align*}
+    %     G(c_0) &= \lnot G(c_2)\\
+    %     G(c_1) &= C.
+    %   \end{align*}
+    %   Let the start state of $s_2$ be $c_0$.
 
-      \begin{figure}
-        \centering
-        \begin{tikzpicture}[node distance=2cm]
-          \node[state, initial] (c_0) {$\bar{\alpha}$};
-          \node[state] (c_1) [below right of=c_0] {$C$};
-          \node[state] (c_prime) [above right of=c_0] {};
-          \node[state] (c) [above left of=c_prime] {$\alpha$};
+    %   \begin{figure}
+    %     \centering
+    %     \begin{tikzpicture}[node distance=2cm]
+    %       \node[state, initial] (c_0) {$\bar{\alpha}$};
+    %       \node[state] (c_1) [below right of=c_0] {$C$};
+    %       \node[state] (c_prime) [above right of=c_0] {};
+    %       \node[state] (c) [above left of=c_prime] {$\alpha$};
 
-          \path[->] (c_0) edge node [above right] {$\bar{\alpha}$} (c_1);
-          \path[->] (c_1) edge [loop right] node [right] {$C,D$} ();
-          \path[->] (c_0) edge node [above left] {$\alpha$} (c_prime);
-          \path[->] (c) edge node [below left] {$\alpha$} (c_prime);
-        \end{tikzpicture}
-        \caption{Constructon of invasion strategy, used in the proof of \cref{evolutionarystable1}}
-        \label{invasionstrategy}
-      \end{figure}
+    %       \path[->] (c_0) edge node [above right] {$\bar{\alpha}$} (c_1);
+    %       \path[->] (c_1) edge [loop right] node [right] {$C,D$} ();
+    %       \path[->] (c_0) edge node [above left] {$\alpha$} (c_prime);
+    %       \path[->] (c) edge node [below left] {$\alpha$} (c_prime);
+    %     \end{tikzpicture}
+    %     \caption{Constructon of invasion strategy, used in the proof of \cref{evolutionarystable1}}
+    %     \label{invasionstrategy}
+    %   \end{figure}
 
-      \begin{claim}
-        Given the above construction of $s_2$, the following inequalities hold:
-      \begin{align*}
-        v_{s_1}(s_1) &\leq (1-p)^2 \gamma + 2(1-p)p R + p^2  R \\ 
-        v_{s_1}(s_2) &\leq (1-p) \gamma + p T \\
-        v_{s_2}(s_1) &\geq (1 - p) \gamma  + p S \\
-        v_{s_2}(s_2) &\geq (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma .
-      \end{align*}
-      \end{claim}
+    %   \begin{claim}
+    %     Given the above construction of $s_2$, the following inequalities hold:
+    %   \begin{align*}
+    %     v_{s_1}(s_1) &\leq (1-p)^2 \gamma + 2(1-p)p R + p^2  R \\ 
+    %     v_{s_1}(s_2) &\leq (1-p) \gamma + p T \\
+    %     v_{s_2}(s_1) &\geq (1 - p) \gamma  + p S \\
+    %     v_{s_2}(s_2) &\geq (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma .
+    %   \end{align*}
+    %   \end{claim}
 
-      Before proving this claim, we will use it to finish our proof of \cref{evolutionarystable1}.
+    %   Before proving this claim, we will use it to finish our proof of \cref{evolutionarystable1}.
 
-      Now, we simply compute $F(s_2) - F(s_1)$, which we want to show is greater than 0.
-      \begin{align*}
-        F(s_2) - F(s_1) &= \\
-        &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
-        &= (1 - \epsilon) (\gamma + p(\ldots)) + \epsilon (R + p(\ldots)) - (1-\epsilon) (\gamma + p(\ldots)) - \epsilon (\gamma + p(\ldots)) \\
-        &= \epsilon (R - \gamma) + p(\ldots)
-      \end{align*}
+    %   Now, we simply compute $F(s_2) - F(s_1)$, which we want to show is greater than 0.
+    %   \begin{align*}
+    %     F(s_2) - F(s_1) &= \\
+    %     &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
+    %     &= (1 - \epsilon) (\gamma + p(\ldots)) + \epsilon (R + p(\ldots)) - (1-\epsilon) (\gamma + p(\ldots)) - \epsilon (\gamma + p(\ldots)) \\
+    %     &= \epsilon (R - \gamma) + p(\ldots)
+    %   \end{align*}
 
-      We know that $R - \gamma > 0$ by our initial assumption.
+    %   We know that $R - \gamma > 0$ by our initial assumption.
 Clearly, since $(\ldots)$ is some polynomial in $p$, given an $\epsilon$ we can find a sufficiently small $p$ such that the full expression is positive.
@@
 
-    \end{proof}
+    % \end{proof}
 
-    \begin{proof}[Proof of \cref{claimpayoffs}]
-      We can prove this using either of the two definitions.
-    \end{proof}
+    % \begin{proof}[Proof of \cref{claimpayoffs}]
+    %   We can prove this using either of the two definitions.
+    % \end{proof}
 
@@
 
-    We note that the $2R > T + P$ condition is necessary.
-Otherwise, the AllD strategy would be able to invade Pavlov.
-We see this by noting that $\lim_{p \to 0} v_{s_2}(s_1) = T + P$ if $s_2$ is AllD and $s_1$ is Pavlov, and that AllD is better against itself than Pavlov is against it.
+    We can note that the $2R > T + P$ condition is necessary.
+Otherwise, the All-D strategy, depicted in \cref{figurealld}, would be able to invade Pavlov.
+We see this by noting that $\lim_{p \to 0} v_{s_2}(s_1) = T + P$ if $s_2$ is All-D and $s_1$ is Pavlov, and that All-D is better against itself than Pavlov is against it.
 
-    \section{Discussion of Model}
+    We have some, admittedly weak, empirical evidence for this theorem to be true.
+First, Pavlov when played against itself gets payoff $R - O(p)$, and our proof of \cref{evolutionarystable1} is very clearly relying heavily on letting $p$ be very close to 0, so it seems unlikely that a similar proof would be able to prove that Pavlov is not evolutionarily stable.
+Moreover, Fudenberg and Maskin stated but did not prove a similar result in their version of the noisy regime \cite{fundenberg1990evolution}.
+We also wrote a Julia notebook for generating $s_1$-$s_2$ Markov chains and the time average distributions, and ran Pavlov against a few variants of Pavlov and other simple strategies like All-D, All-C and Tit-for-tat, and in all cases Pavlov performed better than our potential invaders; this can be found in the accompanying GitHub repository \cite{arvid2020}.
+That repository also contains a document of unorganized notes that may provide hints for proving the conjecture.
+
+\begin{figure}
+\begin{tikzpicture}[node distance=2cm]
+  \node[state,initial] (s_0) {$D$};
+
+  \path[->] (s_0) edge [loop above] node {$C,D$} ();
+\end{tikzpicture}
+  \centering
+  \caption{The always defecting All-D strategy.}
+  \label{figurealld}
+\end{figure}
+
+    \section{Further Work}
     \label{sectiondiscussion}
 
-    \subsection{Other Potential Models}
+    An obvious next step is to prove \cref{pavlovtheorem}. We believe that it should not be too hard.
 
-    Right now we have only modeled noise in perception.
-One could think of another possible kind of noise: a ``failure of the mind,'' which perhaps could be modeled instead by a probability $p$ of being transported to any random state, instead.
-This would create ergodicity which is nice.
+    One may also consider different setups.
+For example, weakening \cref{definitionevolutionarystability} of evolutionary stability by saying that it is only required to not be $\epsilon$-invadable for a fixed $p$, instead of for infinitesimally small $p$, would lead to a more interpretable model.
+Showing that \cref{evolutionarystable1} still holds in this situation would be interesting.
 
-    \iffalse
+    One can also think of other kinds of noise: for example, a ``failure of the mind,'' which could be modeled by a probability $p$ of being transported to any random state in the automaton.
+This would automatically make the Markov chain strongly connected, which can potentially simplify a lot of the analysis.
 
-    \section{Appendix: Time average distributions}
+    % \section{Appendix: Time average distributions}
 
-    We might have periodicity, but for our purposes, we might as well extend the definition and look at periodic distributions as stationary too. The following two lemmas help with that.
+    % We might have periodicity, but for our purposes, we might as well extend the definition and look at periodic distributions as stationary too. The following two lemmas help with that.
 
-    \begin{lemma}
-      Given a starting distribution $v$ and a Markov matrix $M$, for every $\epsilon > 0$, there will exist a $k$ such that $|vP^{nk} - vP^{mk}| < \epsilon$ for all $n$ and $m$ $> 0$.
-    \end{lemma}
+    % \begin{lemma}
+    %   Given a starting distribution $v$ and a Markov matrix $M$, for every $\epsilon > 0$, there will exist a $k$ such that $|vP^{nk} - vP^{mk}| < \epsilon$ for all $n$ and $m$ $> 0$.
+    % \end{lemma}
 
-    This proves that a Markov chain will always reach a periodic state.
+    % This proves that a Markov chain will always reach a periodic state.
 
-    \begin{lemma}
-      Suppose distributions form a chain $p_1 \to p_2 \to \cdots \to p_n \to p_1$. Then $\pi = \frac{p_1 + \ldots p_n}{n}$ is stationary.
-    \end{lemma}
+    % \begin{lemma}
+    %   Suppose distributions form a chain $p_1 \to p_2 \to \cdots \to p_n \to p_1$. Then $\pi = \frac{p_1 + \ldots p_n}{n}$ is stationary.
+    % \end{lemma}
 
-    This proves that we're able to talk about stationary distributions even when they don't really actually exist.
+    % This proves that we're able to talk about stationary distributions even when they don't really actually exist.
 
-    \fi
+    % \section{Appendix: Probablistic Automata}
 
-    \section{Appendix: Probablistic Automata}
+    % OHHHHHHH. THIS WILL HELP WITH MY PROOF????? THIS IS EXACTLY WHAT I'M TRYING TO DO IN MY PROOF RIGHT????? YEAHHHH, NO.
 
-    OHHHHHHH. THIS WILL HELP WITH MY PROOF????? THIS IS EXACTLY WHAT I'M TRYING TO DO IN MY PROOF RIGHT????? YEAHHHH, NO.
-
-    In this paper, we have considered strategies that make a deterministic move based on what they perceive.
+    % In this paper, we have considered strategies that make a deterministic move based on what they perceive.
 One could also imagine strategies that attaches a certain probability distribution to a perceived input, and chooses their next action based on that.
@@
 
-    Proof idea: we can use cycles in the Markov chain with n total outputs, x of which are to state 1, to model getting to state 1 with probability x / n.
+    % Proof idea: we can use cycles in the Markov chain with n total outputs, x of which are to state 1, to model getting to state 1 with probability x / n.
 this assumes that the outputs are of low enough probability, which can be achieved by chaining together lots of (1-p) transitions, which go to 0.
 
-    the hard part of this is showing that modifying the finite automaton like this won't hurt us.
+    % the hard part of this is showing that modifying the finite automaton like this won't hurt us.
 in fact, it would certainly not hurt us if not every state had to give an output.

### 2020-05-22 paper/paper.tex
finish!!!! all proofs are done but probably quite bad

@@
 
-Finally, we can characterize the absorbing strongly connected components of the $s_1$-$s_2$ Markov chain in terms of the absorbing SCCs of $s_1$ and $s_2$ separately.
+% Finally, we can characterize the absorbing strongly connected components of the $s_1$-$s_2$ Markov chain in terms of the absorbing SCCs of $s_1$ and $s_2$ separately.
 
-\begin{lemma}
-  \label{absorbingstronglyconnectedcomponentscartesianproduct}
-  Let $\mathcal{A}_1$ be the set of absorbing strongly connected components of $s_1$, and similarly, let $\mathcal{A}_2$ be the set of absorbing strongly connected components of $s_2$.
+% \begin{lemma}
+%   \label{absorbingstronglyconnectedcomponentscartesianproduct}
+%   Let $\mathcal{A}_1$ be the set of absorbing strongly connected components of $s_1$, and similarly, let $\mathcal{A}_2$ be the set of absorbing strongly connected components of $s_2$.
 If $\mathcal{A}$ is the set of absorbing strongly connected components of the $s_1$-$s_2$ Markov chain, then
-  \begin{equation*}
-    \mathcal{A} = \{H \times K \mid H \in \mathcal{A}_1 \text{ and } K \in \mathcal{A}_2\},
-  \end{equation*}
-  where $A \times B$ is the cartesian product $A \times B = \{(a, b) \mid a \in A \text{ and } b \in B\}$.
-\end{lemma}
+%   \begin{equation*}
+%     \mathcal{A} = \{H \times K \mid H \in \mathcal{A}_1 \text{ and } K \in \mathcal{A}_2\},
+%   \end{equation*}
+%   where $A \times B$ is the cartesian product $A \times B = \{(a, b) \mid a \in A \text{ and } b \in B\}$.
+% \end{lemma}
 
-\begin{proof}
-  First, we note that $H \times K$ for all $H \in \mathcal{A}_1$ and $K \in \mathcal{A}_2$ is an absorbing strongly connected component.
+% \begin{proof}
+%   First, we note that $H \times K$ for all $H \in \mathcal{A}_1$ and $K \in \mathcal{A}_2$ is an absorbing strongly connected component.
 If there is an edge from $(h,k)$ to $(u,v)$ in the $s_1$-$s_2$ Markov chain, there must, by definition, be an edge from $h$ to $u$ and $k$ to $v$, so $(u,v) \in H \times K$.
 It is also easy to see that if there is a path
-\end{proof}
+% \end{proof}
 
@@
       To do that, we want to prove that for all $p_0, \epsilon_0 \in (0,1)$, there exists $p < p_0$ and $\epsilon < \epsilon_0$, such that $s_1$ is $\epsilon$-invadable.
-We choose $\epsilon = \epsilon_0 / 2$, and present a strategy $s_2$ that can invade $s_1$ for all $p$ that are sufficiently small.
+We present a strategy $s_2$ that can invade $s_1$ for all $p$ and $\epsilon$ that are sufficiently small.
 
@@
         until $s_1$ has reached an absorbing strongly connected component. 
-       Once $s_1$ is in an absorbing strongly connected component, \cref{lemmaabsorbinghasuniquetimeaverage} tells us that the time average distribution does not depend on the particular start state in that component, which means that even if $s_1$ after this point ``figures out'' that it is not playing against itself anymore, there is nothing it can do about it (since it also cannot escape the component).
+       Once $s_1$ is in an absorbing strongly connected component, it cannot escape, and $s_2$ can also force it to end up in any of the states in the component, which means that even if $s_1$ after this point ``figures out'' that it is not playing against itself anymore, there's nothing it can do about it.
       The second step in the construction of $K_H$ makes sure that if $s_2$ is playing against itself, it is in sync with its clone; that is, with probability $1 - O(p)$, both copies of $s_2$ will transition to the third component at the exact same time (although not necessarily in the same $K_H$).
@@
       Then, $K_H$ outputs the exact opposite action at time $T$. 
-      At that point, if its opponent outputs $s_1$'s expected action, it transitions into the absorbing strongly connected component $H_\text{copy}$, which is just an exact copy of $H$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and enters an always cooperating state.
+      At that point, if its opponent outputs $s_1$'s expected action, it wants to transition into the absorbing strongly connected component $H_\text{copy}$, which is just an exact copy of $H$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and enters an always cooperating state.
+      Before transitioning into $H_\text{copy}$, $s_2$ makes sure that $s_1$ is in a state that is beneficial for $s_2$.
 
@@
           \node[state] (c_6) [below=of c_5] {$C$};
-          \node[elliptic state] (c_7) [below=of c_4] {$H_\text{copy}$ \\(an exact copy of $H$)};
+          \node[elliptic state] (c_7) [below=of c_4] {4. Force $s_1$ into a good state.};
+          \node[elliptic state] (c_8) [below=of c_7] {$H_\text{copy}$ \\(an exact copy of $H$)};
 
@@
           \path[->] (c_6) edge [loop right] (c_6);
+          \path[->] (c_7) edge (c_8);
           % \path[->] (c_0) edge node [above right] {$\bar{\alpha}$} (c_1);
@@
         \caption{High level construction of $K_H$, the absorbing component replacing $H$ when creating $s_2$ in the proof of \cref{evolutionarystable1}.
-Note that $K_H$ contains exactly 2 absorbing strongly connected components.}
+Note that $K_H$ contains exactly 2 absorbing strongly connected components.
+SCC means strongly connected component.}
         \label{c2flowchart}
@@
 We now want to prove that this $s_2$ can in fact invade $s_1$.
-To do that, we will first assume that the below 3 lemmas are correct, i.e., that the construction is possible and works as specified, and then finish the proof of our theorem using those assumptions.
-Once that's done, we will prove the 3 lemmas.
+To do that, we will first assume that the below four lemmas are correct, i.e., that the construction is possible and works as specified, and then finish the proof of our theorem using those assumptions.
+Once that's done, we will prove the four lemmas.
 
@@
 
+      \begin{lemma}
+        \label{claimcanforcegoodstate}
+        We can construct a non-absorbing component placed after the third component that, supposing that $s_2$ plays against $s_1$, can transition into any given absorbing strongly connected component $S$ in the $s_1$-$s_2$ chain that is a subset of $H_\text{copy} \times H'$ where $H'$ is the absorbing strongly connected component that $s_1$ is in.
+      \end{lemma}
+
       We now finish our proof, starting with a lemma concerning the payoffs.
@@
         v_{s_1}(s_1) &= \gamma \pm O(p) \\
-        v_{s_1}(s_2) &= \gamma \pm O(p) \\
-        v_{s_2}(s_1) &= \gamma \pm O(p) \\
-        v_{s_2}(s_2) &= R \pm O(p)
+        v_{s_2}(s_2) &= R \pm O(p),
+      \end{align*}
+      and either
+      \begin{align*}
+        \lim_{p \to 0} v_{s_2}(s_1) > \gamma,
+      \end{align*}
+      or
+      \begin{align*}
+        v_{s_1}(s_2) = \gamma \pm O(p) \text{ and }
+        v_{s_2}(s_1) = \gamma \pm O(p)
       \end{align*}
@@
       \begin{subproof}
+        By our definition of $\gamma$, we have that $v_{s_1}(s_1) \leq \gamma \pm O(p)$.
 
-        By our definition of $\gamma$, we have that $v_{s_1}(s_1) = \gamma \pm O(p)$.
+        Consider now the $s_2$-$s_2$ Markov chain.
+By \cref{claimcansyncs2}, both clones will transition into the third component at the exact same time.
+The third component spends only a finite amount of time by \cref{claimcanfigureout}, and is the same regardless of which $K_H$ the strategies are in.
+Thus, after leaving the third component of $K_H$, if no mistakes have happened since leaving the second component, both clones of $s_2$ will be in exactly corresponding states, since they will always mirror each other.
+This happens with probability at least $(1-p)^N$ where $N$ is the finite number of states in the third component, and thus with probability $1 - O(p)$.
+Thus, in the identifying stage, both clones will output the same action, and as we can see in \cref{c2flowchart}, they will then both transition to the always cooperating state with probability $1 - O(p)$.
+Therefore, $v_{s_2}(s_2) = R \pm O(p)$.
 
@@
 By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ reaches an absorbing strongly connected component, which we call $H'$.
-        This means that the probability of reaching the absorbing component $H' \times K_H$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$, the same as if we hadn't replaced $H$ by $K_H$ in $s_2$, which is just the probability of reaching the absorbing component $H' \times H$ in the $s_1$-$s_1$ Markov chain.
+        This means that the probability of reaching the absorbing component $H' \times K_H$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we had not replaced $H$ by $K_H$ in $s_2$, which is just the probability of reaching the absorbing component $H' \times H$ in the $s_1$-$s_1$ Markov chain.
 Thus,
@@
         \end{equation}
-        Let $C_H$ be the always cooperating absorbing state of $K_H$. Then, this equation implies that 
+        % Let $C_H$ be the always cooperating absorbing state of $K_H$. Then, this equation implies that 
+        % \begin{equation}
+        %   \label{equationhch}
+        %   P(\text{$s_1$-$s_2$ ends up in $H' \times C_H$}) = O(p).
+        % \end{equation}
+        By \cref{claimcanforcegoodstate}, $s_1$ can choose to end up in any absorbing strongly connected component $S_{H', H}$ that is a subset of $H' \times H_\text{copy}$, with probability $1 - O(p)$.
+        Naturally, $s_1$ chooses an $S_{H', H}^*$ as follows: among the $S$ that have the maximum payoff for $s_2$, not caring about $O(p)$, we choose the one the minimizes the payoff for $s_1$.
+        Let $v_{s}(S)$ denote the payoff that strategy $s$ achieves in the absorbing strongly connected component $S$. Then,
         \begin{equation}
-          \label{equationhch}
-          P(\text{$s_1$-$s_2$ ends up in $H' \times C_H$}) = O(p).
+          \label{imtoootired}
+          v_{s_2}(S_{H', H}^*) \geq \sum_{S} P(\text{$s_1$-$s_1$ ends up in $S$} \mid \text{$s_1$-$s_1$ ends up in $H' \times H$}) \cdot v_{s_1}(S) \pm O(p)
         \end{equation}
-        Thus, when $s_1$ plays against $s_2$, the absorbing strongly connected components are mostly the same, with mostly the same probabilities. 
-        Now, let $\pi$ be the time average distribution of the $s_1$-$s_1$ Markov chain and let $\tau$ be the time average distribution of the $s_1$-$s_2$ Markov chain.
-Let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_1$ chain, and let $\mathcal{B}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ chain.
-Using \cref{timeaveragedistributiondecomposition} and \cref{absorbingstronglyconnectedcomponentscartesianproduct}, we can now calculate $\tau$:
+        since the right hand side is a weighted average of $v_{s_1}(S)$, which is the same as $v_{s_2}(S)$ in this case, which is what we maximized over when we chose $S_{H', H}^*$.
+Also, since $s_1$ could make sure to get into it with probability $1 - O(p)$, we have
+        \begin{equation}
+          \label{eqshh}
+          P(\text{$s_1$-$s_2$ ends up in $S_{H',H}^*$}) = P(\text{$s_1$-$s_2$ ends up in $H' \times H_\text{copy}$})  \pm O(p).
+        \end{equation}
+        Also, as a consequence, the $S_{H,H'}^*$ for all $H$ and $H'$ become the only absorbing strongly connected components with probabilities of ending up in them greater than $O(p)$.
+
+        Now, let $\mathcal{A}$ be the set of absorbing strongly connected components of the $s_1$-$s_1$ chain, and let $\mathcal{B}$ be the set of absorbing strongly connected components of the $s_1$-$s_2$ chain.
+We can now calculate $v_{s_2}(s_1)$.
 
@@
           \begin{WithArrows}
-          \tau &= \sum_{S \in \mathcal{B}} P(\text{$s_1$-$s_2$ ends up in $S$}) \cdot \pi^{(S)} \Arrow{\cref{absorbingstronglyconnectedcomponentscartesianproduct}}\\
-          &= \sum_{H', C_H} P(\text{$s_1$-$s_2$ ends up in $H' \times C_H$}) \cdot \pi^{(H' \times C_H)}\\ &\:\:\:\:\:\:\:+ \sum_{H', H_\text{copy}} P(\text{$s_1$-$s_2$ ends up in $H' \times H_\text{copy}$}) \cdot \pi^{(H' \times H_\text{copy})} \Arrow{\cref{equationhch}}\\
-          &= \sum_{H', H_\text{copy}} P(\text{$s_1$-$s_2$ ends up in $H' \times H_\text{copy}$}) \cdot \pi^{(H' \times H_\text{copy})} \pm O(p) \Arrow{\cref{equationhhcopy}}\\
-          &= \sum_{H', H} P(\text{$s_1$-$s_1$ ends up in $H' \times H$}) \cdot \pi^{(H' \times H)} \pm O(p) \Arrow{\cref{absorbingstronglyconnectedcomponentscartesianproduct}}\\
-          &= \sum_{S \in \mathcal{A}} P(\text{$s_1$-$s_1$ ends up in $S$}) \cdot \pi^{(S)} \pm O(p)\\
-          &= \pi \pm O(p)
+          v_{s_2}(s_1) &= \sum_{S \in \mathcal{B}} P(\text{$s_1$-$s_2$ ends up in $S$}) \cdot v_{s_2}(S) \\
+          &= \sum_{S_{H',H}^*} P(\text{$s_1$-$s_2$ ends up in $S_{H', H}^*$}) \cdot v_{s_2}(S_{H', H}^*) \pm O(p) \Arrow{\cref{eqshh}} \\
+          &= \sum_{S_{H',H}^*} P(\text{$s_1$-$s_2$ ends up in $H' \times H_\text{copy}$}) \cdot v_{s_2}(S_{H', H}^*) \pm O(p) \Arrow{\cref{equationhhcopy}} \\
+          &= \sum_{S_{H',H}^*} P(\text{$s_1$-$s_1$ ends up in $H' \times H$}) \cdot v_{s_2}(S_{H', H}^*) \pm O(p) \Arrow{\cref{imtoootired}}\\
+          &\geq \sum_{S \in \mathcal{A}} P(\text{$s_1$-$s_1$ ends up in $S$}) \cdot v_{s_1}(S) \pm O(p)\\
+          &= v_{s_1}(s_1) \pm O(p)
           \end{WithArrows}
         \end{align*}
-        Thus, $s_1$-$s_1$ and $s_1$-$s_2$ have almost the same time average distribution. This proves that $v_{s_1}(s_2) = \gamma \pm O(p)$ and $v_{s_2}(s_1) = \gamma \pm O(p)$.
+        Thus, we have that $v_{s_2}(s_1) \geq \gamma \pm O(p)$.
+Suppose that $v_{s_2}(s_1) = \gamma \pm O(p)$.
+Then, we must have equality in \cref{imtoootired} for all $H$ and $H'$, which means that all $S$ were equally good for $s_2$, in which case we chose $S_{H',H}^*$ to minimize the payoff for $s_1$.
+Following through with the exact same argument as above, but flipping the inequalities, we then find that $v_{s_1}(s_2) \leq \gamma \pm O(p)$.
 
-        For the last part of the claim, consider the $s_2$-$s_2$ Markov chain.
-By \cref{claimcansyncs2}, both clones will transition enter the third component at the exact same time.
-The third step takes only a finite number of steps by \cref{claimcanfigureout}, and is the same regardless of which $K_H$ the strategies are in.
-Thus, after leaving the third step of $K_H$, if no mistakes have happened since leaving state 2, both clones of $s_2$ will be in exactly corresponding states, since they will always mirror each other.
-This happens with probability at least $(1-p)^N$ where $N$ is the finite number of states in step 3, and thus with probability $1 - O(p)$.
-Thus, in the identifying stage, both clones will output the same thing, and as we can see in \cref{c2flowchart}, they will then both transition to the always cooperating state with probability $1 - O(p)$.
-Therefore, $v_{s_2}(s_2) = R \pm O(p)$, which finishes the proof of the lemma.
       \end{subproof}
@@
       Given \cref{claimpayoffs}, we simply compute $F(s_2) - F(s_1)$, which we want to show is greater than 0.
+We have two cases.
+First, we assume that $\lim_{p \to 0} v_{s_2}(s_1) > \gamma$.
+Let $\lim_{p \to 0} v_{s_2}(s_1) = \beta$.
+Then,
       \begin{align*}
@@
         &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
-        &= (1 - \epsilon) (\gamma \pm O(p)) + \epsilon (R \pm O(p)) - (1-\epsilon) (\gamma \pm O(p)) - \epsilon (\gamma \pm O(p)) \\
+        &= (1 - \epsilon) (\beta \pm O(p)) + \epsilon (R \pm O(p)) - (1-\epsilon) (\gamma \pm O(p)) - \epsilon (v_{s_1}(s_2) \pm O(p)) \\
+        &\geq (1 - \epsilon) (\beta - \gamma) + \epsilon(R - v_{s_1}(s_2)) \pm O(p)
+      \end{align*}
+      Here, we can make $\epsilon$ small enough such that this becomes positive, since $\beta - \gamma > 0$.
+Now, in the second case, we assume that $v_{s_2}(s_1) = \gamma \pm O(p)$ and $v_{s_1}(s_2) \leq \gamma \pm O(p)$.
+Then,
+      \begin{align*}
+        F(s_2) - F(s_1) &= \\
+        &= (1 - \epsilon) \cdot v_{s_2}(s_1) + \epsilon \cdot v_{s_2}(s_2) - (1 - \epsilon) \cdot v_{s_1}(s_1) - \epsilon \cdot v_{s_1}(s_2) \\
+        &\geq (1 - \epsilon) (\gamma \pm O(p)) + \epsilon (R \pm O(p)) - (1-\epsilon) (\gamma \pm O(p)) - \epsilon (\gamma \pm O(p)) \\
         &\geq \epsilon (R - \gamma) \pm O(p)
       \end{align*}
+      We know that $R - \gamma > 0$ by our initial assumption. Here, we can make $p$ small enough such that this becomes positive.
 
-      We know that $R - \gamma > 0$ by our initial assumption.
-For all sufficiently small $p$, thus, $F(s_2) - F(s_1) > 0$.
-This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is not $\epsilon$-invadable some value of $p < p_0$.
+      In conclusion, our constructed $s_2$ has higher fitness than $s_1$.
+This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is not $\epsilon$-invadable some value of $p < p_0$ and some $\epsilon < \epsilon_0$.
 In conclusion, then, $s_1$ is not evolutionarily stable, which concludes the proof of \cref{evolutionarystable1}.
@@
 This is exactly what we wanted, so we transition out of component 3.
+      \end{proof}
+
+      \begin{proof}[Proof of \cref{claimcanforcegoodstate}]
+        Note that an absorbing strongly connected component $S$ of the $s_1$-$s_2$ chain that is a subset of $H_\text{copy} \times H'$ can be gotten to by making $s_1$ transition to some state $h_1$ and $s_2$ transition to some state $h_2 \in H_\text{copy}$ where $(h_1, h_2) \in S$.
+This is simple.
+Note that $s_1$ is in a strongly connected component already, and that $s_2$ knows, by the previous component, which state $s_1$ is in.
+Thus, $s_2$ knows the path for $s_1$ to get to $h_1$, and can just output that.
+Then, $s_1$ will be in $h_1$, and $s_2$ can simply transition to $h_2$.
+This proves the lemma.
       \end{proof}

### 2020-05-22 paper/paper.tex
proofread a lil

@@
-\documentclass[11pt]{amsart}
+\documentclass[12pt]{article}
+\usepackage[letterpaper, margin=1.3in]{geometry}
 \usepackage{amsmath, amssymb, amsthm}
@@
 \usepackage{blkarray}
+\usepackage{tabularx}
 \usetikzlibrary{backgrounds,arrows,automata,positioning,shapes,shapes.geometric,calc, arrows.meta, bending, decorations.pathreplacing}
+
+    \newcolumntype{L}{>{\raggedright\arraybackslash}X}
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
 
@@
 
-\title{Cooperation Is Provably Required in a Version of the Noisy Iterated Prisoner's Dilemma}
-\author{Arvid Lunnemark}
+\title{\Large \textbf{Cooperation Is Provably Required for Success in a Version of the Noisy Iterated Prisoner's Dilemma}}
+\author{Arvid Lunnemark \\
+\url{arvid@mit.edu}\\ \\
+Supervised by: Prof. Michael Sipser}
 
@@
 
+\begin{abstract}
+    We introduce a natural model for the iterated prisoner's dilemma in the presence of noise, where strategies are finite automata that at each step have a probability $p$ of making a mistake.
+This leads us to model the result of two strategies playing against each other as a Markov chain.
+We prove that for a strategy to be evolutionarily stable in our setup, it has to be cooperative, in the sense that it has to cooperate when playing against a clone of itself.
+We conjecture that the Pavlov strategy is evolutionarily stable.
+\end{abstract}
+
 \section{Introduction}
 
-The prisoner's dilemma is a classic symmetric two-player game, where both players acting in their own best interests leads to an outcome that is not optimal for either of them.
-There are two actions, cooperate ($C$) and defect ($D$), with the payoffs represented by the matrix
-\begin{equation*}
-  \begin{matrix}
-      &  C & D \\ 
-    C &  (R, R) & (S, T) \\
-    D & (T, S) & (P, P)
-  \end{matrix}
-  % \begin{blockarray}{cccccc}
-  %   a & b & c & d & e \\
-  %   \begin{block}{c(ccccc)}
-  %     1 & 1 & 1 & 1 & 1 & f \\
-  %     0 & 1 & 0 & 0 & 1 & g \\
-  %     0 & 0 & 1 & 0 & 1 & h \\
-  %     0 & 0 & 0 & 1 & 1 & i \\
-  %     0 & 0 & 0 & 0 & 1 & j \\
-  %   \end{block}
-  %   \end{blockarray}
-\end{equation*} \todo{copy the table from axelrod's book because it's much better and much clearer!!!}
-where $T > R > P > S$ (the row player gets the first in each pair of numbers).
-Given the other player's action, a player always maximizes their payoff by choosing $D$, but with both doing so, the players end up in the $(D, D)$ state, which is worse for both than the $(C, C)$ state since $R > P$.
-The game can been used to model many real-world situations, including countries failing to act to stop climate change, doping in sport, and economic competition.
+The prisoner's dilemma is a classic symmetric two-player game, where both players acting in their own best interests leads them to an outcome that is not optimal for either of them.
+There are two actions: cooperate ($C$) and defect ($D$).
+The payoffs are awarded according to \cref{fig:prisonersdilemmatable}.
 
-To make the game more interesting, one can consider playing it multiple times in a row with multiple players.
-One might loosely connect the repeated multiplayer game to evolution: why have humans evolved to cooperate with each other?
-In a seminal paper in 1980, Axelrod presented evidence that in a repeated setting with multiple players facing off in a tournament, cooperation can arise as the strategy of choice even for a selfish player \cite{axelrod1980effective}.
+\begin{figure}
+    \centering
+    \begin{tabularx}{0.8\textwidth}{|c | Y| Y|}
+    \cline{2-3}
+    \multicolumn{1}{c|}{}
+         & Cooperate ($C$) & Defect ($D$) \\ \hline
+        Cooperate ($C$) & $R=3 , R=3$ & $S=0, T=5$ \\
+         & Reward for mutual cooperation & Sucker's payoff, temptation to defect \\ \hline
+        Defect ($D$) & $T=5, S=0$ & $P=1, P=1$ \\
+        & Temptation to defect, sucker's payoff & Punishment for mutual defection \\ \hline
+    \end{tabularx}
+    \caption{The prisoner's dilemma. The payoff of the player picking the row is listed first.}
+    \label{fig:prisonersdilemmatable}
+\end{figure}
+
+Given the other player's action, a player always maximizes their payoff by choosing to defect, but with both doing so, the players end up in the $(D, D)$ state, which is worse for both than the $(C, C)$ state.
+The game has been used to model many real-world situations, including countries failing to act to stop climate change, doping in sport, and economic competition.
+
+To make the game more interesting, one can consider playing it multiple times in a row, with multiple players.
+One might loosely connect this repeated multiplayer game to evolution: why have humans evolved to cooperate with each other?
+In a seminal paper in 1980, Axelrod presented evidence that in a repeated setting with multiple players facing off in a tournament, cooperation can arise as the strategy of choice, even for a selfish player \cite{axelrod1980effective}.
 In particular, the mostly cooperating Tit-for-tat strategy, depicted in \cref{tftfigure}, won his tournament.
-The idea is that even though Tit-for-tat loses when played against defecting strategies, that is outweighed by it being heavily rewarded when cooperating with other cooperating strategies.
+The idea is that even though Tit-for-tat loses when it plays against defecting strategies, it is being heavily rewarded when it cooperates with other cooperating strategies, giving Tit-for-tat a net advantage.
 
@@
 If all strategies had been of the type to always defect, Tit-for-tat would not have won the Axelrod tournament.
-Therefore, follow-up tournaments and simulations have been run, and perhaps surprisingly, most provide further evidence that cooperation is the prevailing strategy, as Axelrod summarizes in a 1981 book \cite{axelrod1981evolution}.
-There, he also introduces an evolutionary model that can be used for analyzing the game theoretically: strategies are considered to exist in a population that evolves over time in an evolutionary way (more successful strategies reproduce, and mutations can happen).
+Therefore, follow-up tournaments and simulations have been run, and perhaps surprisingly, most provide further evidence that cooperation is the prevailing strategy, as Axelrod summarized in 1981 in his well-cited book on the topic  \cite{axelrod1981evolution}.
+There, he also presented an evolutionary model that can be used for analyzing the game: strategies are considered to exist in a population that evolves over time in an evolutionary way (more successful strategies reproduce, and mutations can happen).
 In that context, a strategy is said to be evolutionarily stable if, supposing it controls a large share of the population, it resists being overtaken by any other strategy entering the population in small numbers.
-That is, populations of evolutionarily stable strategies form the equilibruim states of the evolutionary process, and thus, one may say that evolutionarily stable strategies are the ``best'' strategies.
+That is, populations of evolutionarily stable strategies form the equilibrium states of the evolutionary process, and thus, one may say that the evolutionarily stable strategies are the ``best'' strategies.
 
 In light of Axelrod's tournaments highlighting the effectiveness of cooperation, it has long been a goal to prove that a cooperating strategy like Tit-for-tat is evolutionarily stable.
-A few variants of this has been shown: First, Nowak has shown that in the finitely repeated game, a strategy that always defects is sometimes not evolutionarily stable \cite{nowak2004emergence}.
-Second, Binmore has shown that in the infinitely repeated game where strategies are modeled as finite automatons where having more states comes at a cost, a strategy needs to cooperate with itself to be stable \cite{binmore1992evolutionary}.
-And third, Fundenberg and Maskin have shown that a strategy needs to cooperate with itself to be stable also in the infinitely repeated game in the presence of a certain notion of infinitesimally small noise \cite{fundenberg1990evolution}.
+A few variants on this result has been shown.
+First, Nowak et al.
+showed that in the finitely repeated game, a strategy that always defects is sometimes not evolutionarily stable \cite{nowak2004emergence}.
+Second, Binmore and Samuelson showed that in the infinitely repeated game where strategies are modeled as finite automata where having more states comes at a cost, a strategy needs to cooperate with with its clone to be stable \cite{binmore1992evolutionary}.
+And third, Fudenberg and Maskin showed that a strategy needs to cooperate with its clone to be stable also in the infinitely repeated game in the presence of a certain notion of infinitesimally small noise \cite{fundenberg1990evolution}.
 
-Notably, all these results impose additional restrictions on the setup, and as noted by Fundenberg and Maskin, one has to do that: in the deterministic infinitely repeated game, one could create a strategy that can self-identify — if the opponent ever deviates from the pattern, the strategy resorts to defecting for the rest of the game.
+Notably, all these results impose additional restrictions on the setup, and as noted by Fudenberg and Maskin, restrictions are necessary: in the deterministic infinitely repeated game, one could create a strategy that can self-identify — if the opponent ever deviates from the pattern, the strategy resorts to defecting for the rest of the game.
 This strategy will be evolutionarily stable, but is not cooperative.
 
-In this paper, we, like Fundenberg and Maskin, choose the addition of noise as our restriction.
+In this paper, we, like Fudenberg and Maskin, choose the addition of noise as our restriction.
 In contrast to them, however, we present a model where, at every step, each strategy has a tiny probability $p$ of doing the wrong thing, which is arguably the most natural way of modeling noise.
+
 In \cref{sectionsetup}, we define the setup of our version of the problem in detail.
@@
 In \cref{sectionproofs}, we prove our results.
-Finally, in \cref{sectiondiscussion} we briefly discuss potential directions for future work.
+Finally, in \cref{sectiondiscussion}, we briefly discuss potential directions for future work.
 
@@
 
-In this section we define our setup.
-In summary, we consider the infinitely repeated prisoner's dilemma played by finite automata in infinite populations, evolving evolutionarily in the presence of noise.
+In this section we define our setup for the game.
+In summary, we consider the infinitely repeated prisoner's dilemma played by finite automata in infinite or continuous populations, evolving evolutionarily in the presence of noise.
 
-\subsection{Formal Definition}
+% \subsection{Formal Definition}
 
-First, we define the reward function.
+% First, we define the reward function.
 
@@
 
+Tit-for-tat is depicted as a Moore machine in \cref{tftfigure}.
+
 We will consider strategies in the presence of noise.
-To model that, we will assume that a strategy has a probability $1-p$ of following the correct transition and a probability $p$ of following the incorrect transition, at every step.
+To model that, we will assume that a strategy has a probability of $1-p$ of following the correct transition, and a probability of $p$ of following the incorrect transition, at every step.
 Note that this models noise in \textit{perception}.
-One could also imagine modeling noise in \textit{action taken}, but it is easy to see that the two are equivalent up to a change in the values of $R, T, S$ and $P$.
+One could also imagine modeling noise in \textit{action taken}, but it is not hard to see that the two are equivalent up to a change in the values of $R, T, S$ and $P$.
 
-We can now begin to define the outcome of two strategies playing against each other in the infinitely repeated game. We do this using Markov chains.
+We can now begin to define the outcome of two strategies playing against each other in the infinitely repeated game. Markov chains are well suited for this.
 
 \begin{definition}
-  Suppose that strategy $s_1$ plays against strategy $s_2$.
-This defines an \textit{$s_1$-$s_2$ Markov chain} where each state is a tuple $(c_1,c_2)$ with $c_1$ being a state in $s_1$ and $c_2$ a state in $s_2$.
-The transition probabilities are defined in the obvious way:
-  \begin{equation*}
+  Suppose that a strategy $s_1$ plays against a strategy $s_2$.
+This defines an \textit{$s_1$-$s_2$ Markov chain}, where each state is a tuple $(c_1,c_2)$, with $c_1$ being a state in $s_1$ and $c_2$ a state in $s_2$.
+The transition probabilities are defined in the obvious way: \begin{equation*}
     p_{(c_1,c_2) \to (c_1', c_2')} = p_{c_1 \to c_1'} \cdot p_{c_2 \to c_2'},
-  \end{equation*}
-  where $p_{c_1 \to c_1'}$ is 1 if $s_1$ always transitions from $c_1$ to $c_1'$, 0 if it never does, $1-p$ if the output of $s_2$ at state $c_2$ causes $s_1$ to transition from $c_1$ to $c_1'$, and $p$ if the opposite of that output causes the transition; $p_{c_2 \to c_2'}$ is defined similarly.
+  \end{equation*} where $p_{c_1 \to c_1'}$ is 1 if $s_1$ always transitions from $c_1$ to $c_1'$, 0 if it never does, $1-p$ if the output of $s_2$ at state $c_2$ causes $s_1$ to transition from $c_1$ to $c_1'$, and $p$ if the opposite of that output causes the transition; $p_{c_2 \to c_2'}$ is defined similarly.
 \end{definition}
@@
 The \textit{time average distribution} of the $s_1$-$s_2$ Markov chain, denoted $\pi$, is the distribution such that \begin{equation*}
-    \pi_{c_1, c_2} = E \left[ \text{fraction of time in state $(c_1,c_2)$} \mid \text{start in state } (a_0, b_0)  \right]
+    \pi_{c_1, c_2} = E \left[ \text{fraction of time in state $(c_1,c_2)$} \mid \text{start in state } (a_0, b_0)  \right],
   \end{equation*}
-  where the expected fraction of time is taken over the infinite sequence of states representing a random walk.
+  where the expectation is taken over the infinite sequence of states representing a random walk.
 \end{definition}
 
-Note that the time average distribution given a start state always exists, regardless of the structure of the Markov chain. Now, we can define the payoff.
+Note that the time average distribution given a start state always exists, regardless of the structure of the Markov chain.
+
+Now, we can define the payoff.
 
@@
 
-That is, the payoff that $s_1$ gets when playing against $s_2$ is simply an average of the reward it gets in each possible state of the Markov chain weighted by the fraction of time that is spent there.
+That is, the payoff that $s_1$ gets when playing against $s_2$ is simply an average of the reward it gets in each possible state of the Markov chain, weighted by the fraction of time that is spent there.
 
@@
 
-We're now ready to look at how strategies interact.
+We're now ready to look at how strategies interact in bigger numbers.
 
@@
 We can now use the fitness of a strategy to compare it with other strategies in the same population.
-If a strategy $s_1$ has a higher fitness than another strategy $s_2$, we say that the frequency of $s_1$ will increase at the expense of the frequency of $s_2$, in the next step of the evolutionary process.
+If a strategy $s_1$ has a higher fitness than another strategy $s_2$, we think of the frequency of $s_1$ as being on the increase, at the expense of the frequency of $s_2$.
 This is getting us close to how we want to define stable strategies; our next move is looking not only at a single evolutionary step, but the entire evolutionary process.
@@
     \label{fitnesscond}
-    F(s_2) > F(s_1)
+    F(s_2) > F(s_1).
   \end{equation*}
@@
 
-That is, if $s_1$ is $\epsilon$-invadable, there exists a strategy $s_2$ that can start as only a tiny fraction $\epsilon$ of the total population, and consistently have higher fitness than $s_1$, eventually causing $s_2$ to overtake and then eliminate $s_1$ completely.
-We are now finally ready to define evolutionary stability.
+That is, if $s_1$ is $\epsilon$-invadable, there exists a strategy $s_2$ that can start as only a tiny fraction $\epsilon$ of the total population, and consistently have higher fitness than $s_1$, eventually eliminating $s_1$ completely.
+With this, we can define evolutionary stability.
 
@@
 
-\todo{separate this definition out into a pure evolutionarily stable definition, and then an evolutionarily stable for infinitesimally small probabilities definition??
-would make interpretation much clearer.}
-
 That is, a strategy $s_1$ is evolutionarily stable if, as the noise probability $p$ goes to 0, it can withstand invasion attempts from any strategy that starts off as a tiny fraction of the population.
 
-\subsection{Interpretation}
+% \subsection{Interpretation}
 
-Having concluded our formal definition of evolutionary stability, it might be helpful to see examples of concrete setups in which this definition makes sense.
+% Having concluded our formal definition of evolutionary stability, it might be helpful to see an example of a concrete setup in which our definitions are applicable and make sense.
+This will shed light on which parts of our definitions may not be as natural as we would want them to be.
 
-Consider the following world.
+% Consider a square world, inhabited by
 
-\todo{WRITE THIS SECTION??
-I'M NOT SURE THAT MY MODEL MAKES MUCH SENSE ANYMORE THOUGH, BECAUSE IT REQUIRES PEOPLE TO PLAY AGAINST THEMSELVES WHICH IS A NONO.
-also the probability going to 0 does not really make sense}
+% Consider the following world.
 
@@
 
-We can now state our results! Together, the following theorem and conjecture imply that in the setup described here, mutual cooperation arises as the only stable choice.
+Together, the following theorem and conjecture imply that in the setup described here, mutual cooperation arises as the only stable choice.
 
@@
   When played against itself, it ends up spending just as much time in the defection state as in the cooperation state, because as soon as one mistake is made, it goes into a mutual defection cycle with its clone that is only broken by an additional mistake.
-  Thus, $v_s(s)$ where $s$ is Tit-for-tat is much smaller than $R$, which means that it is not evolutionarily stable by \cref{evolutionarystable1}.
-Notably, this shows that our results are different, and perhaps stronger, than Fudenberg and Maskin's: their criterion for evolutionary stability did not rule out Tit-for-tat \cite{fundenberg1990evolution}.
+  Thus, $v_s(s)$, where $s$ is Tit-for-tat, is much smaller than $R$, which means that it is not evolutionarily stable by \cref{evolutionarystable1}.
+Notably, this shows that these results are different, and perhaps stronger, than Fudenberg and Maskin's: their criterion for evolutionary stability does not rule out Tit-for-tat \cite{fundenberg1990evolution}.
 \end{remark}
@@
 Before we prove \cref{evolutionarystable1}, we need to understand the mechanisms behind the payoff $v_{s_1}(s_2)$.
-In this subsection, we prove a series of lemmas that characterize the time average distribution, and consequently $v_{s_1}(s_2)$.
+In this subsection, we prove a series of lemmas that characterize the time average distribution, and in turn $v_{s_1}(s_2)$.
 
@@
   \label{timeaverageisstationary}
-  The time average distribution $\pi^{(a,b)}$, for any starting state $(a,b)$, is a stationary distribution of the Markov chain.
-That is, if $M$ is the transition matrix of the Markov chain, we have $M \pi^{(a,b)} = \pi^{(a,b)}$.
+  The time average distribution $\pi$, for a given start state $(a,b)$, is a stationary distribution of the Markov chain.
+That is, if $M$ is the transition matrix of the Markov chain, we have $M \pi = \pi$.
 \end{lemma}
@@
 We state this lemma without proof, as it is a fairly standard result.
-Important to note is that the time average distribution is not necessarily a \textit{unique} stationary distribution, as we make no assumptions that our Markov chain be ergodic.
+Important to note is that the time average distribution is not necessarily a \textit{unique} stationary distribution, as we do not assume that our Markov chain is ergodic.
 
@@
 Let $F$ be the fraction of time spent in state $(c_1,c_2)$.
-If $(c_1,c_2)$ is not in an absorbing strongly connected component, it is well known that the fraction of time in that state $F$ will be 0.
-This is also what the lemma says.
+If $(c_1,c_2)$ is not in an absorbing strongly connected component, the fraction of time in state $F$ will then be 0.
+This agrees with what the lemma states.
 Suppose now that $(c_1,c_2)$ is in an absorbing strongly connected component $S$.
   Let $X$ be the event that the $s_1$-$s_2$ Markov chain ends up in $S$. By the law of total probability, \begin{equation*}
-    E[F] = E[F \mid X] \cdot P(X) + E[F \mid \text{not } X] \cdot P(\text{not } X)
+    E[F] = E[F \mid X] \cdot P(X) + E[F \mid \text{not } X] \cdot P(\text{not } X).
   \end{equation*}
@@
 This is continuous for all $p$ where $g(p) \neq 0$, and since $g$ will have a finite degree it is therefore continuous in a small right neighborhood of 0.
-Finally, note that $v_{s_1}(s_2)$ is between $S$ and $T$, which in conclusion means that the limit of it as $p$ goes to 0 tends to a finite number, as desired.
+Finally, note that $v_{s_1}(s_2)$ is at least $S$ and at most $T$, which in conclusion means that the limit of $v_{s_1}(s_2)$ as $p$ goes to 0 tends to a finite number, as desired.
 \end{proof}
@@
   \begin{equation*}
-    v_{s}(s) \leq R
+    v_{s}(s) \leq R.
   \end{equation*}
@@
       \begin{align*}
-        v_{s_1}(s_1) &= \gamma \pm O(p) \\
+        v_{s_1}(s_1) &= \gamma \pm O(p), \\
         v_{s_2}(s_2) &= R \pm O(p),
@@
         v_{s_1}(s_2) = \gamma \pm O(p) \text{ and }
-        v_{s_2}(s_1) = \gamma \pm O(p)
+        v_{s_2}(s_1) = \gamma \pm O(p).
       \end{align*}
@@
 Thus,
-        \begin{multline*}
+        \begin{equation*}
           P(\text{$s_1$-$s_2$ ends up in $H' \times K_H$}) = (1 - O(p)) P(\text{$s_1$-$s_1$ ends up in $H' \times H$})
-        \end{multline*}
+        \end{equation*}
         Now, by \cref{claimcanfigureout}, the probability that $s_2$ enters the absorbing strongly connected component $H_\text{copy}$ after having gotten to $K_H$ is $1 - O(p)$ when playing against $s_1$.

### 2020-05-22 paper/paper.tex
first draft

@@
 
-  We are now finally ready to prove that to be evolutionarily stable, a strategy needs to cooperate with a clone of itself.
+In this subsection, we prove that evolutionary stability requires that a strategy be cooperating with a clone of itself.
 
@@
       The underlying idea is to construct $s_2$ such that $s_1$ can see no difference between itself and $s_2$, while $s_2$, on the other hand, can.
-If we succeed in doing so, we will have that $v_{s_1}(s_2) = v_{s_2}(s_1) = v_{s_1}(s_1)$, and can construct $s_2$ such that it always cooperates when it recognizes itself, thereby yielding $v_{s_2}(s_2) = R$.
+If we succeed in doing so, we will have that $v_{s_1}(s_2) = v_{s_2}(s_1) = v_{s_1}(s_1)$, and will be able to construct $s_2$ such that it always cooperates when it recognizes itself, thereby yielding $v_{s_2}(s_2) = R$.
 This would give $s_2$ a higher fitness than $s_1$.
@@
 
-      We now construct a $K_H$ for each $H$. \Cref{c2flowchart} shows the high-level construction of $K_H$.
+      We now describe how to construct $K_H$ for each $H$. \Cref{c2flowchart} shows the high-level construction of $K_H$.
        All of the steps succeed with probability $1 - O(p)$. The first step makes sure that $K_H$ does not deviate from $s_1$ at all
@@
       The second step in the construction of $K_H$ makes sure that if $s_2$ is playing against itself, it is in sync with its clone; that is, with probability $1 - O(p)$, both copies of $s_2$ will transition to the third component at the exact same time (although not necessarily in the same $K_H$).
-      The third step now figures out what action $s_1$ would take at some large finite time $T$. 
+      The third step figures out what action $s_1$ would take at some large finite time $T$. 
       Then, $K_H$ outputs the exact opposite action at time $T$. 
-      At that point, if its opponent outputs $s_1$'s expected action, it wants to transition into the absorbing strongly connected component $H_\text{copy}$, which is just an exact copy of $H$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and enters an always cooperating state.
+      At that point, if its opponent outputs $s_1$'s expected action, it wants to transition into the absorbing strongly connected component $H_\text{copy}$, which is just an exact copy of $H$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and thus enters an always cooperating state.
       Before transitioning into $H_\text{copy}$, $s_2$ makes sure that $s_1$ is in a state that is beneficial for $s_2$.
@@
           \node[elliptic state, initial, initial where=above] (c_0) {1. Wait until $s_1$ is in an absorbing SCC, while simulating $H$.};
-          \node[elliptic state] (c_1) [below=of c_0] {2. Sync $s_2$.};
+          \node[elliptic state] (c_1) [below=of c_0] {2. Sync up with $s_2$.};
           \node[elliptic state] (c_3) [below=of c_1] {3. Figure out what action that $s_1$ will take right after this.};
@@
         \end{tikzpicture}
-        \caption{High level construction of $K_H$, the absorbing component replacing $H$ when creating $s_2$ in the proof of \cref{evolutionarystable1}.
+        \caption{High-level construction of $K_H$, the absorbing component replacing $H$ in the creation of $s_2$ in the proof of \cref{evolutionarystable1}.
 Note that $K_H$ contains exactly 2 absorbing strongly connected components.
-SCC means strongly connected component.}
+SCC is short for strongly connected component.}
         \label{c2flowchart}
@@
 We now want to prove that this $s_2$ can in fact invade $s_1$.
-To do that, we will first assume that the below four lemmas are correct, i.e., that the construction is possible and works as specified, and then finish the proof of our theorem using those assumptions.
+To do that, we will first assume that the below four lemmas are correct, that is, that the construction is possible and works as specified, and then finish the proof of our theorem using those assumptions.
 Once that's done, we will prove the four lemmas.
@@
       \begin{align*}
-        v_{s_1}(s_2) = \gamma \pm O(p) \text{ and }
-        v_{s_2}(s_1) = \gamma \pm O(p).
+        v_{s_2}(s_1) = \gamma \pm O(p) \text{ and }
+        v_{s_1}(s_2) \leq \gamma \pm O(p).
       \end{align*}
@@
       \begin{subproof}
-        By our definition of $\gamma$, we have that $v_{s_1}(s_1) \leq \gamma \pm O(p)$.
+        By our definition of $\gamma$, we have that $v_{s_1}(s_1) = \gamma \pm O(p)$.
 
@@
 By \cref{claimcanwaitfors1}, with probability $1 - O(p)$, $s_2$ is indistinguishable from $s_1$ until $s_1$ reaches an absorbing strongly connected component, which we call $H'$.
-        This means that the probability of reaching the absorbing component $H' \times K_H$ in the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we had not replaced $H$ by $K_H$ in $s_2$, which is just the probability of reaching the absorbing component $H' \times H$ in the $s_1$-$s_1$ Markov chain.
+        This means that the probability of reaching the absorbing component $H' \times K_H$ of the $s_1$-$s_2$ Markov chain, is up to a factor of $1 - O(p)$ the same as if we had not replaced $H$ by $K_H$ in $s_2$, which is just the probability of reaching the absorbing component $H' \times H$ in the $s_1$-$s_1$ Markov chain.
 Thus,
@@
         By \cref{claimcanforcegoodstate}, $s_1$ can choose to end up in any absorbing strongly connected component $S_{H', H}$ that is a subset of $H' \times H_\text{copy}$, with probability $1 - O(p)$.
-        Naturally, $s_1$ chooses an $S_{H', H}^*$ as follows: among the $S$ that have the maximum payoff for $s_2$, not caring about $O(p)$, we choose the one the minimizes the payoff for $s_1$.
-        Let $v_{s}(S)$ denote the payoff that strategy $s$ achieves in the absorbing strongly connected component $S$. Then,
+        Naturally, $s_1$ chooses an $S_{H', H}^*$ as follows: among the $S$ that have the maximum payoff for $s_2$, in the limit as $p$ goes to 0, we choose the one the minimizes the payoff for $s_1$.
+        Let $v_{s}(S)$ denote the payoff that strategy $s$ achieves in the absorbing strongly connected component $S$ of the Markov chain. Then,
         \begin{equation}
@@
         since the right hand side is a weighted average of $v_{s_1}(S)$, which is the same as $v_{s_2}(S)$ in this case, which is what we maximized over when we chose $S_{H', H}^*$.
-Also, since $s_1$ could make sure to get into it with probability $1 - O(p)$, we have
+Also, since $s_2$ could ensure with probability $1 - O(p)$ that the Markov chain enters $S_{H',H}^*$ (by \cref{claimcanforcegoodstate}), we have
         \begin{equation}
@@
         \end{equation}
-        Also, as a consequence, the $S_{H,H'}^*$ for all $H$ and $H'$ become the only absorbing strongly connected components with probabilities of ending up in them greater than $O(p)$.
+        Also, as a consequence, the collection of $S_{H,H'}^*$ for all $H$ and $H'$ become the only absorbing strongly connected components with probabilities of ending up in them greater than $O(p)$.
 
@@
 Following through with the exact same argument as above, but flipping the inequalities, we then find that $v_{s_1}(s_2) \leq \gamma \pm O(p)$.
-
+This proves the lemma.
       \end{subproof}
@@
         &= (1 - \epsilon) (\beta \pm O(p)) + \epsilon (R \pm O(p)) - (1-\epsilon) (\gamma \pm O(p)) - \epsilon (v_{s_1}(s_2) \pm O(p)) \\
-        &\geq (1 - \epsilon) (\beta - \gamma) + \epsilon(R - v_{s_1}(s_2)) \pm O(p)
+        &= (1 - \epsilon) (\beta - \gamma) + \epsilon(R - v_{s_1}(s_2)) \pm O(p)
       \end{align*}
-      Here, we can make $\epsilon$ small enough such that this becomes positive, since $\beta - \gamma > 0$.
+      Here, we can make $\epsilon$ and $p$ small enough such that this becomes positive, since $\beta - \gamma > 0$.
 Now, in the second case, we assume that $v_{s_2}(s_1) = \gamma \pm O(p)$ and $v_{s_1}(s_2) \leq \gamma \pm O(p)$.
@@
       In conclusion, our constructed $s_2$ has higher fitness than $s_1$.
-This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is not $\epsilon$-invadable some value of $p < p_0$ and some $\epsilon < \epsilon_0$.
+This proves that $s_2$ can invade $s_1$, and thus, that $s_1$ is $\epsilon$-invadable some values of $p < p_0$ and some $\epsilon < \epsilon_0$.
 In conclusion, then, $s_1$ is not evolutionarily stable, which concludes the proof of \cref{evolutionarystable1}.
@@
 
-      We now return to the three lemmas that we left out, which detail the construction of $s_2$.
+      We now return to the four lemmas that we left out, which detail the construction of $s_2$.
 
@@
 
-        Let $T_1$ be the random variable designating the time at which $s_1$ transitions into an absorbing strongly connected component, counting from the time $s_2$ entered $K_H$.
+        Let $T_1$ be a random variable designating the time at which $s_1$ transitions into an absorbing strongly connected component, counting from the time that $s_2$ entered $K_H$.
 (In particular, if $s_1$ is already in an absorbing strongly connected component, $T_1 = 0$.)
-        Let $T_2$ be the random variable designating the time at which $s_2$ transitions out of component 1 of $K_H$.
+        Let $T_2$ be a random variable designating the time at which $s_2$ transitions out of component 1 in $K_H$.
         To prove our lemma, we want to prove that with probability $1 - O(p)$, $T_1 \leq T_2$, and that during all the time that we are in component 1 of $K_H$, the output of $s_2$ is indistinguishable from $H$.
 
-        First, we deal with the indistinguishableness part.
+        First, we deal with the indistinguishability part.
 We create the component by copying $H$.
 Next, we define a new graph $W$, that we will design in the subsequent paragraphs, which will wait for $s_1$ to get into an absorbing component.
-This graph will have edges defined with the labels $C$ and $D$ still, but have no outputs at the states.
-The graph $D$ will also have one designated ``end state.'' We will then construct our component 1 for $K_H$ by replacing every state $h \in H$ with a copy of $D$ that we call $D_h$.
-There will be a $C$ edge from node $u \in D_h$ to node $v \in D_{h'}$ if and only if there is a $C$ edge from $h$ to $h'$, and from $u$ to $v$ in $D$.
+This graph will have edges defined with the labels $C$ and $D$, but will have no outputs at its states.
+The graph $W$ will also have one designated end state.
+We will then construct our component 1 for $K_H$ by replacing every state $h \in H$ with a copy of $W$ that we call $W_h$.
+There will be a $C$ edge from node $u \in W_h$ to node $v \in W_{h'}$ if and only if there is a $C$ edge from $h$ to $h'$ in $H$, and from $u$ to $v$ in $W$.
 Similarly for the $D$ edge.
-This is commonly known as the Kronecker or tensor product of the graphs $D$ and $H$.
-The output at every state $u \in D_h$ is the same as the output of $h$.
-For the end state $e \in D_h$ for each $h$, we will add an edge with label both $C$ and $D$, transitioning out of this component and into component 2.
+This is commonly known as the Kronecker or tensor product of the graphs $H$ and $W$.
+The output at every state $u \in W_h$ is the same as the output of $h$.
+For the end state $e \in W_h$ for each $h$, we will add an edge labeled both $C$ and $D$, transitioning out of this component and into component 2.
 
         It is clear that as long as we don't transition out of this component, the Kronecker product will ensure that the output will be indistinguishable from $H$.
-We now want to design $D$ such that $T_1 \leq T_2$ with high probability.
+We now want to design $W$ such that $T_1 \leq T_2$ with high probability.
 
         Let $d$ be the maximum length of a path from any state in $s_1$ to its closest absorbing strongly connected component.
-Then, for each state in $s_1$, there is a sequence of perceived inputs of length $\leq d$ (the \textit{special sequence} of that state), such that $s_1$ upon seeing that sequence ends up in an absorbing strongly connected component.
+Then, for each state in $s_1$, there is a sequence of perceived inputs of length $\leq d$ (the \textit{special sequence} for that state), such that $s_1$ upon seeing that sequence ends up in an absorbing strongly connected component.
 For all states, the probability that $s_1$
@@
 Consider the geometric random variable $X_1 = d + \operatorname{Geom} (p^d)$.
-For all $t$, then, we have that $P(X_1 = t) \leq P(T_1 = t)$.
-Thus, we may see $X_1$ as some sort of upper bound to $T_1$.
-4
-        We now describe the construction of $D$.
+We may see $X_1$ as some sort of upper bound to $T_1$.
+
+        We now describe the construction of $W$.
 Consider a state $c_1 \in s_1$ and a state $h \in H$, and suppose that we let them evolve with no mistakes happening at all.
-Then, the outputs of $c_1$ will form an infinite string $w(c_1, h)$.
+Then, the outputs of $s_1$ will form an infinite string $l(c_1, h)$.
 Let there be $|s_1|$ states in $s_1$ and $|H|$ states in $H$.
-Now, create a string $w^*$ of length $|s_1| \cdot |H|$, which differs in at least 1 place from each $w(c_1,h)$.\footnote{This is the famous technique of diagonalization.} Thus, if $s_1$ ever outputs the string $w^*$ when we are in this component, we know that at least one mistake has occurred, somewhere.
-Consequently, the probability that, at each timestep, $w^*$ will be perceived by $s_2$ as the next sequence of outputs of $s_1$ is $\leq p$.
-        If we repeat $w^*$ $d+1$ times to form $(w^*)^{d+1}$, the probability that that string is perceived by $s_2$ as the next sequence of outputs of $s_1$ is $\leq p^{d+1}$.
-        Now, create $D$ as follows: it is a chain of states $d_1,d_2,\ldots,d_{(d+1)|w^*|+1}$ where $d_i$ transitions to $d_{i+1}$ on seeing the $i$th character of the string $(w^*)^{d+1}$; $d_i$ transitions to $d_1$ otherwise.
-$d_{(d+1)|w^*|+1}$ is the designated end state of $D$.
-Now, it is clear that each step, the probability that we will reach the end state in the next $(d+1)|w^*|+1$ steps is $\leq p^{d+1}$.
-Consider the geometric random variable $X_2 = (d+1)|w^*| + 1 + \operatorname{Geom} (p^{d+1})$.
-For all $t$, then, we have that $P(X_2 = t) \geq P(T_2 = t)$.
-Thus, we may see $X_2$ as a lower bound to $T_2$.
+Now, create a string $l^*$ of length $|s_1| \cdot |H|$, which differs in at least 1 place from each $l(c_1,h)$.\footnote{This is the famous technique of diagonalization.} Thus, if $s_1$ ever outputs the string $l^*$ when we are in this component, we know that at least one mistake has occurred, somewhere.
+Consequently, the probability that, at each timestep, $l^*$ will be perceived by $s_2$ as the next sequence of outputs of $s_1$ is $\leq p$.
+        If we repeat $l^*$ for $d+1$ times to form $(l^*)^{d+1}$, the probability that that string is perceived by $s_2$ as the next sequence of outputs of $s_1$ is $\leq p^{d+1}$.
+        Now, create $W$ as follows: it is a chain of states $w_1,w_2,\ldots,w_{(d+1)|l^*|+1}$ where $w_i$ transitions to $w_{i+1}$ on seeing the $i$th character of the string $(l^*)^{d+1}$; $w_i$ transitions to $w_1$ otherwise.
+$w_{(d+1)|l^*|+1}$ is the designated end state of $W$.
+Now, it is clear that each step, the probability that we will reach the end state in the next $(d+1)|l^*|+1$ steps is $\leq p^{d+1}$.
+Consider the geometric random variable $X_2 = (d+1)|l^*| + 1 + \operatorname{Geom} (p^{d+1})$.
+We may see $X_2$ as a lower bound to $T_2$.
 
@@
 It is a chain of states where each state transitions to the next on both input $C$ and $D$ (that is, deterministically).
-Let $w_H^*$ be the string used in the proof of \cref{claimcanwaitfors1} for component $H$.
-Let $W = \sum_{H} (d+1) | w_H^* |$.
-Then, the outputs of the first $W$ states form the string $w_H^*$ repeated $d+1$ times, for all $w_H$.
-The following $W+1$ states always output $D$.
+Let $l_H^*$ be the string used in the proof of \cref{claimcanwaitfors1} for component $H$.
+Let $L = \sum_{H} (d+1) | l_H^* |$.
+Then, the outputs of the first $L$ states form the string $l_H^*$ repeated $d+1$ times, for all $l_H$.
+The following $L+1$ states always output $D$.
 This is followed by one last state outputting $C$.
@@
 Then, by \cref{claimcanwaitfors1}, using the fact that $s_2'$ is a copy of $s_1$ before it reaches an absorbing component, $s_2$ will not enter component 2 of $K_H$ until $s_2'$ enters an absorbing component $K_{H'}$, with probability $1 - O(p)$.
-When $s_2$ enters component 2, it will at some point output the entirety of the string $w_{H'}^*$.
+When $s_2$ enters component 2, it will at some point output the entirety of the string $l_{H'}^*$.
 With probability $1 - O(p)$, $s_2'$ will then, by our construction of component 1, enter component 2.
-That is, $s_2'$ enters the first part of component 2 before $s_2$ leaves the first part of component 2, which means that $s_2$ is at most $W$ steps ahead of $s_2$.
+That is, $s_2'$ enters the first part of component 2 before $s_2$ leaves the first part of component 2, which means that $s_2$ is at most $L$ steps ahead of $s_2$.
 This means that $s_2'$ will be in the defection part of the chain when $s_2$ reaches the cooperation state.
@@
 
-    We have some, admittedly weak, empirical evidence for this theorem to be true.
-First, Pavlov when played against itself gets payoff $R - O(p)$, and our proof of \cref{evolutionarystable1} is very clearly relying heavily on letting $p$ be very close to 0, so it seems unlikely that a similar proof would be able to prove that Pavlov is not evolutionarily stable.
-Moreover, Fudenberg and Maskin stated but did not prove a similar result in their version of the noisy regime \cite{fundenberg1990evolution}.
-We also wrote a Julia notebook for generating $s_1$-$s_2$ Markov chains and the time average distributions, and ran Pavlov against a few variants of Pavlov and other simple strategies like All-D, All-C and Tit-for-tat, and in all cases Pavlov performed better than our potential invaders; this can be found in the accompanying GitHub repository \cite{arvid2020}.
-That repository also contains a document of unorganized notes that may provide hints for proving the conjecture.
+    We have some, admittedly weak, empirical evidence supporting the truth of this conjecture.
+First, playing Pavlov against itself yields the payoff $R - O(p)$, and our proof of \cref{evolutionarystable1} is very clearly relying heavily on letting $p$ be very close to 0.
+It therefore seems unlikely that a similar proof would be able to prove that Pavlov is not evolutionarily stable.
+Moreover, Fudenberg and Maskin stated but did not prove a similar result in their version of the noisy game \cite{fundenberg1990evolution}.
+To experiment, we also wrote a Julia notebook for generating $s_1$-$s_2$ Markov chains and time average distributions, and ran Pavlov against a few variants of itself, as well as against other simple strategies like All-D, All-C and Tit-for-tat.
+In all cases, Pavlov performed better than our potential invaders.
+This code notebook can be found in the accompanying GitHub repository \cite{arvid2020}.
 
@@
 
-    An obvious next step is to prove \cref{pavlovtheorem}. We believe that it should not be too hard.
+    An obvious next step is to prove \cref{pavlovtheorem}. We believe that it should not be too hard. The handwritten notes in \cite{arvid2020} might contain some useful ideas.
 
-    One may also consider different setups.
+    One may also consider other setups.
 For example, weakening \cref{definitionevolutionarystability} of evolutionary stability by saying that it is only required to not be $\epsilon$-invadable for a fixed $p$, instead of for infinitesimally small $p$, would lead to a more interpretable model.

### 2020-05-22 workinprogress/latexnotes/notes.tex
organize!

--- 2020-04-05 latexnotes/notes.tex
@@
 \newtheorem{corollary}{Corollary}
+\newtheorem{lemma}{Lemma}
 \newtheorem{problem}{Problem}
@@
   The \textit{fitness} of a strategy $s$ in a population $P = (S, f)$ is \begin{equation*}
-    F(s) = \sum_{s' \in S} f(s) v(s, s').
+    F(s) = \sum_{s' \in S} f(s') v(s, s').
   \end{equation*}
@@
 
+Thus, \textbf{I need to change the definitions} to get something meaningful out of this.
+Actually, what this also means is that it further goes to show that the results presented using the weird model of noise are weird as well, as the result they proved using that model is provably not reproducible in this model.
+
+I can think of two ways of fixing this problem by changing the definitions:
+
+\begin{enumerate}
+  \item Make $p \to 0$ earlier. i.e. instead of fixing $p$ and then letting $\alpha \to 0$, fix $\alpha$ and let $p \to 0$ or something similar.
+  \item Define invasion differently: a strategy is ES iff given an invasion fraction $\alpha$ (for sufficiently small $\alpha$) there exists no strategy that can achieve a higher fitness.
+\end{enumerate}
+
+There might be other ways of solving this; if so, add to this list later.
+For now, I should analyze the two methods in separation.
+Later, I can determine which of the two methods is more reasonable to adopt.
+
+\subsection*{Make $p \to 0$ happen earlier}
+
+The simplest way: define $v(s_1, s_2)$ as the limit $\lim_{p \to 0}$ of the stationary distribution payoff.
+In this situation, theorem 1 is true, and can be proved by using a very simple strategy that just self-identifies on the first turn and then copies.
+\textbf{Pros:} This gives us exactly the result we want.
+\textbf{Cons:} I don't think that result means much, because it is impossible to relate it to a real situation since there is no actual value of $p$ for which the result would hold (not even $p = 0$) which just means that it's some sort of fictional result.
+
+Can we take a middle ground? Could say something like: A strategy $s_2$ can invade a strategy $s_1$ if there exists a real 
+\begin{definition}
+  A strategy $s_2$ can \textit{invade} a strategy $s_1$, if there exists a real number $\alpha \in (0,1]$, such that for all populations $P$ with $S = \{s_1, s_2\}$ and $f(s_2) < \alpha$, there exists a value of $p > 0$ such that
+  \begin{equation*}
+    F(s_2) > F(s_1).
+  \end{equation*}
+\end{definition}
+This means that the same strategy $s_2$ can invade $s_1$ but that the $p$ can possibly change from time to time.
+This would also solve the AllD problem, and it would also prove theorem 1, while making more sense (albeit not much sense either).
+
+\subsection*{Redefine invasion}
+
+This is sort of similar to our second approach above actually. But here we take a different approach: instead of different $p$ for every $\alpha$, we take different $s_2$ for different $\alpha$.
+
+The following is one way of defining invasion in this way.
+\begin{definition}
+  A strategy $s_1$ can be invaded if there exists a real number $\alpha \in (0,1]$ such that for all $\epsilon < \alpha$, there exists a strategy $s_2$ such that in the population $P$ with $S = \{s_1,s_2\}$ and $f(s_2) = \epsilon$, we have \begin{equation*}
+    F(s_2) > F(s_1)
+  \end{equation*}
+\end{definition}
+
+Does this alleviate the AllD problem?
+Yes, and it also proves theorem 1, using the same (boring) construction (except that the self-identification now may take $k$ steps instead of only 1 depending on how small the population is).
+Is this reasonable?
+Not really, because it doesn't say anything about what happens to that strategy next.
+
+\subsection*{Other approaches}
+
+So we can see that all three possible ways to solve the problem are somewhat unsatisfactory. New ideas:
+\begin{enumerate}
+  \item Include evolutionary drift into the picture: a strategy is not evolutionary stable if it is vulnerable to a 2-stage attack whereby 1 startegy first enters and survives because of evolutionary drift, and then another comes in and colludes with the second one to take over.
+  \item Move away from infinitely repeated games (no no seems very messy and sad)
+  \item Think of another model that better captures the finiteness of real population (which also helps with the $\alpha \to 0$ problem).
+  \item A different model of computation.
+Note that the Moore machines I'm using have a very clear start.
+This, one could say, does not really model reality that well, because there will not necessarily be a clear starting point.
+  \item Another way of evaluating fitness: think that the individuals are spread out geographically which means that even in a very large population, the initial strategy mutation injection will first face a small fraction of the population, then take over that part, then face a larger proportion but now be larger itself, and so on and so forth.
+Thinking this way, as long as the span of higher fitness grows by something like $r^2$ in a 2D model, and there is some $\alpha$ where the invader is better, then it can eventually take over most of the population which is very interesting.
+\end{enumerate}
+
+\subsection*{Evolutionary drift and multiple different strategies}
+
+hmm.
+This has the same AllD problem as the original setup, using the same argument.
+The thing with noise really is that you can't really hide a secret that is useful (and when AllD finds out it will end up punishing you for it)
+
+[other approaches here]
+
+\subsection*{2D model}
+
+Okay so this is cool.
+Think of 1 strategy as covering a large circle.
+then start another strategy as a very very tiny circle.
+who does it interact with?
+could also think of this in terms of the grid.
+Hmmmm ok at the very least this would motivate why it is moew reasonable to say that a strategy can invade another strategy as long as there exists some $\alpha$ for which it can invade: because if a strategy only engages with other strategies within a certain fixed radius, the same strategy could then invade a given strategy regardless of how big the population is, which isn't really principally important but still kinda gives the definition a tiny bit more legitimacy.
+The nice thing about the circle 2D model is that it also doesn't require that the population size stays constant throughout -- it could actually fluctuate quite a bit.
+
+\subsection*{Redefine invasion}
+
+Can we use definition 9 but prove that it implies that it works for all $\alpha \geq \varepsilon$ and $\alpha <$ maybe $\frac 1 2$?
+That would be cool because it would show that that model actually makes a lot of sense.
+OK so that is not true.
+But generally, if we find a strategy that can invade using definition 9 we can almost certainly find a strategy that can invade using the natural definition.
+Wait so I think that that means that they're equivalent.
+Let me state it
+
+\begin{definition}
+  A strategy $s_1$ can be invaded if for all $\varepsilon \in [0,1]$ (and for all $p \in (0,1)$????) there exists a strategy $s_2$ such that in all populations $P$ with $S = \{s_1,s_2\}$ and $f(s_2) \geq \epsilon$, we have \begin{equation*}
+    F(s_2) > F(s_1)
+  \end{equation*}
+\end{definition}
+
+So we have the concept of invadable9 and invadable10.
+If something is invadable9 then I think we can also make it be invadable10.
+And if it is invadable10 then certainly it is invadable9.
+So yay then definition 9 and definition 10 are equivalent which is amazing.
+
+(The proof showing invadable9 implies invadable10 is somewhat interesting: you take a strategy that can invade according to 9 and then you modify it so that it is more or less the same but also has $v(s_1,s_1) + v(s_2,s_2) \geq v(s_1,s_2) + v(s_2,s_1)$ which then would show according to ipad calculations that it the new $s_2$ would be able to invade for all bigger $\varepsilon$ which completes the proof.)
+
+So we can use whichever definition we like.
+Note that definition 10 has a very nice interpretation: $\varepsilon$ and $p$ can be thought of as two constants of nature (representing the number of interactions and probability of failure, respectively), and then the definition says that a strategy can be invaded if regardless of what those parameters are, there is some bad strategy that can pop up and take over.
+In other words, a strategy is evolutionary stable if there exists some constants of nature that can make it prevail.
+So in this sense the evolutionary stability becomes a pretty weak concept.
+Showing a strong result about a weak concept -- e.g.
+the utilitarianism result -- would thus be kind of remarkable.
+Is this notion of ESS weaker than the usual notion of ESS?
+No because it excludes AllD.
+But at the same time, the result will be stronger than any results proved earlier, because probably only Pavlov will be good (although I'm still worried about the $p \to 0$ necessities of this statement.
+But we'll see about that when we get there.)
+
+Note also that definition 9 is nice for proving stuff, because the condition is only to find 1 epsilon and not infinitely many.
+
+\subsection*{Implications of definitions 9 and 10}
+
+\section*{Interesting Findings}
+
+Using definition 9/10, we can prove the following:
+
+\begin{enumerate}
+  \item For a given $p$, if a strategy $s$ is ESS then $v(s,s) \geq \frac{S + T}{2}$.
+  \item For small $p$, if a strategy $s$ is ESS then $\lim_{p \to 0} v(s,s) \geq R$.
+\end{enumerate}
+
+This is great.
+The only problematic thing is that not every strategy that tends to $R$ will be a good one I think (although I'm not sure about that actually).
+And I also need to prove that Pavlov, for example, actually is ESS, because otherwise the definition is kinda stupid.
+But this is still very nice and promising.
+
+Note that all of this can be proven using the obvious self-identifying strategy, which goes to show that the original idea of having someone try to mimic the behavior of the other group but at the same time try to exploit it behind their backs actually works in theory too which is really cool I think.
+
+OK how can we prove that Pavlov is actually ESS? Which definition is easier to use? (probably 9a again).
+
+OK I will use definition 9 I think. ahhhh so much to do here and I'm not even sure that my result is valid.
+
+\begin{lemma}
+  Suppose $s_1$ is a strategy such that $v(s_1,s_1) < \frac{S + T}{2}$. Then $s_1$ is not ESS according to definition 9.
+\end{lemma}
+\begin{proof}
+  We will show this by exhibiting a strategy $s_2$ and a parameter $\epsilon$.
+Let the strategy $s_2$ be such that it ``tests'' its opponent for $k$ steps.
+(uhhh ok this gets a little more complicated I think because it is just not a simple exponential).
+\end{proof}
+
+\section*{For reference}
+
+Pavlov against itself gets the payoff: \begin{equation*}
+  v(s_\text{Pavlov}, s_\text{Pavlov}) = (-4p^3+7p^2-4p+1, p(1-p), p(1-p), 4p^3-5p^2+2p) \cdot (R, S, T, P)
+\end{equation*}
+which tends to 1 as $p \to 0$.
+
 \end{document}

### 2020-05-22 workinprogress/realnotes/notes.tex
organize!

--- 2020-04-09 realnotes/notes.tex
@@
 \begin{theorem}
+  \label{evolutionarystable1}
   Suppose $s_1$ is evolutionary stable. Then $v_{s_1}(s_1) \geq \frac{S + T}{2}$.
@@
 \begin{theorem}
+  \label{evolutionarystable2}
   Suppose $s_1$ is evolutionary stable as $p$ goes to 0 (i.e., that it is evolutionary stable if condition \ref{fitnesscond} is replaced by $\lim_{p \to 0} (F (s_2)-F(s_1)) > 0$).
@@
 
+\begin{proof}[Proof of \cref{evolutionarystable1}]
+    Proof idea: Suppose $s_1$ is not utilitarian. Then, for every $\epsilon > 0$, there exists a strategy $s_2$ for which $F(s_2) > F(s_1)$. (where is $p$???)
+
+    OK I think I'm doing: suppose $s_1$ is not utilitarian. Then, we want to show that it can be invaded for some $\epsilon$ (and we'll worry about actually making that work later).
+
+    OK so let's just show this: suppose $s_1$ has value $\gamma < (S + T)/2$ against itself.
+show that there is an $\alpha$ and an $s_2$ such that $s_1$ gets lower fitness than $s_2$.
+THINK ABOUT WHAT THIS MEANS LATER.
+
+    We define $s_2$ as the thing with the lonely branch that we had earlier. Then we get the following.
+    \begin{equation}
+      v_{s_2}(s_1) = (1 - p^k) \gamma  + p^k S
+    \end{equation}
+    \begin{equation}
+      v_{s_2}(s_2) = (1-p)^{2k} R + 2 (1-p)^k (1 - (1-p)^k) (\tfrac{S + T}{2}) + (1 - (1-p)^k)^2 \gamma
+    \end{equation}
+    \begin{equation}
+      v_{s_1}(s_1) = \gamma
+    \end{equation}
+    \begin{equation}
+      v_{s_1}(s_2) = (1-p^k) \gamma + p^k T
+    \end{equation}
+
+    Suppose now that $f(s_2) = \alpha$ and $f(s_1) = 1 - \alpha$.
+
+    We want to prove the following. Note that we are still free to pick both $k$ and $\alpha$ freely, by the way we set up our claim. \begin{equation}
+      (1- \alpha) v_{s_2}(s_1) + \alpha v_{s_2}(s_2) > (1-\alpha) v_{s_1}(s_1) + \alpha v_{s_1}(s_2)
+    \end{equation}
+
+    We substitute in the definitions from before. We get. \begin{multline}
+      (1- \alpha) \big((1 - p^k) \gamma + p^k S\big) \\
+      + \alpha \big( (1-p)^{2k} R + 2 (1-p)^k (1 - (1-p)^k) (\tfrac{S + T}{2}) + (1 - (1-p)^k)^2 \gamma \big) \\
+      > (1-\alpha) \big(  \gamma \big) \\
+      + \alpha \big(  (1-p^k) \gamma + p^k T  \big)
+    \end{multline}
+
+\end{proof}
+
+    OHHHHHH OK SO DO THIS: WE ALSO CHOOSE $p$.
+SUPPOSE THAT $\gamma < R - \beta$ for some $\beta > 0$.
+THEN, we can choose an $\alpha$ and a $k$ and a $p$ such that we can invade.
+Additionally, it will be true that for any $p$ smaller than the one we choose we will be able to invade.
+So that is, for sufficiently small $p$, $s_1$ will be $\epsilon$-invadable for some $\epsilon$.
+THIS MAKES SENSE FINALLY ARVID.
+ok cool now break, talking to dave.
+i'm still not totally convinced that this is correct but oh well.
+
+    Our new definition of evolutionary stable:
+
+    \begin{definition}
+      A strategy $s_1$ is \textit{evolutionary stable} if there exists $p_0$ and $\alpha$, both in $(0,1)$, such that for all $p < p_0$, and all $\epsilon < \alpha$, $s_1$ is not $\epsilon$-invadable.
+    \end{definition}
+
+    \begin{theorem}
+      \label{preciseevolutionarystable}
+      Suppose $s_1$ is evolutionary stable. Then, 
+      \begin{equation*}
+        v_{s_1}(s_1) \geq \left(\frac{1-p}{1+p}\right) R
+      \end{equation*}
+    \end{theorem}
+    \begin{proof}
+      Suppose $s_1$ is such that \begin{equation*}
+        v_{s_1}(s_1) = \gamma < R \left(\frac{1-p}{1+p}\right).
+      \end{equation*}
+      We want to prove that $s_1$ is not evolutionary stable, and to do that, we want to prove that for any sufficiently small $\epsilon$, there exists a strategy $s_2$ that can invade $s_1$.
+
+      We create the strategy $s_2$ as follows.
+First, copy the entire $s_1$ machine into $s_2$.
+Suppose the state corresponding to the start state of $s_1$ is $c_s$.
+Let the output at $c_s$ be $G(c_s)$.
+Let the node it goes to upon perceiving the opponent move $G(c_s)$ be $T(c_s, G(c_s))$.
+Then, create a new state $c_0$ that outputs $\lnot G(c_s)$ and has transition $T(c_0, G(c_s)) = T(c_s, G(c_s))$.
+Create another new state $c_1$.
+Let $T(c_0, \lnot G(c_s)) = c_1$.
+Let $T(c_2, \cdot) = c_2$.
+Let $G(c_2) = C$.
+This completely describes $s_2$.
+
+      Now, we claim that the payoffs are as follows.
+      \begin{equation}
+        v_{s_2}(s_1) = (1 - p) \gamma  + p S
+      \end{equation}
+      \begin{equation}
+        v_{s_2}(s_2) = (1-p)^{2} R + 2 (1-p) p (\tfrac{S + T}{2}) + p^2 \gamma
+      \end{equation}
+      \begin{equation}
+        v_{s_1}(s_1) = \gamma
+      \end{equation}
+      \begin{equation}
+        v_{s_1}(s_2) = (1-p) \gamma + p T
+      \end{equation}
+
+      Now, we simply compute $F(s_2) - F(s_1)$, and want to show that it is greater than 0 for all $\alpha$ that are sufficiently small.
+
+      \begin{align*}
+        F(s_2) - F(s_1) &= \\
+        &= (1 - \alpha) \cdot v_{s_2}(s_1) + \alpha \cdot v_{s_2}(s_2) - (1 - \alpha) \cdot v_{s_1}(s_1) - \alpha \cdot v_{s_1}(s_2) \\
+        &= 
+      \end{align*}
+
+    \end{proof}
+
 \begin{figure}

### 2020-05-25 paper/paper.tex
finishing touches

@@
     \end{tabularx}
-    \caption{The prisoner's dilemma. The payoff of the player picking the row is listed first.}
+    \caption{The prisoner's dilemma. One player picks a row, the other a column. The payoff of the player picking the row is listed first.}
     \label{fig:prisonersdilemmatable}
@@
 One might loosely connect this repeated multiplayer game to evolution: why have humans evolved to cooperate with each other?
-In a seminal paper in 1980, Axelrod presented evidence that in a repeated setting with multiple players facing off in a tournament, cooperation can arise as the strategy of choice, even for a selfish player \cite{axelrod1980effective}.
+In a seminal paper published in 1980, Axelrod \cite{axelrod1980effective} presented evidence that in a repeated setting with multiple players facing off in a tournament, cooperation can arise as the strategy of choice, even for a selfish player.
 In particular, the mostly cooperating Tit-for-tat strategy, depicted in \cref{tftfigure}, won his tournament.
-The idea is that even though Tit-for-tat loses when it plays against defecting strategies, it is being heavily rewarded when it cooperates with other cooperating strategies, giving Tit-for-tat a net advantage.
+The idea is that even though Tit-for-tat loses when it plays against defecting strategies, it is being heavily rewarded when it cooperates with other cooperating strategies, resulting in a net advantage for Tit-for-tat.
 
@@
   \caption{The Tit-for-tat strategy.
-At any one point, it is in one of two states, taking the action corresponding to the label of the state.
-Upon perceiving its opponent's move, it decides to switch state if the opponent does something different.}
+At any one point, it is in one of two states, taking the action corresponding to the label of that state.
+It always switches state so as to copy the opponent's last move.}
   \label{tftfigure}
@@
 Therefore, follow-up tournaments and simulations have been run, and perhaps surprisingly, most provide further evidence that cooperation is the prevailing strategy, as Axelrod summarized in 1981 in his well-cited book on the topic  \cite{axelrod1981evolution}.
-There, he also presented an evolutionary model that can be used for analyzing the game: strategies are considered to exist in a population that evolves over time in an evolutionary way (more successful strategies reproduce, and mutations can happen).
+There, he also presented an evolutionary model that can be used for analyzing the game: strategies are considered to exist in a population that evolves over time in an evolutionary way (more successful strategies reproduce, and random new strategies are introduced in small numbers).
 In that context, a strategy is said to be evolutionarily stable if, supposing it controls a large share of the population, it resists being overtaken by any other strategy entering the population in small numbers.
 That is, populations of evolutionarily stable strategies form the equilibrium states of the evolutionary process, and thus, one may say that the evolutionarily stable strategies are the ``best'' strategies.
+This characterizes the best strategies without relying on a particular composition of strategies to begin with.
 
@@
 First, Nowak et al.
-showed that in the finitely repeated game, a strategy that always defects is sometimes not evolutionarily stable \cite{nowak2004emergence}.
-Second, Binmore and Samuelson showed that in the infinitely repeated game where strategies are modeled as finite automata where having more states comes at a cost, a strategy needs to cooperate with with its clone to be stable \cite{binmore1992evolutionary}.
-And third, Fudenberg and Maskin showed that a strategy needs to cooperate with its clone to be stable also in the infinitely repeated game in the presence of a certain notion of infinitesimally small noise \cite{fundenberg1990evolution}.
+\cite{nowak2004emergence} showed that in the finitely repeated game, a strategy that always defects is sometimes not evolutionarily stable.
+Second, Binmore and Samuelson \cite{binmore1992evolutionary} showed that in the infinitely repeated game where strategies are modeled as finite automata where having more states comes at a cost, a strategy needs to cooperate with its clone to be stable.
+And third, Fudenberg and Maskin \cite{fundenberg1990evolution} showed that a strategy needs to cooperate with its clone to be stable also in the infinitely repeated game in the presence of a certain notion of infinitesimally small noise.
 
-Notably, all these results impose additional restrictions on the setup, and as noted by Fudenberg and Maskin, restrictions are necessary: in the deterministic infinitely repeated game, one could create a strategy that can self-identify — if the opponent ever deviates from the pattern, the strategy resorts to defecting for the rest of the game.
-This strategy will be evolutionarily stable, but is not cooperative.
+Notably, all these results impose additional restrictions on the setup, and as noted by Fudenberg and Maskin, restrictions are necessary: in the deterministic infinitely repeated game, one could create a strategy that cooperates only with itself — if the opponent ever deviates from what the strategy would do, the strategy resorts to defecting for the rest of the game.
+This strategy will be evolutionarily stable, but is not necessarily cooperative.
 
@@
 In \cref{sectionsetup}, we define the setup of our version of the problem in detail.
-Then, in \cref{sectionresults}, we state the two main results: that a strategy needs to be cooperative to be evolutionarily stable, and that such evolutionarily stable strategies exist.
+Then, in \cref{sectionresults}, we state the two main results: a theorem saying that a strategy needs to be cooperative to be evolutionarily stable, and conjecture saying that such evolutionarily stable strategies exist.
 In \cref{sectionproofs}, we prove our results.
@@
   \end{equation*}
-  where the expectation is taken over the infinite sequence of states representing a random walk.
+  where the expectation is taken over the infinite sequence of states representing a random walk in the Markov chain.
 \end{definition}
@@
 
-Now, we can define the payoff.
+Now, we can define the payoff that one strategy gets when playing against another one.
 
@@
 
-\iffalse
-We use the notation $G_{s_1,s_2}(c_1,c_2)$ to refer to the vector $(G_{s_1}(c_1), G_{s_2}(c_2))$, and we use $S_{s_1,s_2}$ to refer to the set of all states in the Markov chain.
-
-\begin{definition}
-  \label{strategypayoffs}
-  Let $X_t$ be the random variable designating which state the $s_1$-$s_2$ Markov chain is in at time $t$. The \text{payoff} of strategy $s_1$ when played against strategy $s_2$ is 
-  \begin{equation*}
-    v_{s_1}(s_2) = E \left[ \lim_{T \to \infty} \frac{1}{T} \sum_{t=0}^T r(G_{s_1,s_2} (X_t)) \Bigm| X_0 = (c_{\text{start}}(s_1), c_{\text{start}}(s_2)) \right]
-  \end{equation*}
-\end{definition}
-
-That is, when $s_1$ plays against $s_2$, we define its payoff to be the average payoff over all possible infinite sequences of moves.
-Note that the expectation is taken over the infinite sequence $(X_0, X_1, \ldots)$.
-The limit inside is thus simply a normal time average limit of bounded real numbers, which clearly exists.
-
-We will now introduce the notion of a time average distribution which will lead us to a second way of defining the payoff $v_{s_1}(s_2)$.
-\todo{is this really necessary??
-should i pick one or the other??
-should i move one of them to the next section?
-which one is easier to understand?
-do they trivially say the same thing?}
-
-\begin{definition}
-  \label{timeaveragedistribution}
-  The \textit{time average distribution} of the $s_1$-$s_2$ Markov chain given the start state $(a, b)$, denoted $\pi^{(a, b)}$, is the distribution such that \begin{equation*}
-    \pi_{c_1, c_2}^{(a, b)} = E \left[ \text{fraction of time in state $(c_1,c_2)$} \mid \text{initial state is } (a, b)  \right]
-  \end{equation*}
-  where the expected fraction of time is taken over the infinite sequence $(X_0, X_1, \ldots)$.
-\end{definition}
-
-We will use $\pi$ to refer to $\pi^{(c_{\text{start}}(s_1), c_\text{start}(s_2))}$.
-
-\begin{lemma}
-  \label{payofftimeaverage}
-  The payoff when $s_1$ plays against $s_2$ is
-  \begin{equation*}
-    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} \pi_{c_1,c_2} \cdot r(G_{s_1}(c_1),G_{s_2}(c_2)).
-  \end{equation*}
-\end{lemma}
-
-We may also make $r(G_{s_1}(c_1),G_{s_2}(c_2))$ into a vector, denoted by $r$, and write this as the dot product \begin{equation*}
-  v_{s_1}(s_2) = \pi \cdot r.
-\end{equation*}
-
-\todo{I think the best way to handle this is to remove the expected value definition altogether.
-just do the time average distribution and skip the proof of the lemma.
-it is honestly a very natural definition}
-
-\begin{proof}[Proof of \cref{payofftimeaverage}]
-  The key idea is that a time average sum where each element is one of finitely many values can be written as a frequency-weighted finite sum instead.
-Let $I_{c_1,c_2,t}$ be the indicator variable that is 1 if $G_{s_1,s_2}(X_t) = (c_1,c_2)$ and 0 otherwise.
-Then, we can write \begin{equation*}
-    \lim_{T \to \infty} \frac{1}{T} \sum_{t = 0}^T r(G_{s_1,s_2}(X_t)) = 
-    \lim_{T \to \infty}\frac{1}{T} \sum_{t = 0}^T \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1,c_2) \cdot I_{c_1,c_2,t}
-  \end{equation*}
-  We may now exchange the order of summation and move the finite sum out of the limit, to get \begin{equation*}
-    \lim_{T \to \infty}\frac{1}{T} \sum_{t = 0}^T \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1,c_2) \cdot I_{c_1,c_2,t}
-    = \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1,c_2) \cdot \lim_{T \to \infty} \sum_{t=0}^T \frac{I_{c_1,c_2,t} }{T}
-  \end{equation*}
-  We can now use \cref{strategypayoffs} and linearity of expectation to find that
-  \begin{equation*}
-    v_{s_1}(s_2) = \sum_{(c_1,c_2) \in S_{s_1,s_2}} r(c_1, c_2) E \left[ \lim_{T \to \infty} \sum_{t=0}^T \frac{I_{c_1,c_2,t}}{T} \bigm| X_0 = (c_{\text{start}}(s_1), c_{\text{start}}(s_2))
-    \right]
-  \end{equation*}
-  Finally, we note that this is exactly the statement of \cref{payofftimeaverage}, which proves our lemma.
-\end{proof}
-
-Appendix A contains more details on time average distributions.
-In particular, if a unique stationary distribution exists, it is equal to the time-average distribution, which enables us to quickly find the time-average distribution in many cases.
-
-\fi
-
-We're now ready to look at how strategies interact in bigger numbers.
+We're now ready to look at how strategies interact in a multiplayer setting.
 
@@
   \label{evolutionarystable1}
-  Suppose that a strategy $s_1$ is evolutionarily stable. Then $\lim_{p \to 0} v_{s_1}(s_1) = R$.
+  Suppose that a strategy $s_1$ is evolutionarily stable. Then \begin{equation*}
+  \lim_{p \to 0} v_{s_1}(s_1) = R.
+  \end{equation*}
 \end{theorem}
@@
   Thus, $v_s(s)$, where $s$ is Tit-for-tat, is much smaller than $R$, which means that it is not evolutionarily stable by \cref{evolutionarystable1}.
-Notably, this shows that these results are different, and perhaps stronger, than Fudenberg and Maskin's: their criterion for evolutionary stability does not rule out Tit-for-tat \cite{fundenberg1990evolution}.
+Notably, this shows that these results are different, and perhaps stronger, than Fudenberg and Maskin's \cite{fundenberg1990evolution}: while Tit-for-tat is not stable in their model either, their criterion for evolutionary stability does not rule out Tit-for-tat.
 \end{remark}
@@
 
-Before we prove \cref{evolutionarystable1}, we need to understand the mechanisms behind the payoff $v_{s_1}(s_2)$.
+Before we prove \cref{evolutionarystable1}, we need to understand the mechanisms governing the payoff $v_{s_1}(s_2)$.
 In this subsection, we prove a series of lemmas that characterize the time average distribution, and in turn $v_{s_1}(s_2)$.
@@
 \begin{definition}
-  A \textit{strongly connected component} of a directed graph is a subgraph where there is a path from every node to every other node.
+  A \textit{strongly connected component} of a directed graph is a maximal subgraph where there is a path from every node to every other node.
 \end{definition}
@@
   \label{lemmaabsorbinghasuniquetimeaverage}
-  An absorbing strongly connected component has a unique time average distribution. That is, the time average distribution does not depend on the start state.
+  An absorbing strongly connected component has a unique time average distribution.
+That is, the time average distribution of the component does not depend on the start state (as long as it is inside the component).
 \end{lemma}
@@
        absorbing strongly connected components of $s_1$.
-The key idea, now, is to replace each original absorbing strongly connected component $H \in \mathcal{A}_1$ with a new absorbing component $K_H$, which has the capability of identifying itself.
+The key idea, now, is to replace each original absorbing strongly connected component $H \in \mathcal{A}_1$ with a new absorbing component $K_H$, which has the capability of distinguishing between itself and $s_1$.
 
@@
       Then, $K_H$ outputs the exact opposite action at time $T$. 
-      At that point, if its opponent outputs $s_1$'s expected action, it wants to transition into the absorbing strongly connected component $H_\text{copy}$, which is just an exact copy of $H$, but if its opponent outputs the opposite of that action, it has successfully recognized itself and thus enters an always cooperating state.
+      At that point, if its opponent outputs $s_1$'s expected action, it wants to transition into the absorbing strongly connected component $H_\text{copy}$, which is just an exact copy of $H$, but if its opponent outputs the opposite of that action, it has successfully distinguished its opponent from $s_1$ and thus enters an always cooperating state.
       Before transitioning into $H_\text{copy}$, $s_2$ makes sure that $s_1$ is in a state that is beneficial for $s_2$.
@@
         % \end{equation}
-        By \cref{claimcanforcegoodstate}, $s_1$ can choose to end up in any absorbing strongly connected component $S_{H', H}$ that is a subset of $H' \times H_\text{copy}$, with probability $1 - O(p)$.
-        Naturally, $s_1$ chooses an $S_{H', H}^*$ as follows: among the $S$ that have the maximum payoff for $s_2$, in the limit as $p$ goes to 0, we choose the one the minimizes the payoff for $s_1$.
+        By \cref{claimcanforcegoodstate}, $s_2$ can choose to end up in any absorbing strongly connected component $S_{H', H}$ that is a subset of $H' \times H_\text{copy}$, with probability $1 - O(p)$.
+        Naturally, $s_2$ chooses an $S_{H', H}^*$ as follows: among the $S$ that have the maximum payoff for $s_2$, in the limit as $p$ goes to 0, we choose the one the minimizes the payoff for $s_1$.
         Let $v_{s}(S)$ denote the payoff that strategy $s$ achieves in the absorbing strongly connected component $S$ of the Markov chain. Then,
@@
 It therefore seems unlikely that a similar proof would be able to prove that Pavlov is not evolutionarily stable.
-Moreover, Fudenberg and Maskin stated but did not prove a similar result in their version of the noisy game \cite{fundenberg1990evolution}.
+Moreover, Fudenberg and Maskin \cite{fundenberg1990evolution} stated but did not prove a similar result in their version of the noisy game.
 To experiment, we also wrote a Julia notebook for generating $s_1$-$s_2$ Markov chains and time average distributions, and ran Pavlov against a few variants of itself, as well as against other simple strategies like All-D, All-C and Tit-for-tat.
@@
 
-    \section{Further Work}
+    \section{Future Work}
     \label{sectiondiscussion}
